### 2015-11-01 abstract.tex
Initial commit

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
invariance, i.e., our algorithms make exactly the same decisions if the
sequence of loss vectors is multiplied by any positive constant.  Our
algorithms work for any decision set, bounded or unbounded.  For unbounded
decisions sets, these are the first truly adaptive algorithms for online linear
optimization.

### 2015-11-01 ada-ftrl-proofs.tex
Initial commit

\section{Proofs for AdaFTRL}
\label{section:ada-ftrl-proof}

\begin{proof}[Proof of Corollary~\ref{corollary:ada-ftrl-regret-bound}]
Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:ada-ftrl-regret-bound}
applied to the regularizer $R(w) = \frac{c}{S} f(w)$ and
Proposition~\ref{proposition:diameter-vs-range} gives
$$
\Regret_T \le \sqrt{3}(1 + c) \max\left\{\sqrt{8}, \frac{1}{\sqrt{2c}} \right\} \sqrt{S \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
It remains to find the minimum of $g(c) = \sqrt{3}(1 + c) \max\{\sqrt{8},
1/\sqrt{2c}\}$.  The function $g$ is strictly convex on $(0, \infty)$ and has
minimum at $c=1/16$ and $g(\frac{1}{16}) = \sqrt{3}(1+\frac{1}{16})\sqrt{8} \le
5.3$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
Let $a_t = \|\ell_t\|_* \max\{D, 1/\sqrt{2\lambda}\}$. The statement of the
lemma is equivalent to $\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2}$ which we
prove by induction on $T$.  The base case $T=0$ is trivial. For $T \ge 1$, we
have
$$
\Delta_T
\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
$$
where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
the second inequality from the induction hypothesis and the fact that $f(x) = x
+ \min\{a_T, a_T^2/x\}$ is an increasing function of $x$.  It remains to prove
that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
Dividing through by $a_T$ and making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
$$
z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

### 2015-11-01 ada-ftrl.tex
Initial commit

\section{AdaFTRL}
\label{section:ada-ftrl}

In this section we generalize the AdaHedge
algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
showing that it retains its scale-free property. The analysis is very general
and based on general properties of strongly convex functions, rather than
specific properties of the entropic regularizer like in AdaHedge.

Assume that $K$ is bounded and that $R(w)$ is a strongly convex lower
semi-continuous function bounded from above.  We instantiate
Algorithm~\ref{algorithm:ftrl-varying-regularizer} with the sequence of
regularizers
\begin{equation}
\label{equation:ada-ftrl}
R_t(w) = \Delta_{t-1} R(w)
\quad \text{where}
\quad \Delta_{t}=\sum_{i=1}^{t} \Delta_{i-1} \Breg_{R^*}\left(- \frac{L_i}{\Delta_{i-1}}, -\frac{L_{i-1}}{\Delta_{i-1}}\right) \; .
\end{equation}

The sequence $\{\Delta_t\}_{t=0}^\infty$ is non-negative and non-decreasing.
Also, $\Delta_t$ as a function of $\{\ell_s\}_{s=1}^t$ is positive homogenous
of degree one, making the algorithm scale-free.

If $\Delta_{i-1} = 0$, we define 
$\Delta_{i-1} \Breg_{R^*}(\frac{-L_i}{\Delta_{i-1}}, \frac{-L_{i-1}}{\Delta_{i-1}})$ 
as $\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which always exists and is
finite; see Appendix~\ref{section:limits}.  Similarly, when $\Delta_{t-1} = 0$, we
define $w_t = \argmin_{w \in K} \langle L_{t-1}, w \rangle$ where ties among
minimizers are broken by taking the one with the smallest value of $R(w)$,
which is unique due to strong convexity; this is the same as $w_t = \lim_{a \to
0^+} \argmin_{w \in K} (\langle L_{t-1}, w \rangle + aR(w))$.

Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on
the regret of the algorithm after $T$ rounds, without the need to know before
hand an upper bound on $\|\ell_t\|_*$.  We prove the theorem in
Section~\ref{section:ada-ftrl-regret-bound}.

\begin{theorem}[Regret Bound]
\label{theorem:ada-ftrl-regret-bound}
Suppose $K \subseteq V$ is a non-empty bounded closed convex subset. Let $D =
\sup_{x,y \in K} \|x - y\|$ be its diameter with respect to a norm $\|\cdot\|$.
Suppose that the regularizer $R:K \to \R$ is a non-negative lower
semi-continuous function that is $\lambda$-strongly convex with respect to
$\|\cdot\|$ and is bounded from above.  The regret of AdaFTRL satisfies
$$
\Regret_T(u) \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \left(1 + R(u) \right) \; .
$$
\end{theorem}

The regret bound can be optimized by choosing the optimal multiple of the
regularizer.  Namely, we choose regularizer of the form $\lambda f(w)$ where
$f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
optimization is the following corollary.  Its proof can be found in
Appendix~\ref{section:ada-ftrl-proof}.

\begin{corollary}[Regret Bound]
\label{corollary:ada-ftrl-regret-bound}
Suppose $K \subseteq V$ is a non-empty bounded closed convex subset. Suppose
$f:K \to \R$ is a non-negative lower semi-continuous function that is
$1$-strongly convex with respect to $\|\cdot\|$ and is bounded from above.  The
regret of AdaFTRL with regularizer
$$
R(w) = \frac{f(w)}{16 \cdot \sup_{v \in K} f(v)}
\qquad \text{satisfies} \qquad
\Regret_T \le
5.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{corollary}

\subsection{Proof of Regret Bound for AdaFTRL}
\label{section:ada-ftrl-regret-bound}

\begin{lemma}[Initial Regret Bound]
\label{lemma:initial-regret-bound}
AdaFTRL, for any $u \in K$ and any $u \ge 0$, satisfies
$\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T$.
\end{lemma}

\begin{proof}
Let $R_t(w) = \Delta_{t-1} R(w)$. Since $R$ is non-negative,
$\{R_t\}_{t=1}^\infty$ is non-decreasing.  Hence, $R_t^*(\ell) \ge
R_{t+1}^*(\ell)$ for every $\ell \in V^*$ and thus $R_t^*(-L_t) -
R_{t+1}^*(-L_t) \ge 0$.  So, by Lemma~\ref{lemma:generic-regret-bound},
\begin{equation}
\label{equation:regret-bound-inequality}
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) \; .
\end{equation}
Since, $\Breg_{R_t^*}(u,v) = \Delta_{t-1} \Breg_{R^*}(\frac{u}{\Delta_{t-1}},
\frac{v}{\Delta_{t-1}})$ by definition of Bregman divergence and Part 8 of
Proposition~\ref{proposition:conjugate-properties}, we have $\sum_{t=1}^T
\Breg_{R_t^*}(-L_t, -L_{t-1}) = \Delta_T$.
\end{proof}

\begin{lemma}[Recurrence]
\label{lemma:gap-recurrence}
Let $D = \sup_{u, v \in K} \|u -v\|$ be the diameter of $K$.
The sequence $\{\Delta_t\}_{t=1}^\infty$ generated by AdaFTRL satisfies for any $t \ge 1$,
$$
\Delta_t \le \Delta_{t-1} + \min \left\{ D\|\ell_t\|_*, \ \frac{\|\ell_t\|_*^2}{2\lambda \Delta_{t-1}} \right\} \; .
$$
\end{lemma}

\begin{proof}
The inequality results from strong convexity of $R_t(w)$ and
Proposition~\ref{proposition:conjugate-properties}.
\end{proof}

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
generated by AdaFTRL satisfies for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to
Appendix~\ref{section:ada-ftrl-proof}.
Theorem~\ref{theorem:ada-ftrl-regret-bound} follows from
Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.

### 2015-11-01 introduction.tex
Initial commit

\section{Introduction}
\label{section:introduction}

Online Linear Optimization (OLO) is a problem where an algorithm repeatedly
chooses a point $w_t$ from a convex decision set $K$, observes an arbitrary, or
even adversarially chosen, loss vector $\ell_t$ and suffers loss $\langle
\ell_t, w_t \rangle$.  The goal of the algorithm is to have a small cumulative
loss. Performance of an algorithm is evaluated by the so-called regret, which
is the difference of cumulative losses of the algorithm and of the
(hypothetical) strategy that would choose in every round the same best point in
hindsight.

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006,Shalev-Shwartz-2011}.  Many learning
problems can be directly phrased as OLO, e.g., learning with expert
advice~\cite{Littlestone-Warmuth-1994,Vovk-1998,Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997}, online combinatorial
optimization~\cite{Koolen-Warmuth-Kivinen-2010}. Other
problems can be reduced to OLO, e.g. online convex
optimization~\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification and
regression~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006}, and batch and stochastic
optimization of convex functions~\cite{Nemirovski-Yudin-1983}.  Hence, a result in OLO
immediately implies other results in all these domains.

The adversarial choice of the loss vectors received by the algorithm is what
makes the OLO problem challenging. In particular, if an OLO algorithm commits
to an upper bound on the norm of future loss vectors, its regret can be made
arbitrarily large through an adversarial strategy that produces loss vectors
with norms that exceed the upper bound.

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer. Examples of
such algorithms include the Hedge algorithm or online projected gradient
descent with fixed learning rate.  However, these algorithms have two obvious
drawbacks.

First, they do not come with any regret guarantee for sequences of loss vectors
with norms exceeding $B$. Second, on sequences where the norm of loss vectors
is bounded by $b \ll B$, these algorithms fail to have an optimal regret
guarantee that depends on $b$ rather than on $B$.

% Change spacing between rows.
\renewcommand{\arraystretch}{1.5}

\begin{table}[t]
\fontsize{8}{8.2}\selectfont
\centering
\begin{tabular}{|p{3.6cm}|c|p{3.4cm}|c|}
\hline
\textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
\textsc{Hedge} \cite{Freund-Schapire-1999} & Probability Simplex & Negative Entropy & No \\ \hline
\textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline
\textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad FTRL} \cite{Duchi-Hazan-Singer-2011} & \textbf{Any} & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & No \\ \hline
\textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
\textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \leq C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
\textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func. \newline $1<p\leq2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\end{tabular}
\caption{Selected results for OLO. Best results in each column are in bold.
\label{table:results}
}
\end{table}

\footnotetext{\label{footnote-label} These algorithms attempt to produce an invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a sequence of invariant $w_t$.}

There is a clear practical need to design algorithms that adapt automatically
to norms of the loss vectors.  A natural, yet overlooked, design method to
achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
algorithm.  That is, the sequence of decisions of the algorithm does not change
if the sequence of loss vectors is multiplied by a positive constant.

A summary of algorithms for OLO is presented in Table~\ref{table:results}.
While the scale-free property has been looked at in the expert setting, in the
general OLO setting this issue has been largely ignored.  In particular, the
AdaHedge~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm, for
prediction with expert advice, is specifically designed to be scale-free.  A
notable exception in the OLO literature is the discussion of the ``off-by-one''
issue in~\cite{McMahan-2014}, where it is explained that even the popular AdaGrad
algorithm~\cite{Duchi-Hazan-Singer-2011} is not completely adaptive; see also
our discussion in Section~\ref{section:solo-ftrl}. In particular, existing
scale-free algorithms cover only some norms/regularizers and \emph{only}
bounded decision sets. The case of \textbf{unbounded decision sets},
practically the most interesting one for machine learning applications, remains
completely unsolved.

Rather than trying to design strategies for a particular form of loss
vectors and/or decision sets, in this paper we explicitly focus on the
scale-free property. Regret of scale-free algorithms is proportional to the
scale of the losses, ensuring optimal linear dependency on the maximum norm of
the loss vectors.

The contribution of this paper is twofold. First, in
Section~\ref{section:ada-ftrl} we show that the analysis and design of
AdaHedge can be generalized to the OLO scenario and to any strongly convex
regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new and
rather interesting way to adapt the learning rates to have scale-free
algorithms.  Second, in Section~\ref{section:solo-ftrl} we propose a new and
simple algorithm, \textsc{SOLO FTRL}, that is scale-free and is the
\textbf{first} scale-free online algorithm for unbounded sets with a
non-vacuous regret bound.  Both algorithms are instances of Follow The
Regularized Leader (FTRL) with an adaptive learning rate.  Moreover, our
algorithms show that scale-free algorithms can be obtained in a ``native'' and
simple way, i.e.  without using ``doubling tricks'' that attempt to fix poorly
designed algorithms rather than directly solving the problem.

For both algorithms, we prove that for bounded decision sets the regret after
$T$ rounds is at most $O(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2})$.  We show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving a $\Omega (D
\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} )$ lower bound on the regret of any algorithm
for OLO for any decision set with diameter $D$ with respect to the primal norm
$\|\cdot\|$. For the \textsc{SOLO FTRL} algorithm, we prove an
$O(\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T})$ regret bound for any unbounded
decision set.

Our algorithms are also \textbf{any-time}, i.e., do not need to know the
number of rounds in advance and our regret bounds hold for all time steps
simultaneously.

### 2015-11-01 limits.tex
Initial commit

\section{Limits}
\label{section:limits}

\begin{lemma}
\label{lemma:limit-existence}
Let $K$ be a non-empty bounded closed convex subset of a finite dimensional
normed real vector space $(V, \|\cdot\|)$.  Let $R:K \to \R$ be a strongly
convex lower semi-continuous function bounded from above. Then, for any $x,y
\in V^*$,
$$
\lim_{a \to 0^+} a \Breg_{R^*}(x/a, y/a) = \langle x, u - v \rangle \\
$$
where
\begin{align*}
u = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle x, w \rangle \right)
\quad \text{and} \quad
v = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle y, w \rangle \right) \; .
\end{align*}
\end{lemma}

\begin{proof}
Using Part 3 of Proposition~\ref{proposition:conjugate-properties} we can write
the divergence
\begin{align*}
a \Breg_{R^*}(x/a, y/a) & = a R^*(x/a) - a R^*(y/a) - \langle x - y, \grad R^*(y/a) \rangle \\
& =
 a \left[ \langle x/a, \grad R^*(x/a) \rangle - R(\grad R^*(x/a)) \right] \\
& \qquad - a \left[ \langle y/a, \grad R^*(y/a) \rangle - R(\grad R^*(y/a)) \right]
- \langle x - y, \grad R^*(y/a) \rangle \\
& =
\langle x, \grad R^*(x/a) - \grad R^*(y/a) \rangle - a R(\grad R^*(x/a))
+ a R(\grad R^*(y/a)) \; .
\end{align*}
Part 2 of Proposition~\ref{proposition:conjugate-properties} implies that
\begin{align*}
u = \lim_{a \to 0^+} \grad R^*(x/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle x, w \rangle \right) \; , \\
v = \lim_{a \to 0^+} \grad R^*(y/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle y, w \rangle \right) \; .
\end{align*}
The limits on the right exist because of compactness of $K$. They are simply
the minimizers $u= \argmin_{w \in K} - \langle x, w \rangle$ and $v= \argmin_{w
\in K} - \langle y, w \rangle$ where ties in $\argmin$ are broken according to
smaller value of $R(w)$.

By assumption $R(w)$ is upper bounded. It is also lower bounded, since it is
defined on a compact set and it is lower semi-continuous. Thus,
\begin{align*}
& \lim_{a \to 0^+} a \Breg_{R^*}(x/a, y/a) \\
& = \lim_{a \to 0^+} \langle x, \grad R^*(x/a) - \grad R^*(y/a) \rangle - a R(\grad R^*(x/a)) + a R(\grad R^*(y/a)) \\
& = \lim_{a \to 0^+} \langle x, \grad R^*(x/a) - \grad R^*(y/a) \rangle = \langle x, u - v \rangle \; .
\end{align*}
\end{proof}

### 2015-11-01 lower-bound-proof.tex
Initial commit

\section{Lower Bound Proof}
\label{section:lower-bound-proof}

\begin{proof}[Proof of Theorem~\ref{theorem:simple-lower-bound}]
Pick $x,y \in K$ such that $\|x - y\| = D$. This is possible since $K$ is compact.
Since $\|x - y\| = \sup \{\langle \ell, x - y \rangle ~:~ \ell \in V^*, \|\ell\|_* = 1\}$
and the set $\{ \ell \in V^* ~:~ \|\ell\|_* = 1 \}$ is compact, there exists $\ell \in V^*$
such that
$$
\|\ell\|_* = 1 \qquad \text{and} \qquad \langle \ell, x - y \rangle = \|x - y\| = D \; .
$$
Let $Z_1, Z_2, \dots, Z_T$ be i.i.d. Rademacher variables, that is,
$\Pr[Z_t = +1] = \Pr[Z_t = -1] = 1/2$. Let $\ell_t = Z_t a_t \ell$.
Clearly, $\|\ell_t\|_* = a_t$. The lemma will be proved if we show that
(\ref{equation:simple-lower-bound}) holds with positive probability.
We show a stronger statement that the inequality holds in expectation, i.e.
$\Exp[\Regret_T] \ge \frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T a_t^2}$. Indeed,
\begin{align*}
\Exp\left[ \Regret_T \right]
& \ge \Exp\left[ \sum_{t=1}^T \langle \ell_t, w_t \rangle \right] - \Exp\left[\min_{u \in \{x,y\}} \sum_{t=1}^T \langle \ell_t, u \rangle \right] \\
& = \Exp\left[ \sum_{t=1}^T Z_t a_t \langle \ell, w_t \rangle \right] + \Exp \left[\max_{u \in \{x,y\}} \sum_{t=1}^T -Z_t a_t \langle \ell, u \rangle \right]  \\
& = \Exp\left[ \max_{u \in \{x,y\}} \sum_{t=1}^T -Z_t a_t \langle \ell, u \rangle \right] = \Exp\left[ \max_{u \in \{x,y\}} \sum_{t=1}^T Z_t a_t \langle \ell, u \rangle \right]  \\
& = \frac{1}{2} \Exp\left[ \sum_{t=1}^T Z_t a_t \langle \ell, x + y \rangle \right]  + \frac{1}{2}\Exp\left[ \left|\sum_{t=1}^T Z_t a_t \langle \ell, x - y \rangle \right| \right] \\
& = \frac{D}{2}\Exp\left[ \left|\sum_{t=1}^T Z_t a_t \right| \right] \ge \frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T a_t^2}
\end{align*}
where we used that $\Exp[Z_t] = 0$, the fact that distributions of $Z_t$ and
$-Z_t$ are the same, the formula $\max\{a,b\} = (a+b)/2 + |a-b|/2$, and
Khinchin's inequality in the last step (Lemma A.9 in
\cite{Cesa-Bianchi-Lugosi-2006}).
\end{proof}

### 2015-11-01 lower-bound.tex
Initial commit

\section{Lower Bound}
\label{section:lower-bound}

We show a lower bound on the worst-case regret of any algorithm for OLO. The
proof is a standard probabilistic argument, which we present in
Appendix~\ref{section:lower-bound-proof}.

\begin{theorem}[Lower Bound]
\label{theorem:simple-lower-bound}
Let $K \subseteq V$ be any non-empty bounded closed convex subset. Let $D =
\sup_{u,v \in K} \|u - v\|$ be the diameter of $K$. Let $A$ be any (possibly
randomized) algorithm for OLO on $K$. Let $T$ be any non-negative integer and
let $a_1, a_2, \dots, a_T$ be any non-negative real numbers.  There exists a
sequence of vectors $\ell_1, \ell_2, \dots, \ell_T$ in the dual vector space
$V^*$ such that $\|\ell_1\|_* = a_1, \|\ell_2\|_* = a_2, \dots, \|\ell_T\|_* =
a_T$ and the regret of algorithm $A$ satisfies
\begin{equation}
\label{equation:simple-lower-bound}
\Regret_T \ge \frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \; .
\end{equation}
\end{theorem}

The upper bounds on the regret, which we have proved for our algorithms, have
the same dependency on the norms of loss vectors.  However, a gap remains
between the lower bound and the upper bounds.

Our upper bounds are of the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
\|\ell_t\|_*^2})$ where $f$ is any $1$-strongly convex function with respect to
$\|\cdot\|$.  The same upper bound is also achieved by FTRL with a constant
learning rate when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$
is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.

The gap between $D$ and $\sqrt{\sup_{v \in K} f(v)}$ can be substantial.  For
example, if $K$ is the probability simplex in $\R^d$ and $f(w) = \ln(d) +
\sum_{i=1}^d w_i \ln w_i$ is the shifted negative entropy, the
$\|\cdot\|_1$-diameter of $K$ is $2$, $f$ is non-negative and $1$-strongly
convex w.r.t. $\|\cdot\|_1$, but $\sup_{v \in K} f(v) = \ln(d)$.  On the other
hand, if the norm $\|\cdot\|_2 = \sqrt{\langle \cdot, \cdot \rangle}$ arises
from an inner product $\langle \cdot, \cdot \rangle$, the lower bound matches
the upper bounds within a constant factor.  The reason is that for any $K$ with
$\|\cdot\|_2$-diameter $D$, the function $f(w) = \frac{1}{2} \|w - w_0\|_2^2$,
where $w_0$ is an arbitrary point in $K$, is $1$-strongly convex w.r.t.
$\|\cdot\|_2$ and satisfies that $\sqrt{\sup_{v \in K} f(v)} \le D$. This leads
to the following open problem (posed also in~\cite{Kwon-Mertikopoulos-2014}):
%
\begin{quotation}
\noindent
\emph{Given a bounded convex set $K$ and a norm $\|\cdot\|$, construct a non-negative
function $f:K \to \R$ that is $1$-strongly convex with respect to $\|\cdot\|$
and minimizes $\sup_{v \in K} f(v)$.}
\end{quotation}
%
As shown in~\cite{Srebro-Sridharan-Tewari-2011}, the existence of $f$ with small
$\sup_{v \in K} f(v)$ is equivalent to the existence of an algorithm for OLO with
$\widetilde O(\sqrt{T \sup_{v \in K} f(v)})$ regret assuming $\|\ell_t\|_* \le 1$.
The $\widetilde O$ notation hides a polylogarithmic factor in $T$.

### 2015-11-01 per-coordinate.tex
Initial commit

\section{Per-Coordinate Learning}

An interesting class of algorithms proposed in~\cite{McMahan-Streeter-2010} and
\cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
learning rates.  As shown in \cite{Streeter-McMahan-2010}, our algorithms, or
in fact any algorithm for OLO, can be used with per-coordinate
learning rates as well.

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_i$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(i)}(u_i)$ its regret with respect to $u_i \in K_i$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{i=1}^d \Regret_T^{(i)}(u_i) \; .
$$
If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
each factor $K_i$, we generalize and improve existing regret bounds
\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate
learning rates.

### 2015-11-01 preliminaries-proofs.tex
Initial commit

\section{Proofs for Preliminaries}
\label{section:definitions-proofs}

\begin{proof}[Proof of Proposition~\ref{proposition:diameter-vs-range}]
Let $S = \sup_{u \in K} f(u)$ and $v^* = \argmin_{v \in K} f(v)$. The minimizer
$v^*$ is guaranteed to exist by lower semi-continuity of $f$ and compactness of
$K$.  Optimality condition for $v^*$ and $1$-strong convexity of $f$ imply that
for any $u \in K$,
$$
S
\ge f(u) - f(v^*)
\ge f(u) - f(v^*) - \langle \grad f(v^*), u - v^* \rangle
\ge \frac{1}{2}\|u - v^*\|^2 \; .
$$
In other words, $\|u - v^*\| \le \sqrt{2S}$. By triangle inequality,
$$
D = \sup_{u,v \in K} \|u - v\| \le \sup_{u,v \in K} \left( \|u - v^*\| + \|v^ * - v\| \right) \le 2\sqrt{2S} = \sqrt{8S} \; .
$$
\end{proof}

\begin{proof}[Proof of Property 6 of Proposition~\ref{proposition:conjugate-properties}]
To bound $\Breg_{f^*}(x,y)$ we add a non-negative divergence term $\Breg_{f^*}(y,x)$.
\begin{align*}
\Breg_{f^*}(x, y)
& \le \Breg_{f^*}(x,y) + \Breg_{f^*}(y,x)
= \langle x - y, \grad f^*(x) - \grad f^*(y) \rangle \\
& \le \|x - y\|_* \cdot \| \grad f^*(x) - \grad f^*(y) \|
\le D \|x - y\|_* \; ,
\end{align*}
where we have used H\"older's inequality and Part 7 of the Proposition.
\end{proof}

### 2015-11-01 preliminaries.tex
Initial commit

\section{Notation and Preliminaries}
\label{section:preliminaries}

Let $V$ be a finite-dimensional real vector space equipped with a norm
$\|\cdot\|$. We denote by $V^*$ its dual vector space.  The bi-linear map
associated with $(V^*, V)$ is denoted by $\langle \cdot, \cdot \rangle:V^*
\times V \to \R$.  The dual norm of $\|\cdot\|$ is $\|\cdot\|_*$.

In OLO, in each round $t=1,2,\dots$, the algorithm chooses a point $w_t$ in the
decision set $K \subseteq V$ and then the algorithm observes a loss vector
$\ell_t \in V^*$. The instantaneous loss of the algorithm in round $t$ is
$\langle \ell_t, w_t \rangle$. The cumulative loss of the algorithm after $T$
rounds is $\sum_{t=1}^T \langle \ell_t, w_t \rangle$.  The regret of the
algorithm with respect to a point $u \in K$ is
$$
\Regret_T(u) = \sum_{t=1}^T \langle \ell_t, w_t \rangle - \sum_{t=1}^T \langle \ell_t, u \rangle,
$$
and the regret with respect to the best point is $\Regret_T= \sup_{u \in K}
\Regret_T(u)$.  We assume that $K$ is a non-empty closed convex subset of $V$.
Sometimes we will assume that $K$ is also bounded. We denote by $D$ its
diameter with respect to $\|\cdot\|$, i.e. $D = \sup_{u,v \in K} \|u - v\|$. If
$K$ is unbounded, $D = +\infty$.

\textbf{Convex Analysis.} The \emph{Bregman divergence} of a convex
differentiable function $f$ is defined as $\Breg_f(u,v) = f(u) - f(v) - \langle
\grad f(v), u - v \rangle$.  Note that $\Breg_f(u,v) \ge 0$ for any $u,v$ which
follows directly from the definition of convexity of $f$.

The \emph{Fenchel conjugate} of a function $f:K \to \R$ is the function
$f^*:V^* \to \R \cup \{+\infty\}$ defined as $f^*(\ell) = \sup_{w \in K} \left(
\langle \ell, w \rangle - f(w) \right)$.  The Fenchel conjugate of any function
is convex (since it is a supremum of affine functions) and satisfies for all $w
\in K$ and all $\ell \in V^*$  the \emph{Fenchel-Young inequality}
$
f(w) + f^*(\ell) \ge \langle \ell, w \rangle
$.

Monotonicity of Fenchel conjugates follows easily from the definition: If
$f,g:K \to \R$ satisfy $f(w) \le g(w)$ for all $w \in K$ then $f^*(\ell) \ge
g^*(\ell)$ for every $\ell \in V^*$.

Given $\lambda > 0$, a function $f:K \to \R$ is called \emph{$\lambda$-strongly convex}
with respect to a norm $\|\cdot\|$ if and only if, for all $x,y \in K$,
$$
f(y) \ge f(x) + \langle \grad f(x), y - x \rangle + \frac{\lambda}{2}\|x - y\|^2 \; ,
$$
where $\grad f(x)$ is any subgradient of $f$ at point $x$.

The following proposition relates the range of values of a strongly convex
function to the diameter of its domain. The proof can be found in
Appendix~\ref{section:definitions-proofs}.

\begin{proposition}[Diameter vs. Range]
\label{proposition:diameter-vs-range}
Let $K \subseteq V$ be a non-empty bounded closed convex subset.  Let $D =
\sup_{u,v \in K} \|u - v\|$ be its diameter with respect to $\|\cdot\|$.  Let
$f:K \to \R$ be a non-negative lower semi-continuous function that is
$1$-strongly convex with respect to $\|\cdot\|$.  Then, $D \le \sqrt{8 \sup_{v
\in K} f(v)}$.
\end{proposition}

Fenchel conjugates and strongly convex functions have certain nice properties,
which we list in Proposition~\ref{proposition:conjugate-properties} below.

\begin{proposition}[Fenchel Conjugates of Strongly Convex Functions]
\label{proposition:conjugate-properties}
Let $K \subseteq V$ be a non-empty closed convex set with diameter $D:=\sup_{u,v \in K} \|u-v\|$.
Let $\lambda > 0$, and let $f:K \to \R$ be a lower semi-continuous function
that is $\lambda$-strongly convex with respect to $\|\cdot\|$.
The Fenchel conjugate of $f$ satisfies:
\begin{enumerate}
\item $f^*$ is finite everywhere and differentiable.
\item $\grad f^*(\ell) = \argmin_{w \in K} \left( f(w) - \langle \ell, w \rangle \right)$
\item For any $\ell \in V^*$,
$f^*(\ell) + f(\grad f^*(\ell)) = \langle \ell, \grad f^*(\ell) \rangle$.
\item $f^*$ is $\frac{1}{\lambda}$-strongly smooth i.e. for any $x,y \in V^*$,
$\Breg_{f^*}(x, y) \le \frac{1}{2\lambda} \|x - y\|_*^2$.

\item $f^*$ has $\frac{1}{\lambda}$-Lipschitz continuous gradients i.e.
$\|\grad f^*(x) - \grad f^*(y)\| \le \frac{1}{\lambda} \|x - y\|_*$
for any $x,y \in V^*$.

\item $\Breg_{f^*}(x,y) \le D\|x-y\|_*$ for any $x,y \in V^*$.

\item $\|\grad f^*(x) - \grad f^*(y)\| \le D$ for any $x,y \in V^*$.

\item For any $c > 0$, $(cf(\cdot))^* = cf^*(\cdot/c)$.
\end{enumerate}
\end{proposition}

Except for properties 6 and 7, the proofs can be found
in~\cite{Shalev-Shwartz-2007}.  Property 6 is proven in
Appendix~\ref{section:definitions-proofs}. Property 7 trivially follows from
property 2.

%\label{section:generic-ftrl}

\begin{algorithm}[t]
\caption{\textsc{FTRL with Varying Regularizer}}
\label{algorithm:ftrl-varying-regularizer}
\begin{algorithmic}[1]
\REQUIRE Sequence of regularizers $\{R_t\}_{t=1}^\infty$
\STATE Initialize $L_0 \leftarrow 0$
\FOR{$t=1,2,3,\dots$}
\STATE $w_t \leftarrow \argmin_{w \in K} \left( \langle L_{t-1}, w \rangle + R_t(w) \right)$
\STATE Predict $w_t$
\STATE Observe $\ell_t \in V^*$
\STATE $L_t \leftarrow L_{t-1} + \ell_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}

\textbf{Generic FTRL with Varying Regularizer.}
Our scale-free online learning algorithms are versions of the \textsc{Follow
The Regularized Leader} (FTRL) algorithm with varying regularizers, presented
as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The following lemma
bounds its regret.

\begin{lemma}[Lemma 1 in \cite{Orabona-Crammer-Cesa-Bianchi-2014}]
\label{lemma:generic-regret-bound}
For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
semi-continuous regularizers, regret of
Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}
%
The lemma allows data dependent regularizers. That is, $R_t$ can depend on the past loss
vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$.

### 2015-11-01 solo-ftrl-proofs.tex
Initial commit

\section{Proofs for SOLO FTRL}
\label{section:solo-ftrl-proof}

\begin{proof}[Proof of Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}]
Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:regret-solo-ftrl} applied
to the regularizer $R(w) = \frac{c}{\sqrt{S}} f(w)$, together with
Proposition~\ref{proposition:diameter-vs-range} and a crude bound
$\max_{t=1,2,\dots,T} \|\ell_t\|_* \le \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$,
give
$$
\Regret_T \le \left(c + \frac{2.75}{c}  + 3.5\sqrt{8} \right) \sqrt{S \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
$$
We choose $c$ by minimizing $g(c) = c + \frac{2.75}{c} + 3.5\sqrt{8}$. Clearly,
$g(c)$ has minimum at $c = \sqrt{2.75}$ and has minimal value $g(\sqrt{2.75}) =
2\sqrt{2.75} + 3.5\sqrt{8} \le 13.3$.
\end{proof}

### 2015-11-01 solo-ftrl.tex
Initial commit

\section{SOLO FTRL}
\label{section:solo-ftrl}

The closest algorithm to a scale-free one in the OLO literature is the AdaGrad
algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a regularizer on each
coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, AdaGrad is not fully adaptive to the norm of
the loss vectors. Identical considerations hold for the FTRL-Proximal in
\cite{McMahan-Streeter-2010,McMahan-2014}: the scale-free setting of the learning rate is
valid only in the bounded case.

One simple approach would be to use a doubling trick on $\delta$ in order to
estimate on the fly the maximum norm of the losses. Note that a naive strategy
would still fail because the initial value of $\delta$ should be data-dependent
in order to have a scale-free algorithm. Moreover, we would have to upper bound
the regret in all the rounds where the norm of the current loss is bigger than
the estimate. Finally, the algorithm would depend on an additional parameter,
the ``doubling'' power. Hence, even guaranteeing a regret bound\footnote{For
lack of space, we cannot include the regret bound for the doubling trick
version. It would be exactly the same as in
Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
the additional parameter of the doubling power.}, such strategy would give the
feeling that FTRL needs to be ``fixed'' in order to obtain a scale-free
algorithm.

In the following, we propose a much simpler and better approach.  We propose to
use Algorithm~\ref{algorithm:ftrl-varying-regularizer} with the regularizer
\begin{equation*}
R_t(w) = R(w) \sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2} \; ,
\end{equation*}
where $R:K \to \R$ is any strongly convex function. Through a refined analysis,
we show that the regularizer suffices to obtain an optimal regret bound for any
decision set, bounded or unbounded.  We call such variant \textsc{Scale-free
Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
result is the following Theorem, which is proven in
Section~\ref{section:solo-ftrl-regret-bound}.

\begin{theorem}[Regret of \textsc{SOLO FTRL}]
\label{theorem:regret-solo-ftrl}
Suppose $K \subseteq V$ is a non-empty closed convex subset.  Let $D =
\sup_{u,v \in K} \|u - v\|$ be its diameter with respect to a norm $\|\cdot\|$.
Suppose that the regularizer $R:K \to \R$ is a non-negative lower
semi-continuous function that is $\lambda$-strongly convex with respect to
$\|\cdot\|$. The regret of SOLO FTRL satisfies
\begin{align*}
\Regret_T(u)
& \le \left( R(u) + \frac{2.75}{\lambda}\right) \sqrt{\sum_{t=1}^{T} \norm{\ell_t}_*^2}
+ 3.5 \min\left\{\frac{\sqrt{T-1}}{\lambda} , D\right\} \max_{t \le T} \|\ell_t\|_*.
\end{align*}
\end{theorem}

When $K$ is bounded, we can choose the optimal multiple of the regularizer.  We
choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
optimize $\lambda$.  The result of the optimization is
Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is in
Appendix~\ref{section:solo-ftrl-proof}.  It is similar to
Corollary~\ref{corollary:ada-ftrl-regret-bound} for AdaFTRL. The scaling
however is different in the two corollaries.  In
Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in
K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we
have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.

\begin{corollary}[Regret Bound for Bounded Decision Sets]
\label{corollary:regret-solo-ftrl-bounded-set}
Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
that $f:K \to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
convex with respect to $\|\cdot\|$. SOLO FTRL with regularizer
$$
R(w) = \frac{f(w)\sqrt{2.75}}{\sqrt{\sup_{v \in K} f(v)}}
\quad \text{satisfies} \quad %
\Regret_T \le 13.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
$$
\end{corollary}

\subsection{Proof of Regret Bound for SOLO FTRL}
\label{section:solo-ftrl-regret-bound}

The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
(Lemma~\ref{lemma:useful}).  Related and weaker inequalities were proved by
~\cite{Auer-Cesa-Bianchi-Gentile-2002} and ~\cite{Jaksch-Ortner-Auer-2010}.
The main property of this inequality is that on the right-hand side $C$ does
\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.  We will also use the
well-known technical Lemma~\ref{lemma:sum-of-square-roots-inverses}.

\begin{lemma}[Useful Inequality]
\label{lemma:useful}
Let $C, a_1, a_2, \dots, a_T\geq0$. Then,
$$
\sum_{t=1}^T \min \left\{ a_t^2 / \sqrt{\sum_{s=1}^{t-1} a_s^2}, \ C a_t \right\}
\le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
\begin{proof}
Without loss of generality, we can assume that $a_t > 0$ for all $t$. Since otherwise we
can remove all $a_t = 0$ without affecting either side of the inequality. Let $M_t = \max\{a_1, a_2, \dots, a_t\}$ and $M_0 = 0$.
We prove that for any $\alpha > 1$
$$
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le 2 \sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) + \frac{C\alpha( M_t  - M_{t-1})}{\alpha - 1}
$$
from which the inequality follows by summing over $t=1,2,\dots,T$ and choosing $\alpha = \sqrt{2}$.
The inequality follows by case analysis. If $a_t^2 \le \alpha^2 \sum_{s=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}
= \frac{a_t^2}{\sqrt{\frac{1}{1+\alpha^2} \left( \alpha^2 \sum_{s=1}^{t-1} a_s^2 + \sum_{s=1}^{t-1} a_s^2 \right)}} \\
\le \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{ a_t^2 + \sum_{s=1}^{t-1} a_s^2 }}
= \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{\sum_{s=1}^t a_s^2}}
\le 2\sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\end{multline*}
where we have used $x^2/\sqrt{x^2+y^2} \le 2(\sqrt{x^2+y^2} - \sqrt{y^2})$ in the last step.
On the other hand, if $a_t^2 > \alpha^2 \sum_{t=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le C a_t
= C \frac{\alpha a_t  - a_t}{\alpha - 1}
\le \frac{C}{\alpha - 1} \left( \alpha a_t  - \alpha \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) \\
= \frac{C\alpha}{\alpha - 1} \left( a_t  - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\le \frac{C\alpha}{\alpha - 1} \left( a_t  - M_{t-1} \right)
= \frac{C\alpha}{\alpha - 1} \left( M_t  - M_{t-1} \right)
\end{multline*}
where we have used that $a_t = M_t$ and $\sqrt{\sum_{s=1}^{t-1} a_s^2} \ge M_{t-1}$.
\end{proof}

\begin{lemma}[Lemma~3.5 in \cite{Auer-Cesa-Bianchi-Gentile-2002}]
\label{lemma:sum-of-square-roots-inverses}
Let $a_1, a_2, \dots, a_T$ be non-negative real numbers. If $a_1 > 0$ then,
$$
\sum_{t=1}^T a_t / \sqrt{\sum_{s=1}^t a_s} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
$$
\end{lemma}

\begin{proof}[Proof of Theorem~\ref{theorem:regret-solo-ftrl}]
Let $\eta_t=\tfrac{1}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}$, hence $R_t(w) = \tfrac{1}{\eta_t} R(w)$.
We assume without loss of
generality that $\|\ell_t\|_* > 0$ for all $t$, since otherwise we can remove
all rounds $t$ where $\ell_t = 0$ without affecting regret and the
predictions of the algorithm on the remaining rounds.
By Lemma~\ref{lemma:generic-regret-bound},
\begin{align*}
\Regret_T(u)
& \le \frac{1}{\eta_{T+1}} R(u) + \sum_{t=1}^T \left( \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \right) \; .
\end{align*}
We upper bound the terms of the sum in two different ways.
First, by Proposition~\ref{proposition:conjugate-properties}, we have
$$
\Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t)
\le \Breg_{R_t^*}(-L_t, -L_{t-1})
\le \frac{\eta_t \|\ell_t\|_*^2}{2\lambda} \; .
$$
Second, we have
\begin{align*}
& \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \\
& = \Breg_{R_{t+1}^*}(-L_t, -L_{t-1}) + R^*_{t+1}(-L_{t-1}) - R_t^*(-L_{t-1}) \\
& \qquad + \langle \nabla R_t^*(-L_{t-1})-\nabla R_{t+1}^*(-L_{t-1}), \ell_t \rangle  \\
& \le \tfrac{1}{2\lambda} \eta_{t+1} \|\ell_t\|_*^2 + \| \nabla R_t^*(-L_{t-1})-\nabla R_{t+1}^*(-L_{t-1})\| \cdot \|\ell_t\|_* \\
& = \tfrac{1}{2\lambda} \eta_{t+1} \|\ell_t\|_*^2 + \| \nabla R^*(- \eta_{t} L_{t-1})-\nabla R^*(- \eta_{t+1} L_{t-1})\| \cdot \|\ell_t\|_* \\
& \le \frac{\eta_{t+1} \|\ell_t\|_*^2}{2\lambda} + \min\left\{\frac{1}{\lambda} \|L_{t-1}\|_* \left(\eta_{t} - \eta_{t+1} \right), D\right\} \|\ell_t\|_* \; ,
\end{align*}
where in the first inequality we have used the fact that $R^*_{t+1}(-L_{t-1})
\le R_t^*(-L_{t-1})$, H\"older's inequality, and
Proposition~\ref{proposition:conjugate-properties}.  In the second inequality
we have used properties 5 and 7 of
Proposition~\ref{proposition:conjugate-properties}. Using the definition of
$\eta_{t+1}$ we have
\begin{align*}
\frac{\|L_{t-1}\|_* (\eta_{t} -\eta_{t+1})}{\lambda}
\le \frac{ \|L_{t-1}\|_*}{\lambda \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}
\le \frac{ \sum_{i=1}^{t-1} \|\ell_i\|_*}{\lambda \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}
\le \frac{\sqrt{t-1}}{\lambda}
\le \frac{\sqrt{T-1}}{\lambda}.
\end{align*}
Denoting by $H=\min\left\{\frac{\sqrt{T-1}}{\lambda},D\right\}$ we have
\begin{align*}
& \Regret_T(u)
\le \frac{1}{\eta_{T+1}} R(u) + \sum_{t=1}^T \min\left\{ \frac{\eta_t \|\ell_t\|_*^2}{2\lambda}, \ H \|\ell_t\|_* + \frac{\eta_{t+1} \|\ell_t\|_*^2}{2\lambda}  \right\} \\
& \le \frac{1}{\eta_{T+1}} R(u) + \frac{1}{2\lambda} \sum_{t=1}^T  \eta_{t+1} \|\ell_t\|_*^2 + \frac{1}{2\lambda} \sum_{t=1}^T \min\left\{ \eta_t \|\ell_t\|_*^2, \ 2 \lambda H \|\ell_t\|_* \right\} \\
& = \frac{1}{\eta_{T+1}} R(u) + \frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^t \|\ell_t\|_*^2}} + \frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}, \ 2 \lambda H \|\ell_t\|_* \right\} \; .
\end{align*}
We bound each of the three terms separately. By definition of $\eta_{T+1}$, the
first term is $\frac{1}{\eta_{T+1}} R(u) = R(u) \sqrt{\sum_{t=1}^T
\|\ell_t\|_*^2}$.  We upper bound the second term using
Lemma~\ref{lemma:sum-of-square-roots-inverses} as
$$
\frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^t \|\ell_t\|_*^2}}
\le \frac{1}{\lambda} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
Finally, by Lemma~\ref{lemma:useful} we upper bound the third term as
$$
\frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}, \ 2 \lambda \|\ell_t\|_* H \right\}
\le 3.5 H \max_{t \le T} \|\ell_t\|_* + \frac{1.75}{\lambda} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
Putting everything together gives the stated bound.
\end{proof}

### 2015-11-20 preliminaries.tex
Few improvements

@@
 
-\textbf{Convex Analysis.} The \emph{Bregman divergence} of a convex
-differentiable function $f$ is defined as $\Breg_f(u,v) = f(u) - f(v) - \langle
-\grad f(v), u - v \rangle$.  Note that $\Breg_f(u,v) \ge 0$ for any $u,v$ which
-follows directly from the definition of convexity of $f$.
+\subsection{Convex Analysis}
+
+The \emph{Bregman divergence} of a convex differentiable function $f$ is
+defined as $\Breg_f(u,v) = f(u) - f(v) - \langle \grad f(v), u - v \rangle$.
+Note that $\Breg_f(u,v) \ge 0$ for any $u,v$ which follows directly from the
+definition of convexity of $f$.
 
@@
 \langle \ell, w \rangle - f(w) \right)$.  The Fenchel conjugate of any function
-is convex (since it is a supremum of affine functions) and satisfies for all $w
-\in K$ and all $\ell \in V^*$  the \emph{Fenchel-Young inequality}
-$
-f(w) + f^*(\ell) \ge \langle \ell, w \rangle
-$.
+is convex (since it is a supremum of affine functions) and satisfies
+ the \emph{Fenchel-Young inequality}
+$$
+\text{$\forall w \in K$, $\forall \ell \in V^*$} \qquad
+f(w) + f^*(\ell) \ge \langle \ell, w \rangle \; .
+$$
 
@@
 \begin{enumerate}
+
 \item $f^*$ is finite everywhere and differentiable.
-\item $\grad f^*(\ell) = \argmin_{w \in K} \left( f(w) - \langle \ell, w \rangle \right)$
-\item For any $\ell \in V^*$,
-$f^*(\ell) + f(\grad f^*(\ell)) = \langle \ell, \grad f^*(\ell) \rangle$.
-\item $f^*$ is $\frac{1}{\lambda}$-strongly smooth i.e. for any $x,y \in V^*$,
-$\Breg_{f^*}(x, y) \le \frac{1}{2\lambda} \|x - y\|_*^2$.
+
+\item For any $\ell \in V^*$, $\grad f^*(\ell) = \argmin_{w \in K} \left( f(w) - \langle \ell, w \rangle \right)$
+
+\item For any $\ell \in V^*$, $f^*(\ell) + f(\grad f^*(\ell)) = \langle \ell, \grad f^*(\ell) \rangle$.
+
+\item $f^*$ is $\frac{1}{\lambda}$-strongly smooth i.e. for any $x,y \in V^*$, $\Breg_{f^*}(x, y) \le \frac{1}{2\lambda} \|x - y\|_*^2$.
 
 \item $f^*$ has $\frac{1}{\lambda}$-Lipschitz continuous gradients i.e.
-$\|\grad f^*(x) - \grad f^*(y)\| \le \frac{1}{\lambda} \|x - y\|_*$
-for any $x,y \in V^*$.
+for any $x,y \in V^*$,
+$\|\grad f^*(x) - \grad f^*(y)\| \le \frac{1}{\lambda} \|x - y\|_*$.
 
@@
 
-%\label{section:generic-ftrl}
-
 \begin{algorithm}[t]
@@
 \begin{algorithmic}[1]
-\REQUIRE Sequence of regularizers $\{R_t\}_{t=1}^\infty$
 \STATE Initialize $L_0 \leftarrow 0$
 \FOR{$t=1,2,3,\dots$}
+\STATE Choose a regularizer $R_t:K \to \R$
 \STATE $w_t \leftarrow \argmin_{w \in K} \left( \langle L_{t-1}, w \rangle + R_t(w) \right)$
@@
 
-\textbf{Generic FTRL with Varying Regularizer.}
+\subsection{Generic FTRL with Varying Regularizer}
+\label{section:generic-ftrl}
+
 Our scale-free online learning algorithms are versions of the \textsc{Follow
@@
 \end{lemma}
-%
-The lemma allows data dependent regularizers. That is, $R_t$ can depend on the past loss
-vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$.
+
+\begin{proof}
+
+\end{proof}

### 2015-11-20 mirror-descent.tex

\section{Mirror Descent}
\label{section:mirror-descent}

\begin{algorithm}[t]
\caption{\textsc{Mirror Descent with Varying Regularizer}}
\label{algorithm:mirror-descent-varying-regularizer}
\begin{algorithmic}[1]
\REQUIRE Non-empty closed convex set $K \subseteq V$
\STATE Choose a regularizer $R_0:K \to \R$
\STATE $w_1 \leftarrow \argmin_{w \in K} R_0(w)$
\FOR{$t=1,2,3,\dots$}
\STATE Predict $w_t$
\STATE Observe $\ell_t \in V^*$
\STATE Choose a regularizer $R_t:K \to \R$
\STATE $w_{t+1} \leftarrow \argmin_{w \in K} \left( \langle \ell, w \rangle + \Breg_{R_t}(w, w_t) \right)$
\ENDFOR
\end{algorithmic}
\end{algorithm}

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
\to \R$ where each $R_t$ can depend on $\ell_1, \ell_2, \dots, \ell_t$ in an
arbitrary way. If $R_t$ is not differentiable\footnote{Note that this can
happen even when $R_t$ is a restriction of a differentiable function defined on
a superset of $K$.  If $K$ is bounded and closed, $R_t$ fails to be
differentiable at the boundary of $K$. If $K$ is a subset of an affine subspace
of a dimension smaller than then dimension of $V$, then $R_t$ fails to be
differentiable everywhere.}, to define Bregman divergence $\Breg_{R_t}$ the
algorithm has to choose together with $R_t$ a subgradient map $\grad R_t:K \to
V$.  If $R_t$ is a restriction of a differentiable function, it is convenient
to take $\grad R_t$ as the gradient of of that function.

Several existing algorithms are special cases of \textsc{Mirror Descent}.
Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (GIGA)
algorithm~\cite{Zinkevich-2003} is an instance of \textsc{Mirror Descent} for
any bounded closed convex $K$ with regularizer
$$
R_t(w) = \frac{\sqrt{t}}{2}\|w\|_2^2 \; .
$$
Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
\textsc{Mirror Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2} \; .
$$
respectively. Notice that when $d=1$, both algorithms are equivalent to
$$
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
$$

Third, Exponentially-Weighted Average Forecaster~\cite[Chapter
2]{Cesa-Bianchi-Lugosi-2006} with constant learning rate $\eta$ for the problem
of prediction with expert advice is an instance of \textsc{Mirror Descent} for
the probability simplex $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0
\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{i=1}^d w_i
\ln w_i$ where $R_t:K \to \R$.

Scale-free version of the algorithm \textsc{Mirror Descent} can be obtained
by using regularizers of the form
$$
\R_t(w) = R(w) \sqrt{\sum_{i=1}^t }
$$

### 2015-11-21 mirror-descent.tex
Mirror descent section

@@
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
-parametrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
-\to \R$ where each $R_t$ can depend on $\ell_1, \ell_2, \dots, \ell_t$ in an
+parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
+\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on
+past loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an
 arbitrary way. If $R_t$ is not differentiable\footnote{Note that this can
@@
 differentiable at the boundary of $K$. If $K$ is a subset of an affine subspace
-of a dimension smaller than then dimension of $V$, then $R_t$ fails to be
-differentiable everywhere.}, to define Bregman divergence $\Breg_{R_t}$ the
+of a dimension smaller than the dimension of $V$, then $R_t$ fails to be
+differentiable everywhere.}, to define Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v \rangle$, the
 algorithm has to choose together with $R_t$ a subgradient map $\grad R_t:K \to
-V$.  If $R_t$ is a restriction of a differentiable function, it is convenient
-to take $\grad R_t$ as the gradient of of that function.
+V$.  If $R_t$ is a restriction of a differentiable function $R'_t$, it is convenient
+to define $\grad R_t(w) = \grad R'_t(w)$ for all $w \in K$. The following lemma
+bounds instantaneous regret of \textsc{Mirror Descent}.
 
-Several existing algorithms are special cases of \textsc{Mirror Descent}.
-Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (GIGA)
-algorithm~\cite{Zinkevich-2003} is an instance of \textsc{Mirror Descent} for
-any bounded closed convex $K$ with regularizer
+\begin{lemma}[Regret of Mirror Descent]
+Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} for any $u \in K$ satisfies
 $$
-R_t(w) = \frac{\sqrt{t}}{2}\|w\|_2^2 \; .
+\Regret_T(u) \le \sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
 $$
-Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
+\end{lemma}
+
+Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
+prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
+- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
+$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
+\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
+A special case of projected gradient descent with step size $\eta_t = \frac{1}{\sqrt{t}}$
+is Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (GIGA)
+algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
 Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
@@
 $$
-respectively. Notice that when $d=1$, both algorithms are equivalent to
-$$
+respectively. When $d=1$, both algorithms are equivalent to
+\begin{equation}
+\label{equation:ada-grad-regularizer}
 R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
-$$
+\end{equation}
+All three sequences of regularizers give a scale-free algorithm.
+Also note that regularizer \eqref{equation:ada-grad} can be viewed
+as a scale-free generalization of Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
+\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer} coincides with GIGA.
 
-Third, Exponentially-Weighted Average Forecaster~\cite[Chapter
+\textsc{Exponentially-Weighted Average Forecaster}~\cite[Chapter
 2]{Cesa-Bianchi-Lugosi-2006} with constant learning rate $\eta$ for the problem
@@
 \}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{i=1}^d w_i
-\ln w_i$ where $R_t:K \to \R$.
+\ln w_i$ where $R_t:K \to \R$. Note that this algorithm is \emph{not} scale-free.
 
-Scale-free version of the algorithm \textsc{Mirror Descent} can be obtained
-by using regularizers of the form
+Replacing $\frac{1}{2}\|w\|_2^2$ in equation $\label{equation:ada-grad}$
+with an arbitrary convex function $R:K \to \R$ yields
+\begin{equation}
+\label{equation:scale-free-mirror-descent}
+R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
+\end{equation}
+We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with regularizer $R$.
+Generalizing existing regret bound for
+\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
+
+\begin{theorem}[Regret of Scale-Free Mirror Descent]
+\label{theorem:regret-scale-free-mirror-descent}
+Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose
+that $R:K \to \R$ is a $\lambda$-strongly convex function with respect to a norm $\|\cdot\|$.
+\textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
 $$
-\R_t(w) = R(w) \sqrt{\sum_{i=1}^t }
+\Regret_T(u) \le \left( \frac{1}{\lambda} + \sup_{v \in K} \Breg_R(u,v) \right) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
 $$
+\end{theorem}
+
+We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function
+and optimize $\lambda$.
+
+\begin{corollary}[Regret of Scale-Free Mirror Descent]
+\label{corollary:regret-scale-free-mirror-descent}
+Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
+that $f:K \to \R$ is a $1$-strongly convex function with respect to a norm $\|\cdot\|$.
+\textsc{Scale-Free Mirror Descent} with regularizer
+$$
+R(w) = \frac{f(w)}{\displaystyle \sqrt{\sup_{u,v \in K} \Breg_f(u,v)}}
+\quad \text{satisfies} \quad %
+\Regret_T(u) \le 2 \sqrt{\sup_{u,v \in K} \Breg_f(u,v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
+$$
+\end{corollary}
+
+\subsection{Lower Bounds}
+
+The bounds in Theorem~\ref{theorem:regret-scale-free-mirror-descent} and
+Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
+$\Breg_R(u,v)$ is not bounded. One might wonder if the
+assumption that $\Breg_R(u,v)$ is bounded is necessary. We show necessity of
+this assumption on two counter-examples where $\Breg_R(u,v)$ is not bounded.
+In these counter-examples, we construct sequences of loss vectors $\ell_1,
+\ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots =
+\|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$
+or worse regret.
+
+The first counter-example is stated as
+Theorem~\ref{theorem:first-counter-example} below. The decision set is the
+whole space $K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note
+that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
+norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. Corresponding Bregman divergence is
+$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
+sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
+the first vector of the standard unit basis.  On such sequence, Zinkevich's
+\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror Update} have
+predictions identical to \textsc{Scale-Free Mirror Descent}. Hence the lower
+bound applies to all four algorithms.
+
+\begin{theorem}[First Counter-Example]
+\label{theorem:first-counter-example}
+Suppose $K = V$. For any $T \ge 42$ there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
+such that $\|\ell_1\|_2 = \|\ell_2\|_2 = \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free Mirror Descent}
+with regularizer $R(w) = \frac{1}{2}\|w\|_w^2$, Zinkevich's \textsc{GIGA} algorithm, and
+both versions of \textsc{AdaGrad with Composite Mirror Descent Update} satisfy
+$$
+\Regret_T(0) \ge \frac{T^{3/2}}{20} \; .
+$$
+\end{theorem}
+
+The second counter-example is as Theorem~\ref{theorem:second-counter-example}
+below.  The decision set is the $d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_i
+\ge 0, \sum_{i=1}^d w_i = 0 \}$ and the regularizer is the negative entropy
+$R(w) = \sum_{i=1}^d w_i \ln w_i$.  Negative entropy is $1$-strongly convex
+with respect to $\|\cdot\|_1$ and the dual norm of $\|\cdot\|_1$ is
+\mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman divergence is the
+Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{i=1}^d u_i \ln(u_i/v_i)$.
+Note that despite that negative entropy is upper- and lower-bounded,
+Kullback-Leibler divergence is not.
+
+\begin{theorem}[Second Counter-Example]
+\label{theorem:second-counter-example}
+Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \} \subseteq V$ be the $d$-dimensional
+probability simplex.
+For any $T \ge 120$ there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
+such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty = 1$ and \textsc{Scale-Free Mirror Descent}
+with regularizer $R(w) = \sum_{i=1}^d w_i \ln w_i$ satisfies
+$$
+\Regret_T \ge \frac{T}{6} \; .
+$$
+\end{theorem}

### 2015-11-22 mirror-descent.tex
Few iprovements to Mirror Descent chapter.

@@
 parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
-\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on
-past loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an
-arbitrary way. If $R_t$ is not differentiable\footnote{Note that this can
-happen even when $R_t$ is a restriction of a differentiable function defined on
-a superset of $K$.  If $K$ is bounded and closed, $R_t$ fails to be
-differentiable at the boundary of $K$. If $K$ is a subset of an affine subspace
-of a dimension smaller than the dimension of $V$, then $R_t$ fails to be
-differentiable everywhere.}, to define Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v \rangle$, the
-algorithm has to choose together with $R_t$ a subgradient map $\grad R_t:K \to
-V$.  If $R_t$ is a restriction of a differentiable function $R'_t$, it is convenient
-to define $\grad R_t(w) = \grad R'_t(w)$ for all $w \in K$. The following lemma
-bounds instantaneous regret of \textsc{Mirror Descent}.
+\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
+loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
+not differentiable\footnote{Note that this can happen even when $R_t$ is a
+restriction of a differentiable function defined on a superset of $K$.  If $K$
+is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
+If $K$ is a subset of an affine subspace of a dimension smaller than the
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, to define
+Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
+\rangle$, the algorithm has to choose together with $R_t$ a subgradient map
+$\grad R_t:K \to V$.  If $R_t$ is a restriction of a differentiable function
+$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
+\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.
 
 \begin{lemma}[Regret of Mirror Descent]
-Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} for any $u \in K$ satisfies
+Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
+$u \in K$,
 $$
-\Regret_T(u) \le \sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
+\Regret_T(u)
+\le
+\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
 $$
@@
 \textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
-A special case of projected gradient descent with step size $\eta_t = \frac{1}{\sqrt{t}}$
-is Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (GIGA)
-algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
-Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
-\textsc{Mirror Descent} for any bounded $K$ with regularizers
+A special case of projected gradient descent with step size $\eta_t =
+\frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized Infinitesimal Gradient
+Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal
+versions of \textsc{AdaGrad with Composite Mirror Descent
+Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{Mirror
+Descent} for any bounded $K$ with regularizers
 $$
@@
 \end{equation}
-All three sequences of regularizers give a scale-free algorithm.
-Also note that regularizer \eqref{equation:ada-grad} can be viewed
-as a scale-free generalization of Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
-\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer} coincides with GIGA.
+All three sequences of regularizers give a scale-free algorithm. Regularizer
+\eqref{equation:ada-grad} can be viewed as a scale-free generalization of
+Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
+\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
+coincides with GIGA.
 
@@
 \}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{i=1}^d w_i
-\ln w_i$ where $R_t:K \to \R$. Note that this algorithm is \emph{not} scale-free.
+\ln w_i$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
+scale-free.
 
-Replacing $\frac{1}{2}\|w\|_2^2$ in equation $\label{equation:ada-grad}$
-with an arbitrary convex function $R:K \to \R$ yields
+Replacing $\frac{1}{2}\|w\|_2^2$ in equation $\label{equation:ada-grad}$ with
+an arbitrary convex function $R:K \to \R$ yields
 \begin{equation}
@@
 \end{equation}
-We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with regularizer $R$.
-Generalizing existing regret bound for
+We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
+regularizer $R$.  Generalizing existing regret bound for
 \textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
@@
 \label{theorem:regret-scale-free-mirror-descent}
-Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose
-that $R:K \to \R$ is a $\lambda$-strongly convex function with respect to a norm $\|\cdot\|$.
-\textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
+Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose that $R:K
+\to \R$ is a $\lambda$-strongly convex function with respect to a norm
+$\|\cdot\|$.  \textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
+for any $u \in K$,
 $$
@@
 
-We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function
-and optimize $\lambda$.
+We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
+function and optimize $\lambda$. The result is the following Corollary. Its proof is trivial.
 
@@
 Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
-that $f:K \to \R$ is a $1$-strongly convex function with respect to a norm $\|\cdot\|$.
-\textsc{Scale-Free Mirror Descent} with regularizer
+that $f:K \to \R$ is a $1$-strongly convex function with respect to a norm
+$\|\cdot\|$.  \textsc{Scale-Free Mirror Descent} with regularizer
 $$
@@
 Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
-$\Breg_R(u,v)$ is not bounded. One might wonder if the
-assumption that $\Breg_R(u,v)$ is bounded is necessary. We show necessity of
-this assumption on two counter-examples where $\Breg_R(u,v)$ is not bounded.
-In these counter-examples, we construct sequences of loss vectors $\ell_1,
-\ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots =
-\|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$
-or worse regret.
+$\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
+$\Breg_R(u,v)$ is bounded is necessary. We show necessity of this assumption on
+two counter-examples.  In these
+counter-examples, we consider strongly convex regularizers $R$ such that
+$\Breg_R(u,v)$ is not bounded and we construct sequences of loss vectors
+$\ell_1, \ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots
+= \|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$ or
+worse regret.
 
@@
 the first vector of the standard unit basis.  On such sequence, Zinkevich's
-\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror Update} have
-predictions identical to \textsc{Scale-Free Mirror Descent}. Hence the lower
-bound applies to all four algorithms.
+\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror
+Update} have predictions identical to \textsc{Scale-Free Mirror Descent}. Hence
+the lower bound applies to all four algorithms.
 
@@
 \label{theorem:first-counter-example}
-Suppose $K = V$. For any $T \ge 42$ there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
-such that $\|\ell_1\|_2 = \|\ell_2\|_2 = \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free Mirror Descent}
-with regularizer $R(w) = \frac{1}{2}\|w\|_w^2$, Zinkevich's \textsc{GIGA} algorithm, and
-both versions of \textsc{AdaGrad with Composite Mirror Descent Update} satisfy
+Suppose $K = V$. For any $T \ge 42$ there exists a sequence of loss vectors
+$\ell_1, \ell_2, \dots, \ell_T \in V^*$ such that $\|\ell_1\|_2 = \|\ell_2\|_2
+= \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free Mirror Descent} with
+regularizer $R(w) = \frac{1}{2}\|w\|_w^2$, Zinkevich's \textsc{GIGA} algorithm,
+and both versions of \textsc{AdaGrad with Composite Mirror Descent Update}
+satisfy
 $$
@@
 The second counter-example is as Theorem~\ref{theorem:second-counter-example}
-below.  The decision set is the $d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_i
-\ge 0, \sum_{i=1}^d w_i = 0 \}$ and the regularizer is the negative entropy
-$R(w) = \sum_{i=1}^d w_i \ln w_i$.  Negative entropy is $1$-strongly convex
-with respect to $\|\cdot\|_1$ and the dual norm of $\|\cdot\|_1$ is
-\mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman divergence is the
-Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{i=1}^d u_i \ln(u_i/v_i)$.
-Note that despite that negative entropy is upper- and lower-bounded,
-Kullback-Leibler divergence is not.
+below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
+\in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}$ and the regularizer is the
+negative entropy $R(w) = \sum_{i=1}^d w_i \ln w_i$.  Negative entropy is
+$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
+$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
+divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{i=1}^d u_i
+\ln(u_i/v_i)$.  Note that despite that negative entropy is upper- and
+lower-bounded, Kullback-Leibler divergence is not.
 
@@
 \label{theorem:second-counter-example}
-Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \} \subseteq V$ be the $d$-dimensional
-probability simplex.
-For any $T \ge 120$ there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
-such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty = 1$ and \textsc{Scale-Free Mirror Descent}
-with regularizer $R(w) = \sum_{i=1}^d w_i \ln w_i$ satisfies
+Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}
+\subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
+there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
+such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
+1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{i=1}^d
+w_i \ln w_i$ satisfies
 $$

### 2015-11-23 ada-ftrl.tex
Added proofs for mirror descent section.

@@
 
-If $\Delta_{i-1} = 0$, we define 
-$\Delta_{i-1} \Breg_{R^*}(\frac{-L_i}{\Delta_{i-1}}, \frac{-L_{i-1}}{\Delta_{i-1}})$ 
-as $\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which always exists and is
-finite; see Appendix~\ref{section:limits}.  Similarly, when $\Delta_{t-1} = 0$, we
-define $w_t = \argmin_{w \in K} \langle L_{t-1}, w \rangle$ where ties among
-minimizers are broken by taking the one with the smallest value of $R(w)$,
-which is unique due to strong convexity; this is the same as $w_t = \lim_{a \to
-0^+} \argmin_{w \in K} (\langle L_{t-1}, w \rangle + aR(w))$.
+If $\Delta_{i-1} = 0$, we define $\Delta_{i-1}
+\Breg_{R^*}(\frac{-L_i}{\Delta_{i-1}}, \frac{-L_{i-1}}{\Delta_{i-1}})$ as
+$\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which
+always exists and is finite; see Appendix~\ref{section:limits}.  Similarly,
+when $\Delta_{t-1} = 0$, we define $w_t = \argmin_{w \in K} \langle L_{t-1}, w
+\rangle$ where ties among minimizers are broken by taking the one with the
+smallest value of $R(w)$, which is unique due to strong convexity; this is the
+same as $w_t = \lim_{a \to 0^+} \argmin_{w \in K} (\langle L_{t-1}, w \rangle +
+aR(w))$.
 
@@
 $f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
-optimization is the following corollary.  Its proof can be found in
-Appendix~\ref{section:ada-ftrl-proof}.
+optimization is the following corollary.  Its proof can be found in~\ref{section:ada-ftrl-proofs}.
 
@@
 %
-Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to
-Appendix~\ref{section:ada-ftrl-proof}.
+Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to~\ref{section:ada-ftrl-proofs}.
 Theorem~\ref{theorem:ada-ftrl-regret-bound} follows from

### 2015-11-23 mirror-descent.tex
Added proofs for mirror descent section.

@@
 If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, to define
-Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
-\rangle$, the algorithm has to choose together with $R_t$ a subgradient map
-$\grad R_t:K \to V$.  If $R_t$ is a restriction of a differentiable function
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, Bregman
+divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
+\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
+R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
+$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
 $R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.
+\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
+can be found in~\ref{section:mirror-descent-proofs}.
 
 \begin{lemma}[Regret of Mirror Descent]
+\label{lemma:mirror-descent-regret}
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
@@
 \qquad \text{and} \qquad
-R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2} \; .
+R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
 $$
@@
 All three sequences of regularizers give a scale-free algorithm. Regularizer
-\eqref{equation:ada-grad} can be viewed as a scale-free generalization of
-Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
+\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free generalization of
+Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2 = 1$ for all $i$,
 \textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
@@
 of prediction with expert advice is an instance of \textsc{Mirror Descent} for
-the probability simplex $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0
-\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{i=1}^d w_i
-\ln w_i$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
+the probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1
+\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j
+\ln w_j$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
 scale-free.
 
-Replacing $\frac{1}{2}\|w\|_2^2$ in equation $\label{equation:ada-grad}$ with
+Replacing $\frac{1}{2}\|w\|_2^2$ in equation \eqref{equation:ada-grad-regularizer} with
 an arbitrary convex function $R:K \to \R$ yields
@@
 \textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
+Proof can be found in~\ref{section:mirror-descent-proofs}.
 
@@
 
-The second counter-example is as Theorem~\ref{theorem:second-counter-example}
+The second counter-example is stated as Theorem~\ref{theorem:second-counter-example}
 below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
-\in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}$ and the regularizer is the
-negative entropy $R(w) = \sum_{i=1}^d w_i \ln w_i$.  Negative entropy is
+\in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ and the regularizer is the
+negative entropy $R(w) = \sum_{j=1}^d w_j \ln w_j$.  Negative entropy is
 $1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
 $\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
-divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{i=1}^d u_i
-\ln(u_i/v_i)$.  Note that despite that negative entropy is upper- and
+divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j
+\ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
 lower-bounded, Kullback-Leibler divergence is not.
@@
 \label{theorem:second-counter-example}
-Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}
+Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}
 \subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
@@
 such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
-1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{i=1}^d
-w_i \ln w_i$ satisfies
+1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{j=1}^d
+w_j \ln w_j$ satisfies
 $$

### 2015-11-23 per-coordinate.tex
Added proofs for mirror descent section.

@@
 \times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
-factor $K_i$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
-denote by $\Regret_T^{(i)}(u_i)$ its regret with respect to $u_i \in K_i$. The
+factor $K_j$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
+denote by $\Regret_T^{(j)}(u_j)$ its regret with respect to $u_j \in K_j$. The
 overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
@@
 $$
-\Regret_T(u) = \sum_{i=1}^d \Regret_T^{(i)}(u_i) \; .
+\Regret_T(u) = \sum_{j=1}^d \Regret_T^{(j)}(u_j) \; .
 $$
@@
 clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
-each factor $K_i$, we generalize and improve existing regret bounds
+each factor $K_j$, we generalize and improve existing regret bounds
 \cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate

### 2015-11-23 preliminaries-proofs.tex
Added proofs for mirror descent section.

@@
 \begin{proof}[Proof of Property 6 of Proposition~\ref{proposition:conjugate-properties}]
-To bound $\Breg_{f^*}(x,y)$ we add a non-negative divergence term $\Breg_{f^*}(y,x)$.
+To bound $\Breg_{f^*}(x,y)$ we add a non-negative divergence term
+$\Breg_{f^*}(y,x)$.
 \begin{align*}
@@
 \end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:generic-regret-bound}]
+By Fenchel-Young inequality,
+\begin{align*}
+\sum_{t=1}^T \left( R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) \right)
+& = R_{T+1}^*(-L_T) - R_1^*(0) \\
+& \ge - \langle L_T, u \rangle - R_{T+1}(u)  - R_1^*(0) \\
+& = - R_{T+1}(u)  - R_1^*(0) - \sum_{t=1}^T \langle \ell_t, u \rangle \; .
+\end{align*}
+We add $\sum_{t=1}^T \langle \ell_t, w_t \rangle$ to both sides and we obtain
+$\Regret_T(u)$ on the right side. After rearrangement of the terms, we get an upper bound
+on the regret:
+\begin{align*}
+\Regret_T(u)
+& = \sum_{t=1}^T \langle \ell_t, w_t \rangle - \sum_{t=1}^T \langle \ell_t, u \rangle \\
+& \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^T \left( R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, w_t \rangle \right) \; .
+\end{align*}
+By Proposition~\ref{proposition:conjugate-properties}, part 2, we have $w_t =
+\grad R_t^*(-L_{t-1})$ and therefore we can rewrite the sum in last expression as
+\begin{align*}
+& \sum_{t=1}^T R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, w_t \rangle \\
+& = \sum_{t=1}^T R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, \grad R_t^*(-L_{t-1}) \rangle \\
+& = \sum_{t=1}^T R_t^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, \grad R_t^*(-L_{t-1}) \rangle - R_t^*(-L_t) + R_{t+1}^*(-L_t) \\
+& = \sum_{t=1}^T \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
+\end{align*}
+This finishes the proof.
+\end{proof}

### 2015-11-23 preliminaries.tex
Added proofs for mirror descent section.

@@
 Except for properties 6 and 7, the proofs can be found
-in~\cite{Shalev-Shwartz-2007}.  Property 6 is proven in
-Appendix~\ref{section:definitions-proofs}. Property 7 trivially follows from
-property 2.
+in~\cite{Shalev-Shwartz-2007}.  Property 6 is proven
+in~\ref{section:definitions-proofs}. Property 7 trivially follows from property
+2.
 
@@
 
-\begin{lemma}[Lemma 1 in \cite{Orabona-Crammer-Cesa-Bianchi-2014}]
+\begin{lemma}[Regret of Follow The Regularized Leader]
 \label{lemma:generic-regret-bound}
@@
 
-\begin{proof}
-
-\end{proof}
+Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}. For completeness,
+we include it in~\ref{section:definitions-proofs}.

### 2015-11-23 solo-ftrl-proofs.tex
Added proofs for mirror descent section.

@@
 \section{Proofs for SOLO FTRL}
-\label{section:solo-ftrl-proof}
+\label{section:solo-ftrl-proofs}
+
+\begin{proof}[Proof of Lemma~\ref{lemma:sum-of-square-roots-inverses}]
+We use the inequality $x/\sqrt{x+y} \le 2(\sqrt{x+y} - \sqrt{y})$ which holds
+for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{s=1}^{t-1} a_s$,
+we get that for any $t \ge 1$,
+$$
+\frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{s=1}^t a_s} \ - \ 2 \sqrt{\sum_{s=1}^{t-1} a_s} \; .
+$$
+Summing the above inequality over all $t=1,2,\dots,T$, the right side telescopes to
+$2 \sqrt{\sum_{t=1}^T a_t}$.
+\end{proof}
 

### 2015-11-23 solo-ftrl.tex
Few improvements
Added proofs for mirror descent section.

@@
 the loss vectors. Identical considerations hold for the FTRL-Proximal in
-\cite{McMahan-Streeter-2010,McMahan-2014}: the scale-free setting of the learning rate is
-valid only in the bounded case.
+\cite{McMahan-Streeter-2010,McMahan-2014}: the scale-free setting of the
+learning rate is valid only in the bounded case.
 
@@
 Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is in
-Appendix~\ref{section:solo-ftrl-proof}.  It is similar to
+Appendix~\ref{section:solo-ftrl-proofs}.  It is similar to
 Corollary~\ref{corollary:ada-ftrl-regret-bound} for AdaFTRL. The scaling
@@
 R(w) = \frac{f(w)\sqrt{2.75}}{\sqrt{\sup_{v \in K} f(v)}}
-\quad \text{satisfies} \quad %
+\quad \text{satisfies} \quad
 \Regret_T \le 13.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
@@
 $$
-\sum_{t=1}^T \min \left\{ a_t^2 / \sqrt{\sum_{s=1}^{t-1} a_s^2}, \ C a_t \right\}
+\sum_{t=1}^T \min \left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, \ C a_t \right\}
 \le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
@@
 
-\begin{lemma}[Lemma~3.5 in \cite{Auer-Cesa-Bianchi-Gentile-2002}]
+\begin{lemma}
 \label{lemma:sum-of-square-roots-inverses}
@@
 $$
-\sum_{t=1}^T a_t / \sqrt{\sum_{s=1}^t a_s} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
+\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
 $$
 \end{lemma}
+%
+Lemma~\ref{lemma:sum-of-square-roots-inverses} is from \cite[Lemma
+3.5]{Auer-Cesa-Bianchi-Gentile-2002}.  We give
+a proof in~\ref{section:solo-ftrl-proofs} for completeness.
 

### 2015-11-24 mirror-descent.tex
small typo

@@
 = \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free Mirror Descent} with
-regularizer $R(w) = \frac{1}{2}\|w\|_w^2$, Zinkevich's \textsc{GIGA} algorithm,
+regularizer $R(w) = \frac{1}{2}\|w\|_2^2$, Zinkevich's \textsc{GIGA} algorithm,
 and both versions of \textsc{AdaGrad with Composite Mirror Descent Update}

### 2015-11-25 ada-ftrl.tex
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.

@@
 $\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which
-always exists and is finite; see Appendix~\ref{section:limits}.  Similarly,
+always exists and is finite; see~\ref{section:limits}.  Similarly,
 when $\Delta_{t-1} = 0$, we define $w_t = \argmin_{w \in K} \langle L_{t-1}, w

### 2015-11-25 mirror-descent.tex
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.

@@
 \quad \text{satisfies} \quad %
-\Regret_T(u) \le 2 \sqrt{\sup_{u,v \in K} \Breg_f(u,v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
+\Regret_T \le 2 \sqrt{\sup_{u,v \in K} \Breg_f(u,v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
 $$
 \end{corollary}
+
+The regret bound for \textsc{Scale-Free Mirror Descent}
+in the Corollary~\ref{corollary:regret-scale-free-mirror-descent}
+depends on $\sup_{u,v \in K} \Breg_f(u,v)$. In contrast,
+the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}
+in Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
+depend on $\sup_{u \in K} f(u)$.
+Similarly, regret bound in Theorem~\ref{theorem:regret-scale-free-mirror-descent} for
+\textsc{Scale-Free Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and
+regret bounds in Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
+for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to show that
+$$
+\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
+$$
+provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
+$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
+\sup_{v \in K} \Breg_R(u,v)$.
+
+The assumption $R(v^*) = 0$ and $\grad R(v^*) = 0$
+are easy to achieve by adding an affine function to the regularizer:
+$$
+R'(u) = R(u) - \langle \grad R(v^*), u - v^* \rangle - R(v^*) \; .
+$$
+The regularizer $R'$ has the same parameter of strong convexity as $R$, the
+associated Bregman divergences $D_{R'}$ and $D_{R}$ are equal, $R$ and $R'$
+have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are both zero.
+
+Thus, ignoring constant factors, regret bound for \textsc{Scale-Free Mirror
+Descent} is worse than the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO
+FTRL}.  In fact, there are examples where $R(u)$ is finite whereas $\sup_{v \in
+K} \Breg_R(u,v)$ is infinite. We mention two examples. The first example is $R(w) =
+\frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u \in V$,
+$R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V}
+\frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative entropy
+regularizer $R(w) = \ln d  + \sum_{j=1}^d w_j \ln w_j$ defined on the
+$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
+\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
+lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K}
+\sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples in the
+following subsection.
 
@@
 $\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
-$\Breg_R(u,v)$ is bounded is necessary. We show necessity of this assumption on
+$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free Mirror Descent}
+to have a ``small'' regret. We show necessity of this assumption on
 two counter-examples.  In these
@@
 $\ell_1, \ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots
-= \|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$ or
-worse regret.
+= \|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has regret
+$\Omega(T)$ or worse.
 

### 2015-12-07 introduction.tex
Added proofs for mirror descent section.
Add references to introduction

@@
 OLO is a fundamental problem in machine
-learning~\cite{Cesa-Bianchi-Lugosi-2006,Shalev-Shwartz-2011}.  Many learning
-problems can be directly phrased as OLO, e.g., learning with expert
-advice~\cite{Littlestone-Warmuth-1994,Vovk-1998,Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997}, online combinatorial
-optimization~\cite{Koolen-Warmuth-Kivinen-2010}. Other
-problems can be reduced to OLO, e.g. online convex
-optimization~\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification and
-regression~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
-problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006}, and batch and stochastic
-optimization of convex functions~\cite{Nemirovski-Yudin-1983}.  Hence, a result in OLO
-immediately implies other results in all these domains.
+learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
+Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
+e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
+Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
+online combinatorial optimization~\cite{Kalai-Vempala-2005,
+Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
+reduced to OLO, e.g. online convex
+optimization~\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
+~\cite{Rosenblatt-1958,Freund-Schapire-1999} and
+regression~\cite{Kivinen-Warmuth-1997},
+~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
+problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
+\cite{Bubeck-Cesa-Bianchi-2012}, and batch and stochastic optimization of
+convex functions~\cite{Nemirovski-Yudin-1983, Bubeck-2015}.  Hence, a result in
+OLO immediately implies other results in all these domains.
 
@@
 % Change spacing between rows.
-\renewcommand{\arraystretch}{1.5}
+\renewcommand{\arraystretch}{1.8}
 
@@
 \textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
-\textsc{Hedge} \cite{Freund-Schapire-1999} & Probability Simplex & Negative Entropy & No \\ \hline
+\textsc{Hedge} \cite{Freund-Schapire-1997} & Probability Simplex & Negative Entropy & No \\ \hline
 \textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline

### 2015-12-07 mirror-descent.tex
Add references to Mirror Descent section

@@
 \in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
-can be found in~\ref{section:mirror-descent-proofs}.
+can be found in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.
+For completeness we give a proof in~\ref{section:mirror-descent-proofs}.
 
@@
 \textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
-Proof can be found in~\ref{section:mirror-descent-proofs}.
+Proof is a generalization of the proof
+of~\cite{Duchi-Shalev-Shwartz-Singer-Tewari-2010,Duchi-Hazan-Singer-2011} and
+it can be found can be found in~\ref{section:mirror-descent-proofs}.
 

### 2015-12-07 mirror-descent.tex
Fixed typos in mirror descent section.

@@
 for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to show that
-$$
+\begin{equation}
+\label{equation:regularizer-vs-divergence}
 \forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
-$$
+\end{equation}
 provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
@@
 The regularizer $R'$ has the same parameter of strong convexity as $R$, the
-associated Bregman divergences $D_{R'}$ and $D_{R}$ are equal, $R$ and $R'$
-have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are both zero.
+associated Bregman divergences $\Breg_{R'}$ and $\Breg_{R}$ are equal, $R'$ and
+$R$ have the same minimizer $v^*$, and $\grad R'(v^*)$ and $\grad R'(v^*)$ are
+both zero.
 
-Thus, ignoring constant factors, regret bound for \textsc{Scale-Free Mirror
-Descent} is worse than the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO
-FTRL}.  In fact, there are examples where $R(u)$ is finite whereas $\sup_{v \in
-K} \Breg_R(u,v)$ is infinite. We mention two examples. The first example is $R(w) =
-\frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u \in V$,
-$R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V}
-\frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative entropy
-regularizer $R(w) = \ln d  + \sum_{j=1}^d w_j \ln w_j$ defined on the
-$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
-\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
-lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K}
-\sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples in the
-following subsection.
+Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
+factors, regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
+better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
+\textsc{SOLO FTRL}.  In fact, it is not hard to come up with examples where
+$R(u)$ is finite whereas $\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention
+two such examples. The first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined
+on the whole space $V$, where for any $u \in V$, $R(u)$ is a finite value but
+$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 =
++\infty$. The second example is the negative entropy regularizer $R(w) = \ln d
++ \sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex
+$K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u
+\in K$, $R(u)$ is finite and in fact lies in the interval $[-\ln d, 0]$ but
+$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
+\infty$. We return to these examples in the following subsection.
 
@@
 \ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
-lower-bounded, Kullback-Leibler divergence is not.
+lower-bounded, Kullback-Leibler divergence is unbounded from above.
 
@@
 \label{theorem:second-counter-example}
-Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}
-\subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
+Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$
+be the $d$-dimensional probability simplex.  For any $T \ge 120$
 there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$

### 2015-12-07 ada-ftrl.tex
Use small-caps for all algorithms names

@@
-\section{AdaFTRL}
+\section{\textsc{AdaFTRL}}
 \label{section:ada-ftrl}
 
-In this section we generalize the AdaHedge
+In this section we generalize the \textsc{AdaHedge}
 algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
@@
 and based on general properties of strongly convex functions, rather than
-specific properties of the entropic regularizer like in AdaHedge.
+specific properties of the entropic regularizer like in \textsc{AdaHedge}.
 
@@
 semi-continuous function that is $\lambda$-strongly convex with respect to
-$\|\cdot\|$ and is bounded from above.  The regret of AdaFTRL satisfies
+$\|\cdot\|$ and is bounded from above.  The regret of \textsc{AdaFTRL} satisfies
 $$
@@
 $1$-strongly convex with respect to $\|\cdot\|$ and is bounded from above.  The
-regret of AdaFTRL with regularizer
+regret of \textsc{AdaFTRL} with regularizer
 $$
@@
 
-\subsection{Proof of Regret Bound for AdaFTRL}
+\subsection{Proof of Regret Bound for \textsc{AdaFTRL}}
 \label{section:ada-ftrl-regret-bound}
@@
 \label{lemma:initial-regret-bound}
-AdaFTRL, for any $u \in K$ and any $u \ge 0$, satisfies
+\textsc{AdaFTRL}, for any $u \in K$ and any $u \ge 0$, satisfies
 $\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T$.
@@
 Let $D = \sup_{u, v \in K} \|u -v\|$ be the diameter of $K$.
-The sequence $\{\Delta_t\}_{t=1}^\infty$ generated by AdaFTRL satisfies for any $t \ge 1$,
+The sequence $\{\Delta_t\}_{t=1}^\infty$ generated by \textsc{AdaFTRL} satisfies for any $t \ge 1$,
 $$
@@
 Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
-generated by AdaFTRL satisfies for any $T \ge 0$,
+generated by \textsc{AdaFTRL} satisfies for any $T \ge 0$,
 $$

### 2015-12-07 introduction.tex
Add references to Mirror Descent section
Use small-caps for all algorithms names

@@
 parameter, or the parameter of strong convexity of the regularizer. Examples of
-such algorithms include the Hedge algorithm or online projected gradient
+such algorithms include the \textsc{Hedge} algorithm or online projected gradient
 descent with fixed learning rate.  However, these algorithms have two obvious
@@
 \textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func.
-\newline $1<p\leq2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
+\newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
 \textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
@@
 
-\footnotetext{\label{footnote-label} These algorithms attempt to produce an invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a sequence of invariant $w_t$.}
+\footnotetext{\label{footnote-label} These algorithms attempt to produce an
+invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a
+sequence of invariant $w_t$.}
 
@@
 general OLO setting this issue has been largely ignored.  In particular, the
-AdaHedge~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm, for
+\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm, for
 prediction with expert advice, is specifically designed to be scale-free.  A
 notable exception in the OLO literature is the discussion of the ``off-by-one''
-issue in~\cite{McMahan-2014}, where it is explained that even the popular AdaGrad
+issue in~\cite{McMahan-2014}, where it is explained that even the popular \textsc{AdaGrad}
 algorithm~\cite{Duchi-Hazan-Singer-2011} is not completely adaptive; see also
@@
 Section~\ref{section:ada-ftrl} we show that the analysis and design of
-AdaHedge can be generalized to the OLO scenario and to any strongly convex
+\textsc{AdaHedge} can be generalized to the OLO scenario and to any strongly convex
 regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new and
@@
 \textbf{first} scale-free online algorithm for unbounded sets with a
-non-vacuous regret bound.  Both algorithms are instances of Follow The
-Regularized Leader (FTRL) with an adaptive learning rate.  Moreover, our
+non-vacuous regret bound.  Both algorithms are instances of \textsc{Follow The
+Regularized Leader} (\textsc{FTRL}) with an adaptive learning rate.  Moreover, our
 algorithms show that scale-free algorithms can be obtained in a ``native'' and

### 2015-12-07 preliminaries.tex
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.
Use small-caps for all algorithms names

@@
 The following proposition relates the range of values of a strongly convex
-function to the diameter of its domain. The proof can be found in
-Appendix~\ref{section:definitions-proofs}.
+function to the diameter of its domain. The proof can be found
+in~\ref{section:definitions-proofs}.
 
@@
 Our scale-free online learning algorithms are versions of the \textsc{Follow
-The Regularized Leader} (FTRL) algorithm with varying regularizers, presented
-as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The following lemma
-bounds its regret.
+The Regularized Leader} (\textsc{FTRL}) algorithm with varying regularizers,
+presented as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The following
+lemma bounds its regret.
 

### 2015-12-07 solo-ftrl.tex
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.
Use small-caps for all algorithms names

@@
-\section{SOLO FTRL}
+\section{\textsc{SOLO FTRL}}
 \label{section:solo-ftrl}
 
-The closest algorithm to a scale-free one in the OLO literature is the AdaGrad
+The closest algorithm to a scale-free one in the OLO literature is the \textsc{AdaGrad}
 algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a regularizer on each
@@
 knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
-other words, despite of its name, AdaGrad is not fully adaptive to the norm of
-the loss vectors. Identical considerations hold for the FTRL-Proximal in
+other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the norm of
+the loss vectors. Identical considerations hold for the \textsc{FTRL-Proximal} in
 \cite{McMahan-Streeter-2010,McMahan-2014}: the scale-free setting of the
@@
 the additional parameter of the doubling power.}, such strategy would give the
-feeling that FTRL needs to be ``fixed'' in order to obtain a scale-free
+feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a scale-free
 algorithm.
@@
 semi-continuous function that is $\lambda$-strongly convex with respect to
-$\|\cdot\|$. The regret of SOLO FTRL satisfies
+$\|\cdot\|$. The regret of \textsc{SOLO FTRL} satisfies
 \begin{align*}
@@
 optimize $\lambda$.  The result of the optimization is
-Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is in
-Appendix~\ref{section:solo-ftrl-proofs}.  It is similar to
-Corollary~\ref{corollary:ada-ftrl-regret-bound} for AdaFTRL. The scaling
+Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is
+in~\ref{section:solo-ftrl-proofs}.  It is similar to
+Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}. The scaling
 however is different in the two corollaries.  In
@@
 that $f:K \to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
-convex with respect to $\|\cdot\|$. SOLO FTRL with regularizer
+convex with respect to $\|\cdot\|$. \textsc{SOLO FTRL} with regularizer
 $$
@@
 
-\subsection{Proof of Regret Bound for SOLO FTRL}
+\subsection{Proof of Regret Bound for \textsc{SOLO FTRL}}
 \label{section:solo-ftrl-regret-bound}

### 2015-12-07 mirror-descent.tex
Replace subscript "s" with "i"

@@
 associated Bregman divergences $\Breg_{R'}$ and $\Breg_{R}$ are equal, $R'$ and
-$R$ have the same minimizer $v^*$, and $\grad R'(v^*)$ and $\grad R'(v^*)$ are
+$R$ have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are
 both zero.

### 2015-12-07 solo-ftrl.tex
Replace subscript "s" with "i"

@@
 \begin{equation*}
-R_t(w) = R(w) \left(\delta + \sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2} \right).
+R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
 \end{equation*}
@@
 \begin{equation*}
-R_t(w) = R(w) \sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2} \; ,
+R_t(w) = R(w) \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \; ,
 \end{equation*}
@@
 $$
-\sum_{t=1}^T \min \left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, \ C a_t \right\}
+\sum_{t=1}^T \min \left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, \ C a_t \right\}
 \le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
@@
 $$
-\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
-\le 2 \sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) + \frac{C\alpha( M_t  - M_{t-1})}{\alpha - 1}
+\min\left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, C a_t \right\}
+\le 2 \sqrt{1+\alpha^2} \left( \sqrt{\sum_{i=1}^t a_i^2} - \sqrt{\sum_{i=1}^{t-1} a_i^2} \right) + \frac{C\alpha( M_t  - M_{t-1})}{\alpha - 1}
 $$
 from which the inequality follows by summing over $t=1,2,\dots,T$ and choosing $\alpha = \sqrt{2}$.
-The inequality follows by case analysis. If $a_t^2 \le \alpha^2 \sum_{s=1}^{t-1} a_s^2$, we have
+The inequality follows by case analysis. If $a_t^2 \le \alpha^2 \sum_{i=1}^{t-1} a_i^2$, we have
 \begin{multline*}
-\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
-\le \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}
-= \frac{a_t^2}{\sqrt{\frac{1}{1+\alpha^2} \left( \alpha^2 \sum_{s=1}^{t-1} a_s^2 + \sum_{s=1}^{t-1} a_s^2 \right)}} \\
-\le \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{ a_t^2 + \sum_{s=1}^{t-1} a_s^2 }}
-= \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{\sum_{s=1}^t a_s^2}}
-\le 2\sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
+\min\left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, C a_t \right\}
+\le \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}
+= \frac{a_t^2}{\sqrt{\frac{1}{1+\alpha^2} \left( \alpha^2 \sum_{i=1}^{t-1} a_i^2 + \sum_{i=1}^{t-1} a_i^2 \right)}} \\
+\le \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{ a_t^2 + \sum_{i=1}^{t-1} a_i^2 }}
+= \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{\sum_{i=1}^t a_i^2}}
+\le 2\sqrt{1+\alpha^2} \left( \sqrt{\sum_{i=1}^t a_i^2} - \sqrt{\sum_{i=1}^{t-1} a_i^2} \right)
 \end{multline*}
 where we have used $x^2/\sqrt{x^2+y^2} \le 2(\sqrt{x^2+y^2} - \sqrt{y^2})$ in the last step.
-On the other hand, if $a_t^2 > \alpha^2 \sum_{t=1}^{t-1} a_s^2$, we have
+On the other hand, if $a_t^2 > \alpha^2 \sum_{t=1}^{t-1} a_i^2$, we have
 \begin{multline*}
-\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
+\min\left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, C a_t \right\}
 \le C a_t
 = C \frac{\alpha a_t  - a_t}{\alpha - 1}
-\le \frac{C}{\alpha - 1} \left( \alpha a_t  - \alpha \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) \\
-= \frac{C\alpha}{\alpha - 1} \left( a_t  - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
+\le \frac{C}{\alpha - 1} \left( \alpha a_t  - \alpha \sqrt{\sum_{i=1}^{t-1} a_i^2} \right) \\
+= \frac{C\alpha}{\alpha - 1} \left( a_t  - \sqrt{\sum_{i=1}^{t-1} a_i^2} \right)
 \le \frac{C\alpha}{\alpha - 1} \left( a_t  - M_{t-1} \right)
@@
 \end{multline*}
-where we have used that $a_t = M_t$ and $\sqrt{\sum_{s=1}^{t-1} a_s^2} \ge M_{t-1}$.
+where we have used that $a_t = M_t$ and $\sqrt{\sum_{i=1}^{t-1} a_i^2} \ge M_{t-1}$.
 \end{proof}
@@
 $$
-\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
+\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{i=1}^t a_i}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
 $$
@@
 \begin{proof}[Proof of Theorem~\ref{theorem:regret-solo-ftrl}]
-Let $\eta_t=\tfrac{1}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}$, hence $R_t(w) = \tfrac{1}{\eta_t} R(w)$.
+Let $\eta_t=\tfrac{1}{\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}$, hence $R_t(w) = \tfrac{1}{\eta_t} R(w)$.
 We assume without loss of
@@
 & \qquad + \langle \nabla R_t^*(-L_{t-1})-\nabla R_{t+1}^*(-L_{t-1}), \ell_t \rangle  \\
-& \le \tfrac{1}{2\lambda} \eta_{t+1} \|\ell_t\|_*^2 + \| \nabla R_t^*(-L_{t-1})-\nabla R_{t+1}^*(-L_{t-1})\| \cdot \|\ell_t\|_* \\
-& = \tfrac{1}{2\lambda} \eta_{t+1} \|\ell_t\|_*^2 + \| \nabla R^*(- \eta_{t} L_{t-1})-\nabla R^*(- \eta_{t+1} L_{t-1})\| \cdot \|\ell_t\|_* \\
+& \le \frac{\eta_{t+1} \|\ell_t\|_*^2}{2\lambda} + \| \nabla R_t^*(-L_{t-1})-\nabla R_{t+1}^*(-L_{t-1})\| \cdot \|\ell_t\|_* \\
+& = \frac{\eta_{t+1} \|\ell_t\|_*^2}{2\lambda} + \| \nabla R^*(- \eta_{t} L_{t-1})-\nabla R^*(- \eta_{t+1} L_{t-1})\| \cdot \|\ell_t\|_* \\
 & \le \frac{\eta_{t+1} \|\ell_t\|_*^2}{2\lambda} + \min\left\{\frac{1}{\lambda} \|L_{t-1}\|_* \left(\eta_{t} - \eta_{t+1} \right), D\right\} \|\ell_t\|_* \; ,
@@
 & \le \frac{1}{\eta_{T+1}} R(u) + \frac{1}{2\lambda} \sum_{t=1}^T  \eta_{t+1} \|\ell_t\|_*^2 + \frac{1}{2\lambda} \sum_{t=1}^T \min\left\{ \eta_t \|\ell_t\|_*^2, \ 2 \lambda H \|\ell_t\|_* \right\} \\
-& = \frac{1}{\eta_{T+1}} R(u) + \frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^t \|\ell_t\|_*^2}} + \frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}, \ 2 \lambda H \|\ell_t\|_* \right\} \; .
+& = \frac{1}{\eta_{T+1}} R(u) + \frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}} + \frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}, \ 2 \lambda H \|\ell_t\|_* \right\} \; .
 \end{align*}
@@
 $$
-\frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^t \|\ell_t\|_*^2}}
+\frac{1}{2\lambda} \sum_{t=1}^T  \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}}
 \le \frac{1}{\lambda} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
@@
 $$
-\frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{s=1}^{t-1} \|\ell_s\|_*^2}}, \ 2 \lambda \|\ell_t\|_* H \right\}
+\frac{1}{2 \lambda} \sum_{t=1}^T \min\left\{ \frac{\|\ell_t\|_*^2}{\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}, \ 2 \lambda \|\ell_t\|_* H \right\}
 \le 3.5 H \max_{t \le T} \|\ell_t\|_* + \frac{1.75}{\lambda} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .

### 2015-12-07 ada-ftrl.tex
Replace subscript "s" with "i"
Add limit explanations and limit lemma

@@
 The sequence $\{\Delta_t\}_{t=0}^\infty$ is non-negative and non-decreasing.
-Also, $\Delta_t$ as a function of $\{\ell_s\}_{s=1}^t$ is positive homogenous
+Also, $\Delta_t$ as a function of $\ell_1, \ell_2, \dots, \ell_t$ is positive homogenous
 of degree one, making the algorithm scale-free.
@@
 $\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which
-always exists and is finite; see~\ref{section:limits}.  Similarly,
-when $\Delta_{t-1} = 0$, we define $w_t = \argmin_{w \in K} \langle L_{t-1}, w
-\rangle$ where ties among minimizers are broken by taking the one with the
-smallest value of $R(w)$, which is unique due to strong convexity; this is the
-same as $w_t = \lim_{a \to 0^+} \argmin_{w \in K} (\langle L_{t-1}, w \rangle +
-aR(w))$.
+always exists and is finite; see Lemma~\ref{lemma:limit-bregman-divergence}
+in~\ref{section:limits}.  Similarly, when $\Delta_{t-1} = 0$, we define $w_t =
+\argmin_{w \in K} \langle L_{t-1}, w \rangle$ where ties among minimizers are
+broken by taking the one with the smallest value of $R(w)$, which is unique due
+to strong convexity; as we show in Lemma~\ref{lemma:prediction-limit-existence}
+in~\ref{section:limits}, this is the same as $w_t = \lim_{a \to 0^+}
+\argmin_{w \in K} (\langle L_{t-1}, w \rangle + aR(w))$.
 
@@
 \end{equation}
+Technically, \eqref{equation:regret-bound-inequality} is not justified since $R_t$
+might not be strongly convex. This happens when $\Delta_{t-1} = 0$. In order to justify
+\eqref{equation:regret-bound-inequality}, we consider a different algorithm
+that initializes $\Delta_0 = \epsilon$ where $\epsilon > 0$; that ensures that $\Delta_{t-1} > 0$ and $R_t$ is strongly convex.
+Applying  Lemma~\ref{lemma:generic-regret-bound} and then taking limit $\epsilon \to 0$,
+yields \eqref{equation:regret-bound-inequality}.
+
 Since, $\Breg_{R_t^*}(u,v) = \Delta_{t-1} \Breg_{R^*}(\frac{u}{\Delta_{t-1}},

### 2015-12-07 limits.tex
Add limit explanations and limit lemma

@@
 
-\begin{lemma}
-\label{lemma:limit-existence}
+In this section we show that prediction of \textsc{AdaFTRL} is correctly
+defined when the regularizer is multiplied by zero.
+
+\begin{lemma}[Prediction for Zero Regularizer]
+\label{lemma:prediction-limit-existence}
+Let $K$ be non-empty bounded closed convex subset of a finite dimensional normed real vector space $(V, \|\cdot\|)$.
+Let $R:K \to \R$ be strictly convex and lower semi-continuous, and let $L \in V^*$. The limit
+\begin{equation}
+\label{equation:limit}
+\lim_{\eta \to +\infty}
+\argmin_{w \in K} \left( \langle L, w \rangle + \frac{1}{\eta} R(w) \right)
+\end{equation}
+exists and it is equal to the unique minimizer of $R(w)$ over the set (of minimizers)
+$$
+\left\{ w \in K ~:~ \langle L, w \rangle = \inf_{v \in K} \langle L, v \rangle \right\} \; .
+$$
+\end{lemma}
+
+Before we give the proof, we illustrate the lemma on a simple
+example.  Let $K = [-1,1]^2$ be the unit square in $\R^2$ and let
+$R(w) = \|w\|_2^2$. Let $L = (1,0)$. The minimizers are
+$$
+\argmin_{w \in K} \langle L, w \rangle = \{ (-1,y) ~:~ y \in [-1,1] \} \; .
+$$
+The minimizer with smallest value of $R(w)$ is $(-1,0)$. Hence the lemma implies
+that
+$$
+\lim_{\eta \to +\infty} \argmin_{w \in K}
+\left( \langle L, w \rangle + \frac{1}{\eta} \|w\|_2^2 \right) = (-1, 0) \; .
+$$
+
+\begin{proof}
+Without loss of generality, we can assume that $R(w)$ is non-negative for any $w \in K$.
+For otherwise, we can replace $R(w)$ with $R'(w) = R(w) - \inf_{v \in K} R(v)$.
+
+Since $K$ is a non-empty bounded closed convex subset of a finite dimensional normed vector space, it is compact
+and $r^* = \min_{w \in K} \langle L, w \rangle$ exists and is attained at some $w \in K$. Consider the hyperplane
+$$
+H = \{ w \in V ~:~ \langle L, w \rangle = r^* \} \; .
+$$
+The intersection $H \cap K$ is a non-empty compact convex set.
+Let
+$$
+v^* = \argmin_{v \in K \cap H} R(v) \; .
+$$
+The existence of $v^*$ follows from compactness of $H \cap K$ and lower
+semi-continuity of $R(v)$.  Uniqueness of $v^*$ follows from
+strict convexity of $R(v)$. We show that the limit (\ref{equation:limit})
+equals $v^*$.
+
+By the definition of $H$,
+\begin{equation}
+\label{equation:limit-is-minimizer}
+v^* \in \argmin_{w \in K} \ \langle L, w \rangle \; .
+\end{equation}
+Let $S = \{ w \in K ~:~ R(w) \le R(v^*) \}$. Since $R(w)$ is lower semi-continuous $S$
+is closed. Since $R(w)$ is strictly convex, $S \cap H = \{v^*\}$.
+
+For any $\eta > 0$, let
+$$
+w(\eta) = \argmin_{w \in K} \left( \langle L, w \rangle + \frac{1}{\eta} R(w) \rangle \right) \; .
+$$
+We prove that $w(\eta) \in S$. Indeed, by optimality of $v^*$ and $w(\eta)$
+$$
+\frac{1}{\eta} R(w(\eta))  + \langle L, w(\eta) \rangle
+\le
+\frac{1}{\eta} R(v^*)  + \langle L, v^* \rangle
+\le
+\frac{1}{\eta} R(v^*)  + \langle L, w(\eta) \rangle \; .
+$$
+and hence $R(w(\eta)) \le R(v^*)$.
+
+By non-negativity of $R$ and optimality of $w(\eta)$ we have
+$$
+\langle L, w(\eta) \rangle
+\le \langle L, w(\eta) \rangle + \frac{1}{\eta} R(w(\eta))
+\le \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \; .
+$$
+Taking the limit $\eta \to +\infty$, we see that
+$$
+\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle \le
+\lim_{\eta \to +\infty} \left( \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \right)
+= \langle L, v^* \rangle \; .
+$$
+From~(\ref{equation:limit-is-minimizer}) we have
+$\langle L, v^* \rangle \le \langle L, w \rangle$ for any $w$, and therefore
+\begin{equation}
+\label{equation:limit-equality}
+\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle = \langle L, v^* \rangle \; .
+\end{equation}
+
+Consider any sequence $\{\eta_t\}_{t=1}^\infty$ of positive numbers approaching
+$+\infty$.  Since $K$ is compact, $w(\eta_t)$ has a convergent subsequence.
+Thus $\{w(\eta_t)\}_{t=1}^\infty$ has at least one accumulation point; let
+$w^*$ be any of them. We will show that $w^* = v^*$.
+
+Consider a subsequence $\{\xi_t\}_{t=1}^\infty$ of
+$\{\eta_t\}_{t=1}^\infty$ such that $\lim_{t \to \infty} w(\xi_t) = w^*$.
+Since $w(\xi_t) \in S$ and $S$ is closed, $w^* \in S$.
+From (\ref{equation:limit-equality}) we have $\langle L, w^*
+\rangle = \langle L, v^* \rangle$ and hence $w^* \in H$. Thus $w^* \in S \cap H$.
+Since $v^*$ is the only point in $S \cap H$ we must have $w^* = v^*$.
+\end{proof}
+
+\begin{lemma}[Limit of Bregman Divergence]
+\label{lemma:limit-bregman-divergence}
 Let $K$ be a non-empty bounded closed convex subset of a finite dimensional

### 2015-12-08 ada-ftrl.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 The sequence $\{\Delta_t\}_{t=0}^\infty$ is non-negative and non-decreasing.
-Also, $\Delta_t$ as a function of $\ell_1, \ell_2, \dots, \ell_t$ is positive homogenous
-of degree one, making the algorithm scale-free.
+Also, $\Delta_t$ as a function of $\ell_1, \ell_2, \dots, \ell_t$ is positive
+homogeneous of degree one, making the algorithm scale-free.
 
@@
 broken by taking the one with the smallest value of $R(w)$, which is unique due
-to strong convexity; as we show in Lemma~\ref{lemma:prediction-limit-existence}
-in~\ref{section:limits}, this is the same as $w_t = \lim_{a \to 0^+}
-\argmin_{w \in K} (\langle L_{t-1}, w \rangle + aR(w))$.
+to strong convexity. As we show in Lemma~\ref{lemma:prediction-limit-existence}
+in~\ref{section:limits}, this is the same as $w_t = \lim_{a \to 0^+} \argmin_{w
+\in K} (\langle L_{t-1}, w \rangle + aR(w))$.
 
@@
 $$
-\Regret_T(u) \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \left(1 + R(u) \right) \; .
+\Regret_T(u)
+\le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\}
+    \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \left(1 + R(u) \right) \; .
 $$
@@
 $f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
-optimization is the following corollary.  Its proof can be found in~\ref{section:ada-ftrl-proofs}.
+optimization is the following corollary.  Its proof can be found
+in~\ref{section:ada-ftrl-proofs}.
 
@@
 \end{equation}
-Technically, \eqref{equation:regret-bound-inequality} is not justified since $R_t$
-might not be strongly convex. This happens when $\Delta_{t-1} = 0$. In order to justify
-\eqref{equation:regret-bound-inequality}, we consider a different algorithm
-that initializes $\Delta_0 = \epsilon$ where $\epsilon > 0$; that ensures that $\Delta_{t-1} > 0$ and $R_t$ is strongly convex.
-Applying  Lemma~\ref{lemma:generic-regret-bound} and then taking limit $\epsilon \to 0$,
-yields \eqref{equation:regret-bound-inequality}.
+Technically, \eqref{equation:regret-bound-inequality} is not justified since
+$R_t$ might not be strongly convex. This happens when $\Delta_{t-1} = 0$. In
+order to justify \eqref{equation:regret-bound-inequality}, we consider a
+different algorithm that initializes $\Delta_0 = \epsilon$ where $\epsilon >
+0$; that ensures that $\Delta_{t-1} > 0$ and $R_t$ is strongly convex.
+Applying  Lemma~\ref{lemma:generic-regret-bound} and then taking limit
+$\epsilon \to 0$, yields \eqref{equation:regret-bound-inequality}.
 
 Since, $\Breg_{R_t^*}(u,v) = \Delta_{t-1} \Breg_{R^*}(\frac{u}{\Delta_{t-1}},
-\frac{v}{\Delta_{t-1}})$ by definition of Bregman divergence and Part 8 of
+\frac{v}{\Delta_{t-1}})$ by definition of Bregman divergence and property 8 of
 Proposition~\ref{proposition:conjugate-properties}, we have $\sum_{t=1}^T
@@
 %
-Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to~\ref{section:ada-ftrl-proofs}.
-Theorem~\ref{theorem:ada-ftrl-regret-bound} follows from
+Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred
+to~\ref{section:ada-ftrl-proofs}.  Theorem~\ref{theorem:ada-ftrl-regret-bound}
+follows from
 Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.

### 2015-12-08 conclusions.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

\section{Conclusions}
\label{section:conclusions}

We have investigated scale-free algorithms for online linear optimization and
we have shown that scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors, i.e. without any assumption on the norms of the loss vectors
and the number of rounds. We have designed two scale-free algorithms for online linear
optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
\|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly convex
function with respect to a norm $\|\cdot\|$ defined on the decision set and
where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
based on \textsc{Follow The Regularizer Leader}.  We have also designed an
algorithm with regret $O\left((R(u) + \frac{1}{\lambda}) \sqrt{T}
\max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any competitor $u
\in K$, where $R$ is any non-negative $\lambda$-strongly convex function
defined on the decision set.  The latter result makes sense even when the
decision set $K$ is unbounded.

Similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
we gave examples of regularizers for which the associated Bregman divergence is
unounded and the scale-fre version of \textsc{Mirror Descent} has $\Omega(T)$
regret or worse.

We have shown an $\Omega\left( D \sqrt{\sum_{t=1}^T \|\ell\|_*^2}\right)$
lower bound on regret of any algorithm for decision set with diameter $D$.

### 2015-12-08 introduction.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 reduced to OLO, e.g. online convex
-optimization~\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
+optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
+\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
 ~\cite{Rosenblatt-1958,Freund-Schapire-1999} and
@@
 problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
-\cite{Bubeck-Cesa-Bianchi-2012}, and batch and stochastic optimization of
-convex functions~\cite{Nemirovski-Yudin-1983, Bubeck-2015}.  Hence, a result in
-OLO immediately implies other results in all these domains.
+\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch
+and stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
+Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
+these domains.
 
@@
 parameter, or the parameter of strong convexity of the regularizer. Examples of
-such algorithms include the \textsc{Hedge} algorithm or online projected gradient
-descent with fixed learning rate.  However, these algorithms have two obvious
-drawbacks.
+such algorithms include the \textsc{Hedge} algorithm or online projected
+gradient descent with fixed learning rate.  However, these algorithms have two
+obvious drawbacks.
 
@@
 \textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
-\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \leq C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
+\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \le C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
 \textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func.
@@
 notable exception in the OLO literature is the discussion of the ``off-by-one''
-issue in~\cite{McMahan-2014}, where it is explained that even the popular \textsc{AdaGrad}
-algorithm~\cite{Duchi-Hazan-Singer-2011} is not completely adaptive; see also
-our discussion in Section~\ref{section:solo-ftrl}. In particular, existing
-scale-free algorithms cover only some norms/regularizers and \emph{only}
-bounded decision sets. The case of \textbf{unbounded decision sets},
-practically the most interesting one for machine learning applications, remains
-completely unsolved.
+issue in~\cite{McMahan-2014}, where it is explained that even the popular
+\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011} is not completely
+adaptive; see also our discussion in Section~\ref{section:solo-ftrl}. In
+particular, existing scale-free algorithms cover only some norms/regularizers
+and \emph{only} bounded decision sets. The case of \textbf{unbounded decision
+sets}, practically the most interesting one for machine learning applications,
+remains completely unsolved.
 
-Rather than trying to design strategies for a particular form of loss
-vectors and/or decision sets, in this paper we explicitly focus on the
-scale-free property. Regret of scale-free algorithms is proportional to the
-scale of the losses, ensuring optimal linear dependency on the maximum norm of
-the loss vectors.
+Rather than trying to design strategies for a particular form of loss vectors
+and/or decision sets, in this paper we explicitly focus on the scale-free
+property. Regret of scale-free algorithms is proportional to the scale of the
+losses, ensuring optimal linear dependency on the maximum norm of the loss
+vectors.
 
@@
 Section~\ref{section:ada-ftrl} we show that the analysis and design of
-\textsc{AdaHedge} can be generalized to the OLO scenario and to any strongly convex
-regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new and
-rather interesting way to adapt the learning rates to have scale-free
+\textsc{AdaHedge} can be generalized to the OLO scenario and to any strongly
+convex regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new
+and rather interesting way to adapt the learning rates to have scale-free
 algorithms.  Second, in Section~\ref{section:solo-ftrl} we propose a new and
@@
 non-vacuous regret bound.  Both algorithms are instances of \textsc{Follow The
-Regularized Leader} (\textsc{FTRL}) with an adaptive learning rate.  Moreover, our
-algorithms show that scale-free algorithms can be obtained in a ``native'' and
-simple way, i.e.  without using ``doubling tricks'' that attempt to fix poorly
-designed algorithms rather than directly solving the problem.
+Regularized Leader} (\textsc{FTRL}) with an adaptive learning rate.  Moreover,
+our algorithms show that scale-free algorithms can be obtained in a ``native''
+and simple way, i.e.  without using ``doubling tricks'' that attempt to fix
+poorly designed algorithms rather than directly solving the problem.
 

### 2015-12-08 limits.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 \label{lemma:prediction-limit-existence}
-Let $K$ be non-empty bounded closed convex subset of a finite dimensional normed real vector space $(V, \|\cdot\|)$.
-Let $R:K \to \R$ be strictly convex and lower semi-continuous, and let $L \in V^*$. The limit
+Let $K$ be non-empty bounded closed convex subset of a finite dimensional
+normed real vector space $(V, \|\cdot\|)$.  Let $R:K \to \R$ be strictly convex
+and lower semi-continuous, and let $L \in V^*$. The limit
 \begin{equation}
@@
 Before we give the proof, we illustrate the lemma on a simple
-example.  Let $K = [-1,1]^2$ be the unit square in $\R^2$ and let
+example.  Let $K = [-1,1]^2$ be a closed square in $\R^2$ and let
 $R(w) = \|w\|_2^2$. Let $L = (1,0)$. The minimizers are
@@
 \begin{proof}
-Without loss of generality, we can assume that $R(w)$ is non-negative for any $w \in K$.
-For otherwise, we can replace $R(w)$ with $R'(w) = R(w) - \inf_{v \in K} R(v)$.
+Without loss of generality, we can assume that $R(w)$ is non-negative for any
+$w \in K$.  For otherwise, we can replace $R(w)$ with $R'(w) = R(w) - \inf_{v
+\in K} R(v)$.
 
-Since $K$ is a non-empty bounded closed convex subset of a finite dimensional normed vector space, it is compact
-and $r^* = \min_{w \in K} \langle L, w \rangle$ exists and is attained at some $w \in K$. Consider the hyperplane
+Since $K$ is a non-empty bounded closed convex subset of a finite dimensional
+normed vector space, it is compact and $r^* = \min_{w \in K} \langle L, w
+\rangle$ exists and is attained at some $w \in K$. Consider the hyperplane
 $$
@@
 The existence of $v^*$ follows from compactness of $H \cap K$ and lower
-semi-continuity of $R(v)$.  Uniqueness of $v^*$ follows from
-strict convexity of $R(v)$. We show that the limit (\ref{equation:limit})
-equals $v^*$.
+semi-continuity of $R(v)$.  Uniqueness of $v^*$ follows from strict convexity
+of $R(v)$. We show that the limit (\ref{equation:limit}) equals $v^*$.
 
@@
 \end{equation}
-Let $S = \{ w \in K ~:~ R(w) \le R(v^*) \}$. Since $R(w)$ is lower semi-continuous $S$
-is closed. Since $R(w)$ is strictly convex, $S \cap H = \{v^*\}$.
+Let $S = \{ w \in K ~:~ R(w) \le R(v^*) \}$. Since $R(w)$ is lower
+semi-continuous $S$ is closed. Since $R(w)$ is strictly convex, $S \cap H =
+\{v^*\}$.
 
@@
 $$
-\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle \le
-\lim_{\eta \to +\infty} \left( \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \right)
+\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle
+\le\lim_{\eta \to +\infty} \left( \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \right)
 = \langle L, v^* \rangle \; .
@@
 
-Consider a subsequence $\{\xi_t\}_{t=1}^\infty$ of
-$\{\eta_t\}_{t=1}^\infty$ such that $\lim_{t \to \infty} w(\xi_t) = w^*$.
-Since $w(\xi_t) \in S$ and $S$ is closed, $w^* \in S$.
-From (\ref{equation:limit-equality}) we have $\langle L, w^*
-\rangle = \langle L, v^* \rangle$ and hence $w^* \in H$. Thus $w^* \in S \cap H$.
-Since $v^*$ is the only point in $S \cap H$ we must have $w^* = v^*$.
+Consider a subsequence $\{\xi_t\}_{t=1}^\infty$ of $\{\eta_t\}_{t=1}^\infty$
+such that $\lim_{t \to \infty} w(\xi_t) = w^*$.  Since $w(\xi_t) \in S$ and $S$
+is closed, $w^* \in S$.  From (\ref{equation:limit-equality}) we have $\langle
+L, w^* \rangle = \langle L, v^* \rangle$ and hence $w^* \in H$. Thus $w^* \in S
+\cap H$.  Since $v^*$ is the only point in $S \cap H$ we must have $w^* = v^*$.
 \end{proof}
@@
 \begin{proof}
-Using Part 3 of Proposition~\ref{proposition:conjugate-properties} we can write
+Using property 3 of Proposition~\ref{proposition:conjugate-properties} we can write
 the divergence
@@
 \end{align*}
-Part 2 of Proposition~\ref{proposition:conjugate-properties} implies that
+Property 2 of Proposition~\ref{proposition:conjugate-properties} implies that
 \begin{align*}

### 2015-12-08 lower-bound.tex
Added proofs for mirror descent section.
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 We show a lower bound on the worst-case regret of any algorithm for OLO. The
-proof is a standard probabilistic argument, which we present in
-Appendix~\ref{section:lower-bound-proof}.
+proof is a standard probabilistic argument, which we present
+in~\ref{section:lower-bound-proof}.
 
@@
 \|\ell_t\|_*^2})$ where $f$ is any $1$-strongly convex function with respect to
-$\|\cdot\|$.  The same upper bound is also achieved by FTRL with a constant
-learning rate when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$
-is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
-$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.
+$\|\cdot\|$.  The same upper bound is also achieved by \textsc{FTRL} with a
+constant learning rate when the number of rounds $T$ and $\sum_{t=1}^T
+\|\ell_t\|_*^2$ is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The
+lower bound is $\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.
 
@@
 example, if $K$ is the probability simplex in $\R^d$ and $f(w) = \ln(d) +
-\sum_{i=1}^d w_i \ln w_i$ is the shifted negative entropy, the
+\sum_{j=1}^d w_j \ln w_j$ is the shifted negative entropy, the
 $\|\cdot\|_1$-diameter of $K$ is $2$, $f$ is non-negative and $1$-strongly
-convex w.r.t. $\|\cdot\|_1$, but $\sup_{v \in K} f(v) = \ln(d)$.  On the other
-hand, if the norm $\|\cdot\|_2 = \sqrt{\langle \cdot, \cdot \rangle}$ arises
-from an inner product $\langle \cdot, \cdot \rangle$, the lower bound matches
-the upper bounds within a constant factor.  The reason is that for any $K$ with
-$\|\cdot\|_2$-diameter $D$, the function $f(w) = \frac{1}{2} \|w - w_0\|_2^2$,
-where $w_0$ is an arbitrary point in $K$, is $1$-strongly convex w.r.t.
-$\|\cdot\|_2$ and satisfies that $\sqrt{\sup_{v \in K} f(v)} \le D$. This leads
-to the following open problem (posed also in~\cite{Kwon-Mertikopoulos-2014}):
+convex with respect to $\|\cdot\|_1$, but $\sup_{v \in K} f(v) = \ln(d)$.  On
+the other hand, if the norm $\|\cdot\|_2 = \sqrt{\langle \cdot, \cdot \rangle}$
+arises from an inner product $\langle \cdot, \cdot \rangle$, the lower bound
+matches the upper bounds within a constant factor.  The reason is that for any
+$K$ with $\|\cdot\|_2$-diameter $D$, the function $f(w) = \frac{1}{2} \|w -
+w_0\|_2^2$, where $w_0$ is an arbitrary point in $K$, is $1$-strongly convex
+with respect to $\|\cdot\|_2$ and satisfies that $\sqrt{\sup_{v \in K} f(v)}
+\le D$. This leads to the following open problem (posed also
+in~\cite{Kwon-Mertikopoulos-2014}):
 %
@@
 \noindent
-\emph{Given a bounded convex set $K$ and a norm $\|\cdot\|$, construct a non-negative
-function $f:K \to \R$ that is $1$-strongly convex with respect to $\|\cdot\|$
-and minimizes $\sup_{v \in K} f(v)$.}
+\emph{Given a bounded convex set $K$ and a norm $\|\cdot\|$, construct a
+non-negative function $f:K \to \R$ that is $1$-strongly convex with respect to
+$\|\cdot\|$ and minimizes $\sup_{v \in K} f(v)$.}
 \end{quotation}
 %
-As shown in~\cite{Srebro-Sridharan-Tewari-2011}, the existence of $f$ with small
-$\sup_{v \in K} f(v)$ is equivalent to the existence of an algorithm for OLO with
-$\widetilde O(\sqrt{T \sup_{v \in K} f(v)})$ regret assuming $\|\ell_t\|_* \le 1$.
-The $\widetilde O$ notation hides a polylogarithmic factor in $T$.
+As shown in~\cite{Srebro-Sridharan-Tewari-2011}, the existence of $f$ with
+small $\sup_{v \in K} f(v)$ is equivalent to the existence of an algorithm for
+OLO with $\widetilde O(\sqrt{T \sup_{v \in K} f(v)})$ regret assuming
+$\|\ell_t\|_* \le 1$.  The $\widetilde O$ notation hides a polylogarithmic
+factor in $T$.

### 2015-12-08 mirror-descent.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 
-\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
-Regularized Leader} but quite different in details. The algorithm is stated as
+\textsc{Mirror Descent}\footnote{The algorithm was originally proposed in
+\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}.}
+is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
+quite different in details. The algorithm is stated as
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
@@
 \in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
-can be found in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.
-For completeness we give a proof in~\ref{section:mirror-descent-proofs}.
+can be found
+in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
+completeness we give a proof in~\ref{section:mirror-descent-proofs}.
 
@@
 Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
-prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
-- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
-$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
-\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
-A special case of projected gradient descent with step size $\eta_t =
-\frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized Infinitesimal Gradient
-Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal
-versions of \textsc{AdaGrad with Composite Mirror Descent
-Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{Mirror
-Descent} for any bounded $K$ with regularizers
+prominent example is \textsc{Projected Gradient Descent} defined by the recurrence
+$$
+w_{t+1} = \Pi_K(w_t - \eta_t \ell_t)
+$$
+where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
+$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
+an instance of \textsc{Mirror Descent} with regularizer $R_t(w) =
+\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of projected gradient descent with
+step size $\eta_t = \frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized
+Infinitesimal Gradient Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}.
+Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
+Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
+\textsc{Mirror Descent} for any bounded $K$ with regularizers
 $$
@@
 All three sequences of regularizers give a scale-free algorithm. Regularizer
-\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free generalization of
-Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2 = 1$ for all $i$,
-\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
-coincides with GIGA.
+\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
+generalization of Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2
+= 1$ for all $i$, \textsc{Mirror Descent} with regularizer
+\eqref{equation:ada-grad-regularizer} coincides with GIGA.
 
-\textsc{Exponentially-Weighted Average Forecaster}~\cite[Chapter
-2]{Cesa-Bianchi-Lugosi-2006} with constant learning rate $\eta$ for the problem
-of prediction with expert advice is an instance of \textsc{Mirror Descent} for
-the probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1
-\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j
-\ln w_j$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
-scale-free.
+\textsc{Exponentially-Weighted Average
+Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with constant learning
+rate $\eta$ for the problem of prediction with expert advice is an instance of
+\textsc{Mirror Descent} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
+\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
+\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
+algorithm is \emph{not} scale-free.
 
-Replacing $\frac{1}{2}\|w\|_2^2$ in equation \eqref{equation:ada-grad-regularizer} with
-an arbitrary convex function $R:K \to \R$ yields
+Replacing $\frac{1}{2}\|w\|_2^2$ in equation
+\eqref{equation:ada-grad-regularizer} with an arbitrary convex function $R:K
+\to \R$ yields
 \begin{equation}
@@
 We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
-regularizer $R$.  Generalizing existing regret bound for
-\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
-Proof is a generalization of the proof
-of~\cite{Duchi-Shalev-Shwartz-Singer-Tewari-2010,Duchi-Hazan-Singer-2011} and
-it can be found can be found in~\ref{section:mirror-descent-proofs}.
+regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
+\eqref{equation:scale-free-mirror-descent} does not uniquely define
+the \textsc{Mirror Descent}
+minimizer
+$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
+w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
+This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
+zero (and in particular for $t=0$). In this case, we define
+$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
+$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
+
+Generalizing existing regret bound for
+\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} (see
+also~\cite{Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}) we
+can upper bound regret of \textsc{Scale-Free Mirror Descent}.  Proof is
+in~\ref{section:mirror-descent-proofs}.
 
@@
 $$
-\Regret_T(u) \le \left( \frac{1}{\lambda} + \sup_{v \in K} \Breg_R(u,v) \right) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
+\Regret_T(u)
+\le
+\left( \frac{1}{\lambda} + \sup_{v \in K} \Breg_R(u,v) \right) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
 $$
@@
 We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
-function and optimize $\lambda$. The result is the following Corollary. Its proof is trivial.
+function and optimize $\lambda$. The result is the following Corollary. Its
+proof is trivial.
 
@@
 
-The regret bound for \textsc{Scale-Free Mirror Descent}
-in the Corollary~\ref{corollary:regret-scale-free-mirror-descent}
-depends on $\sup_{u,v \in K} \Breg_f(u,v)$. In contrast,
-the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}
-in Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
-depend on $\sup_{u \in K} f(u)$.
-Similarly, regret bound in Theorem~\ref{theorem:regret-scale-free-mirror-descent} for
-\textsc{Scale-Free Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and
-regret bounds in Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
-for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to show that
+The regret bound for \textsc{Scale-Free Mirror Descent} in the
+Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
+$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
+\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
+Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
+depend on $\sup_{u \in K} f(u)$.  Similarly, regret bound in
+Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
+Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and regret bounds in
+Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
+for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
+show that
 \begin{equation}
@@
 associated Bregman divergences $\Breg_{R'}$ and $\Breg_{R}$ are equal, $R'$ and
-$R$ have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are
-both zero.
+$R$ have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are both
+zero.
 
@@
 $\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
-$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free Mirror Descent}
-to have a ``small'' regret. We show necessity of this assumption on
-two counter-examples.  In these
-counter-examples, we consider strongly convex regularizers $R$ such that
-$\Breg_R(u,v)$ is not bounded and we construct sequences of loss vectors
-$\ell_1, \ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots
-= \|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has regret
-$\Omega(T)$ or worse.
+$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free Mirror
+Descent} to have a ``small'' regret. We show necessity of this assumption on
+two counter-examples.  In these counter-examples, we consider strongly convex
+regularizers $R$ such that $\Breg_R(u,v)$ is not bounded and we construct
+sequences of loss vectors $\ell_1, \ell_2, \dots, \ell_T$ such that
+$\|\ell_1\|_* = \|\ell_2\|_* = \dots = \|\ell_T\|_* = 1$ and \textsc{Scale-Free
+Mirror Descent} has regret $\Omega(T)$ or worse.
 
@@
 \label{theorem:first-counter-example}
-Suppose $K = V$. For any $T \ge 42$ there exists a sequence of loss vectors
+Suppose $K = V$. For any $T \ge 42$, there exists a sequence of loss vectors
 $\ell_1, \ell_2, \dots, \ell_T \in V^*$ such that $\|\ell_1\|_2 = \|\ell_2\|_2
@@
 
-The second counter-example is stated as Theorem~\ref{theorem:second-counter-example}
-below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
-\in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ and the regularizer is the
-negative entropy $R(w) = \sum_{j=1}^d w_j \ln w_j$.  Negative entropy is
-$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
-$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
-divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j
-\ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
-lower-bounded, Kullback-Leibler divergence is unbounded from above.
+The second counter-example is stated as
+Theorem~\ref{theorem:second-counter-example} below.  The decision set is the
+$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
+\sum_{j=1}^d w_j = 1 \}$ and the regularizer is the negative entropy $R(w) =
+\sum_{j=1}^d w_j \ln w_j$.  Negative entropy is $1$-strongly convex with
+respect to $\|\cdot\|_1$ and the dual norm of $\|\cdot\|_1$ is
+\mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman divergence is the
+Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j \ln(u_j/v_j)$.
+Note that despite that negative entropy is upper- and lower-bounded,
+Kullback-Leibler divergence can be arbitrarily large.
 
@@
 \label{theorem:second-counter-example}
-Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$
-be the $d$-dimensional probability simplex.  For any $T \ge 120$
-there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
-such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
-1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{j=1}^d
-w_j \ln w_j$ satisfies
+Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0,
+\sum_{j=1}^d w_j = 1 \}$ be the $d$-dimensional probability simplex.  For any
+$T \ge 120$, there exists a sequence of loss vectors $\ell_1, \ell_2, \dots,
+\ell_T \in V^*$ such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots =
+\|\ell_T\|_\infty = 1$ and \textsc{Scale-Free Mirror Descent} with regularizer
+$R(w) = \sum_{j=1}^d w_j \ln w_j$ satisfies
 $$

### 2015-12-08 per-coordinate.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 learning rates.  As shown in \cite{Streeter-McMahan-2010}, our algorithms, or
-in fact any algorithm for OLO, can be used with per-coordinate
-learning rates as well.
+in fact any algorithm for OLO, can be used with per-coordinate learning rates
+as well.
 
@@
 \times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
-factor $K_j$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
+factor $K_j$, $j=1,2,\dots,d$, we can run any OLO algorithm separately and we
 denote by $\Regret_T^{(j)}(u_j)$ its regret with respect to $u_j \in K_j$. The
@@
 each factor $K_j$, we generalize and improve existing regret bounds
-\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate
-learning rates.
+\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
+per-coordinate learning rates.

### 2015-12-08 preliminaries-proofs.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 \end{align*}
-where we have used H\"older's inequality and Part 7 of the Proposition.
+where we have used H\"older's inequality and property 7 of the Proposition.
 \end{proof}
@@
 \end{align*}
-By Proposition~\ref{proposition:conjugate-properties}, part 2, we have $w_t =
+By Proposition~\ref{proposition:conjugate-properties}, property 2, we have $w_t =
 \grad R_t^*(-L_{t-1})$ and therefore we can rewrite the sum in last expression as

### 2015-12-08 preliminaries.tex
Replace subscript "s" with "i"
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 
-Let $V$ be a finite-dimensional real vector space equipped with a norm
-$\|\cdot\|$. We denote by $V^*$ its dual vector space.  The bi-linear map
-associated with $(V^*, V)$ is denoted by $\langle \cdot, \cdot \rangle:V^*
-\times V \to \R$.  The dual norm of $\|\cdot\|$ is $\|\cdot\|_*$.
+Let $V$ be a finite-dimensional\footnote{Many, but not all, of our
+results can be extended to more general normed vector spaces.} real vector
+space equipped with a norm $\|\cdot\|$. We denote by $V^*$ its dual vector
+space.  The bi-linear map associated with $(V^*, V)$ is denoted by $\langle
+\cdot, \cdot \rangle:V^* \times V \to \R$.  The dual norm of $\|\cdot\|$ is
+$\|\cdot\|_*$.
 
@@
 
-Given $\lambda > 0$, a function $f:K \to \R$ is called \emph{$\lambda$-strongly convex}
-with respect to a norm $\|\cdot\|$ if and only if, for all $x,y \in K$,
+Given $\lambda > 0$, a function $f:K \to \R$ is called \emph{$\lambda$-strongly
+convex} with respect to a norm $\|\cdot\|$ if and only if, for all $x,y \in K$,
 $$
@@
 \label{proposition:conjugate-properties}
-Let $K \subseteq V$ be a non-empty closed convex set with diameter $D:=\sup_{u,v \in K} \|u-v\|$.
-Let $\lambda > 0$, and let $f:K \to \R$ be a lower semi-continuous function
-that is $\lambda$-strongly convex with respect to $\|\cdot\|$.
-The Fenchel conjugate of $f$ satisfies:
+Let $K \subseteq V$ be a non-empty closed convex set with diameter
+$D:=\sup_{u,v \in K} \|u-v\|$.  Let $\lambda > 0$, and let $f:K \to \R$ be a
+lower semi-continuous function that is $\lambda$-strongly convex with respect
+to $\|\cdot\|$.  The Fenchel conjugate of $f$ satisfies:
 \begin{enumerate}
@@
 \begin{algorithmic}[1]
+\REQUIRE Non-empty closed convex set $K \subseteq V$
 \STATE Initialize $L_0 \leftarrow 0$
@@
 Our scale-free online learning algorithms are versions of the \textsc{Follow
-The Regularized Leader} (\textsc{FTRL}) algorithm with varying regularizers,
-presented as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The following
-lemma bounds its regret.
+The Regularized Leader}\footnote{The name \textsc{Follow The Regularized
+Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}. Independently, the
+algorithm was proposed by~\cite{Nestorov-2009} under the name \textsc{Dual
+Averaging} and was later anlyzed in~\cite{Xiao-2010}.} (\textsc{FTRL})
+algorithm with varying regularizers, presented as
+Algorithm~\ref{algorithm:ftrl-varying-regularizer}. The algorithm is
+paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
+\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
+past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
+following lemma bounds its regret.
 
@@
 
-Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}. For completeness,
-we include it in~\ref{section:definitions-proofs}.
+Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}.
+For completeness, we include it in~\ref{section:definitions-proofs}.

### 2015-12-08 solo-ftrl.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

@@
 
-The closest algorithm to a scale-free one in the OLO literature is the \textsc{AdaGrad}
-algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a regularizer on each
-coordinate of the form
+The closest algorithm to a scale-free one in the OLO literature is the
+\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
+regularizer on each coordinate of the form
 \begin{equation*}
@@
 knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
-other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the norm of
-the loss vectors. Identical considerations hold for the \textsc{FTRL-Proximal} in
-\cite{McMahan-Streeter-2010,McMahan-2014}: the scale-free setting of the
-learning rate is valid only in the bounded case.
+other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
+norm of the loss vectors. Identical considerations hold for the
+\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
+scale-free setting of the learning rate is valid only in the bounded case.
 
@@
 the additional parameter of the doubling power.}, such strategy would give the
-feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a scale-free
-algorithm.
+feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a
+scale-free algorithm.
 
@@
 \begin{equation*}
+\label{equation:solo-ftrl-regularizer}
 R_t(w) = R(w) \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \; ,
@@
 Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
-result is the following Theorem, which is proven in
+result is the Theorem~\ref{theorem:regret-solo-ftrl} below, which is proven in
 Section~\ref{section:solo-ftrl-regret-bound}.
+
+The regularizer~\eqref{equation:solo-ftrl-regularizer} does not uniquely define
+the \textsc{FTRL} minimizer $w_t = \argmin_{w \in K} R_t(w)$ when
+$\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}$ is zero. This happens if $\ell_1,
+\ell_2, \dots, \ell_{t-1}$ are all zero (and in particular for $t=1$).  In that
+case, we define the $w_t = \argmin_{w \in K} R(w)$ which is consistent with
+$w_t = \lim_{a \to 0^+}  \argmin_{w \in K} a R(w)$.
 
@@
 & \le \left( R(u) + \frac{2.75}{\lambda}\right) \sqrt{\sum_{t=1}^{T} \norm{\ell_t}_*^2}
-+ 3.5 \min\left\{\frac{\sqrt{T-1}}{\lambda} , D\right\} \max_{t \le T} \|\ell_t\|_*.
++ 3.5 \min\left\{\frac{\sqrt{T-1}}{\lambda} , D\right\} \max_{t \le T} \|\ell_t\|_* \; .
 \end{align*}
@@
 in~\ref{section:solo-ftrl-proofs}.  It is similar to
-Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}. The scaling
-however is different in the two corollaries.  In
+Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}. The
+scaling however is different in the two corollaries.  In
 Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in
@@
 Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
-that $f:K \to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
-convex with respect to $\|\cdot\|$. \textsc{SOLO FTRL} with regularizer
+that $f:K \to \R$ is a non-negative lower semi-continuous function that is
+$1$-strongly convex with respect to $\|\cdot\|$. \textsc{SOLO FTRL} with
+regularizer
 $$
@@
 \label{lemma:useful}
-Let $C, a_1, a_2, \dots, a_T\geq0$. Then,
+Let $C, a_1, a_2, \dots, a_T \ge 0$. Then,
 $$
@@
 
-\begin{lemma}
+\begin{lemma}[Simple Inequality]
 \label{lemma:sum-of-square-roots-inverses}
@@
 %
-Lemma~\ref{lemma:sum-of-square-roots-inverses} is from \cite[Lemma
-3.5]{Auer-Cesa-Bianchi-Gentile-2002}.  We give
-a proof in~\ref{section:solo-ftrl-proofs} for completeness.
+Lemma~\ref{lemma:sum-of-square-roots-inverses} is from
+\cite[Lemma~3.5]{Auer-Cesa-Bianchi-Gentile-2002}.  We give a proof
+in~\ref{section:solo-ftrl-proofs} for completeness.
 
 \begin{proof}[Proof of Theorem~\ref{theorem:regret-solo-ftrl}]
-Let $\eta_t=\tfrac{1}{\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}$, hence $R_t(w) = \tfrac{1}{\eta_t} R(w)$.
-We assume without loss of
-generality that $\|\ell_t\|_* > 0$ for all $t$, since otherwise we can remove
-all rounds $t$ where $\ell_t = 0$ without affecting regret and the
-predictions of the algorithm on the remaining rounds.
-By Lemma~\ref{lemma:generic-regret-bound},
+Let $\eta_t = \frac{1}{\sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2}}$, hence $R_t(w)
+= \frac{1}{\eta_t} R(w)$.  We assume without loss of generality that
+$\|\ell_t\|_* > 0$ for all $t$, since otherwise we can remove all rounds $t$
+where $\ell_t = 0$ without affecting regret and the predictions of the
+algorithm on the remaining rounds.  By Lemma~\ref{lemma:generic-regret-bound},
 \begin{align*}

### 2015-12-08 ada-ftrl-proofs.tex
Added proofs for mirror descent section.
Use small-caps for all algorithms names
Make AdaFTRL recurrence a separate lemma

@@
-\section{Proofs for AdaFTRL}
-\label{section:ada-ftrl-proof}
+\section{Proofs for \textsc{AdaFTRL}}
+\label{section:ada-ftrl-proofs}
 
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
-Let $a_t = \|\ell_t\|_* \max\{D, 1/\sqrt{2\lambda}\}$. The statement of the
-lemma is equivalent to $\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2}$ which we
-prove by induction on $T$.  The base case $T=0$ is trivial. For $T \ge 1$, we
-have
+We prove the lemma by induction on $T$.  The base case $T=0$ is trivial. For $T
+\ge 1$, we have
 $$

### 2015-12-08 ada-ftrl.tex
Make AdaFTRL recurrence a separate lemma

@@
 \label{lemma:initial-regret-bound}
-\textsc{AdaFTRL}, for any $u \in K$ and any $u \ge 0$, satisfies
-$\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T$.
+\textsc{AdaFTRL} satisfies, for any $u \in K$ and any $T \ge 0$,
+$$
+\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T \; .
+$$
 \end{lemma}
@@
 \label{lemma:gap-recurrence}
-Let $D = \sup_{u, v \in K} \|u -v\|$ be the diameter of $K$.
-The sequence $\{\Delta_t\}_{t=1}^\infty$ generated by \textsc{AdaFTRL} satisfies for any $t \ge 1$,
+Let $D = \sup_{u, v \in K} \|u -v\|$ be the diameter of $K$.  The sequence
+$\{\Delta_t\}_{t=1}^\infty$ generated by \textsc{AdaFTRL} satisfies for any $t
+\ge 1$,
 $$
@@
 \begin{proof}
-The inequality results from strong convexity of $R_t(w)$ and
-Proposition~\ref{proposition:conjugate-properties}.
+By definition, $\Delta_t$ satisfies the recurrence
+$$
+\Delta_t = \Delta_{t-1} + \Delta_{t-1} \Breg_{R^*}\left( - \frac{L_t}{\Delta_{t-1}}, - \frac{L_{t-1}}{\Delta_{t-1}} \right) \; .
+$$
+Using parts 4 and 6 of Proposition~\ref{proposition:conjugate-properties}, we
+can upper bound $\Breg_{R^*}\left( - \frac{L_t}{\Delta_{t-1}}, -
+\frac{L_{t-1}}{\Delta_{t-1}} \right)$ with two different quantities.  Taking
+minimum of the two quantities finishes the proof.
 \end{proof}
 
-\begin{lemma}[Solution of the Recurrence]
+The recurrence of Lemma~\ref{lemma:gap-recurrence} can be simplified. Defining
+$$
+a_t = \|\ell_t\|_* \max \left\{D, \frac{1}{\sqrt{2 \lambda}} \right\} \; ,
+$$
+we get a recurrence
+$$
+\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \; .
+$$
+The next lemma solves this recurrence, by giving an explicit upper bound on
+$\Delta_T$ in terms of $a_1, a_2, \dots, a_T$.
+
+\begin{lemma}[Recurrence]
 \label{lemma:recurrence-solution}
-Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
-generated by \textsc{AdaFTRL} satisfies for any $T \ge 0$,
+Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
+Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of non-negative real
+numbers satisfying
 $$
-\Delta_T \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
+\Delta_0 = 0 \qquad \text{and} \qquad
+\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
+$$
+Then, for any $T \ge 0$,
+$$
+\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
 $$
@@
 follows from
-Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.
+Lemmas~\ref{lemma:initial-regret-bound},~\ref{lemma:gap-recurrence}~and~\ref{lemma:recurrence-solution}.

### 2015-12-08 abstract.tex
Added one sentence to the abstract.
Fix abstract
Added sentence explaining lower bounds for mirror decsent.
Reworked introduction

@@
 algorithms work for any decision set, bounded or unbounded.  For unbounded
-decisions sets, these are the first truly adaptive algorithms for online linear
+decisions sets, these are the first adaptive algorithms for online linear
 optimization.
+
+We also study a popular scale-free variant of online mirror descent algorithm,
+and we show that in two natural settings it has linear or worse regret.

### 2015-12-08 introduction.tex
Reworked introduction

@@
 
-The contribution of this paper is twofold. First, in
-Section~\ref{section:ada-ftrl} we show that the analysis and design of
-\textsc{AdaHedge} can be generalized to the OLO scenario and to any strongly
-convex regularizer, in an algorithm we call \textsc{AdaFTRL}, providing a new
-and rather interesting way to adapt the learning rates to have scale-free
-algorithms.  Second, in Section~\ref{section:solo-ftrl} we propose a new and
-simple algorithm, \textsc{SOLO FTRL}, that is scale-free and is the
-\textbf{first} scale-free online algorithm for unbounded sets with a
-non-vacuous regret bound.  Both algorithms are instances of \textsc{Follow The
-Regularized Leader} (\textsc{FTRL}) with an adaptive learning rate.  Moreover,
-our algorithms show that scale-free algorithms can be obtained in a ``native''
-and simple way, i.e.  without using ``doubling tricks'' that attempt to fix
-poorly designed algorithms rather than directly solving the problem.
+We design and analyze three scale-free algorithms. First two algorithms, which
+we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
+\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive learning
+rate.  \textsc{AdaFTRL} is a generalization of
+\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
+strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
+``correct'' scale-free version of \textsc{AdaGrad
+FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
+regularizers.  The third algorithm, which we call \textsc{Scale-Free Mirror
+Descent}, is a variant of \textsc{Mirror Descent}. It is a generalization of
+\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to arbitrary strongly convex
+regularizers, similar to \textsc{Optimistic MD}~\cite{Rakhlin-Sridharan-2013}.
+The three algorithms are presented and analyzed in
+Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
+\ref{section:mirror-descent} respectively.
 
-For both algorithms, we prove that for bounded decision sets the regret after
-$T$ rounds is at most $O(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2})$.  We show that the
-$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving a $\Omega (D
-\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} )$ lower bound on the regret of any algorithm
-for OLO for any decision set with diameter $D$ with respect to the primal norm
-$\|\cdot\|$. For the \textsc{SOLO FTRL} algorithm, we prove an
-$O(\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T})$ regret bound for any unbounded
-decision set.
+For all three algorithms, we prove that for bounded decision sets the regret
+after $T$ rounds is at most $O \left(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$
+where the constant hidden in $O(\cdot)$ depends on the decision set and the
+regularizer.  In Section~\ref{section:lower-bound}, we show that the
+$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving $\Omega
+\left(D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound on the regret of
+any algorithm for OLO for any decision set with diameter $D$ with respect to
+the primal norm $\|\cdot\|$.
 
-Our algorithms are also \textbf{any-time}, i.e., do not need to know the
-number of rounds in advance and our regret bounds hold for all time steps
+For the \textsc{SOLO FTRL} algorithm, we prove an $O \left(\max_{t=1,2,\dots,T}
+\|\ell_t\|_* \sqrt{T} \right)$ regret bound for any unbounded decision set. The
+constant hidden in $O(\cdot)$ depends not only on the decision set and the
+regularizer, but also on the competitor $u \in K$.~\footnote{An upper bound
+that does not depend on $u$ is impossible. It would contradict the $\Omega
+\left( D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound.} This is the
+\textbf{first adaptive algorithm for unbounded decision sets} with a
+non-trivial upper bound.
+
+All three algorithms are \textbf{any-time}, i.e., they do not need to know the
+number of rounds in advance and the regret bounds hold for all time steps
 simultaneously.
+
+Our proofs of the regret upper bounds rely on homogeneous inequalities
+(Lemmas~\ref{lemma:recurrence-solution},
+\ref{lemma:sum-of-square-roots-inverses}, \ref{lemma:useful}).  Two of the
+inequalities are new. Previously, many proofs of regret bounds relied on ad-hoc
+non-homogeneous inequalities, or the existing algorithms used quantities (e.g.
+$\sum_{t=1}^T \|\ell_t\|_*^2$) that are not available until all loss vectors
+are processed.
+
+Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
+negative results for existing popular variants of \textsc{Mirror Descent}. We
+show that variants of \textsc{Mirror Descent} have $\Omega(T)$ regret if the
+Bregman divergence associated with the regularizer is unbounded on the decision
+set.  The setting includes the standard gradient descent with step size
+$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free Mirror Descent} on a
+unbounded decision set. This result suggests that \textsc{Follow The
+Regularized Leader} is superior to \textsc{Mirror Descent}.

### 2015-12-09 abstract.tex
typos and small changes

@@
 and at the same time do not need to know any upper or lower bounds on the norm
-of the loss vectors.  We achieve adaptiveness to norms of loss vectors by scale
+of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors by scale
 invariance, i.e., our algorithms make exactly the same decisions if the
@@
 decisions sets, these are the first adaptive algorithms for online linear
-optimization.
+optimization with a non-vacuous bound.
 

### 2015-12-09 conclusions.tex
typos and small changes

@@
 we gave examples of regularizers for which the associated Bregman divergence is
-unounded and the scale-fre version of \textsc{Mirror Descent} has $\Omega(T)$
+unounded and the scale-free version of \textsc{Mirror Descent} has $\Omega(T)$
 regret or worse.

### 2015-12-09 introduction.tex
typos and small changes

@@
 chooses a point $w_t$ from a convex decision set $K$, observes an arbitrary, or
-even adversarially chosen, loss vector $\ell_t$ and suffers loss $\langle
+even adversarially chosen, loss vector $\ell_t$ and suffers the loss $\langle
 \ell_t, w_t \rangle$.  The goal of the algorithm is to have a small cumulative
-loss. Performance of an algorithm is evaluated by the so-called regret, which
-is the difference of cumulative losses of the algorithm and of the
+loss. The performance of an algorithm is evaluated by the so-called regret, which
+is the difference of the cumulative losses of the algorithm and of the
 (hypothetical) strategy that would choose in every round the same best point in
@@
 and/or decision sets, in this paper we explicitly focus on the scale-free
-property. Regret of scale-free algorithms is proportional to the scale of the
+property. The regret of scale-free algorithms is proportional to the scale of the
 losses, ensuring optimal linear dependency on the maximum norm of the loss
@@
 we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
-\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive learning
-rate.  \textsc{AdaFTRL} is a generalization of
+\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive regularizers.
+\textsc{AdaFTRL} is a generalization of
 \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary

### 2015-12-09 mirror-descent.tex
Reworked introduction
typos and small changes

@@
 \textsc{Mirror Descent}\footnote{The algorithm was originally proposed in
-\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}.}
+\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for convex optimization.
+In the online literature it makes the first appeareance, under a different name, in \cite{Kivinen-Warmuth-1997}.}
 is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
@@
 If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, Bregman
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the Bregman
 divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
@@
 
-Generalizing existing regret bound for
-\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} (see
-also~\cite{Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}) we
-can upper bound regret of \textsc{Scale-Free Mirror Descent}.  Proof is
+We can upper bound regret of \textsc{Scale-Free Mirror Descent} (see also \cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).  Proof is
 in~\ref{section:mirror-descent-proofs}.
@@
 
-We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
+We choose the regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
 function and optimize $\lambda$. The result is the following Corollary. Its
@@
 Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
-depend on $\sup_{u \in K} f(u)$.  Similarly, regret bound in
+depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
 Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
-Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and regret bounds in
+Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
 Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
@@
 Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
-factors, regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
+factors, the regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
 better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
@@
 \subsection{Lower Bounds}
+\label{subsection:mirror-descent-lower-bound}
 
@@
 that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
-norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. Corresponding Bregman divergence is
+norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
 $\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs

### 2015-12-09 per-coordinate.tex
typos and small changes

@@
-\section{Per-Coordinate Learning}
+\subsection{Per-Coordinate Learning}
 
@@
 \cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
-learning rates.  As shown in \cite{Streeter-McMahan-2010}, our algorithms, or
-in fact any algorithm for OLO, can be used with per-coordinate learning rates
-as well.
+learning rates.  As shown in \cite{Streeter-McMahan-2010}, any algorithm for OLO can be used with per-coordinate learning rates as well.
 
@@
 $$
+
 If the algorithm for each factor is scale-free, the overall algorithm is
-clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
-each factor $K_j$, we generalize and improve existing regret bounds
-\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
-per-coordinate learning rates.
+clearly scale-free as well. Hence, even if not explicitly mentioned in the text, any algorithm we present can be trivially transformed in a per-coordinate version.
+
+%Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
+%each factor $K_j$, we generalize and improve existing regret bounds
+%\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
+%per-coordinate learning rates.

### 2015-12-09 preliminaries.tex
typos and small changes

@@
 The Regularized Leader}\footnote{The name \textsc{Follow The Regularized
-Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}. Independently, the
-algorithm was proposed by~\cite{Nestorov-2009} under the name \textsc{Dual
-Averaging} and was later anlyzed in~\cite{Xiao-2010}.} (\textsc{FTRL})
-algorithm with varying regularizers, presented as
+Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
+It dates back to the potential-based forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was heavily developed by \cite{Shalev-Shwartz-2007}.
+Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for convex optimization under the name \textsc{Dual
+Averaging} and later rediscovered in~\cite{Xiao-2010} for online convex optimization. The version with time-varying regularizers was analyzed in \cite{Orabona-Crammer-Cesa-Bianchi-2014}.
+} (\textsc{FTRL})
+algorithm with \emph{varying regularizers}, presented as
 Algorithm~\ref{algorithm:ftrl-varying-regularizer}. The algorithm is
@@
 For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
-semi-continuous regularizers, regret of
+semi-continuous regularizers, the regret of
 Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
@@
 
-Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}.
+The proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}.
 For completeness, we include it in~\ref{section:definitions-proofs}.

### 2015-12-09 solo-ftrl.tex
typos and small changes

@@
 other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
-norm of the loss vectors. Identical considerations hold for the
+norm of the gradient vectors. Identical considerations hold for the
 \textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
@@
 the estimate. Finally, the algorithm would depend on an additional parameter,
-the ``doubling'' power. Hence, even guaranteeing a regret bound\footnote{For
-lack of space, we cannot include the regret bound for the doubling trick
-version. It would be exactly the same as in
-Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
-the additional parameter of the doubling power.}, such strategy would give the
+the ``doubling'' power. Hence, even in the case one would prove a regret bound,
+%\footnote{For
+%lack of space, we cannot include the regret bound for the doubling trick
+%version. It would be exactly the same as in
+%Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
+%the additional parameter of the doubling power.},
+such strategy would give the
 feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a

### 2015-12-11 introduction.tex
Expand introduction

@@
 \ell_t, w_t \rangle$.  The goal of the algorithm is to have a small cumulative
-loss. The performance of an algorithm is evaluated by the so-called regret, which
-is the difference of the cumulative losses of the algorithm and of the
+loss. The performance of an algorithm is evaluated by the so-called regret,
+which is the difference of the cumulative losses of the algorithm and of the
 (hypothetical) strategy that would choose in every round the same best point in
@@
 
-A summary of algorithms for OLO is presented in Table~\ref{table:results}.
-While the scale-free property has been looked at in the expert setting, in the
-general OLO setting this issue has been largely ignored.  In particular, the
-\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm, for
-prediction with expert advice, is specifically designed to be scale-free.  A
-notable exception in the OLO literature is the discussion of the ``off-by-one''
-issue in~\cite{McMahan-2014}, where it is explained that even the popular
-\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011} is not completely
-adaptive; see also our discussion in Section~\ref{section:solo-ftrl}. In
-particular, existing scale-free algorithms cover only some norms/regularizers
-and \emph{only} bounded decision sets. The case of \textbf{unbounded decision
-sets}, practically the most interesting one for machine learning applications,
-remains completely unsolved.
+% \begin{definition}[Scale-Free Algorithm]
+% \label{definition:scale-free-algorithm}
+% An algorithm for OLO is \emph{scale-free} if for any sequence of loss vectors
+% $\{\ell_t\}_{t=1}^\infty$ and any real number $c > 0$, on the sequences of
+% $\{\ell_t\}_{t=1}^\infty$ and $\{c \cdot \ell_t\}_{t=1}^\infty$, the algorithm
+% chooses the same sequence of points $\{w_t\}_{t=1}^\infty$.
+% \end{definition}
 
-Rather than trying to design strategies for a particular form of loss vectors
-and/or decision sets, in this paper we explicitly focus on the scale-free
-property. The regret of scale-free algorithms is proportional to the scale of the
-losses, ensuring optimal linear dependency on the maximum norm of the loss
-vectors.
+The most important property of scale-free algorithms is that both their loss
+and their regret is scales linearly with the maximum norm of the loss vector
+appearing in the sequence.
 
-We design and analyze three scale-free algorithms. First two algorithms, which
-we call \textsc{AdaFTRL} and \textsc{SOLO FTRL}, are instances of
-\textsc{Follow The Regularized Leader} (\textsc{FTRL}) with adaptive regularizers.
-\textsc{AdaFTRL} is a generalization of
+\subsection{Previous results}
+
+Majority of existing algorithms for OLO are based on two generic algorithms:
+\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
+Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
+forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
+developed by \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
+Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
+Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for
+convex optimization under the name \textsc{Dual Averaging} and later
+rediscovered in~\cite{Xiao-2010} for online convex optimization. The general
+version with time-varying regularizers was analyzed in
+\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
+in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
+for convex optimization. In the online literature it makes the first
+appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. Good
+description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.
+
+Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called a
+\emph{regularizer}. The function is assumed to be strongly convex with respect
+to a given norm $\|\cdot\|$. Based on different regularizers different
+algorithms with different properties can be instantiated. A summary of
+algorithms for OLO is presented in Table~\ref{table:results}.  All of them,
+except for \textsc{Hedge}, are instances of \textsc{FTRL} or \textsc{MD}.
+
+Scale-free versions of \textsc{MD} include \textsc{AdaGrad
+MD}~\cite{Duchi-Hazan-Singer-2011} and~\textsc{Optimistic
+MD}~\cite{Rakhlin-Sridharan-2013}. However, these algorithms have non-trivial
+regret bounds only when the Bregman divergence associated with the regularizer
+is bounded, which happens only when the decision is bounded as well. As we show
+in Section~\ref{subsection:mirror-descent-lower-bound}, this is unavoidable,
+since \textsc{AdaGrad MD} and similar algorithms based on \textsc{MD} can have
+$\Omega(T)$ regret if Bregman divergence is not bounded.
+
+Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
+\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
+learning with expert advice, where the decision set is bounded.
+Algorithms based on \textsc{FTRL} that are ``almost'' scale-free include
+\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} and
+\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  These algorithms fail to
+be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and also our
+discussion in Section~\ref{section:solo-ftrl}.
+
+For unbounded decision sets no scale-free algorithm with a non-trivial regret
+bound was known. Unbounded decision sets are practically important, since
+learning large-scale linear models (e.g. logistic regression) can be reduced to
+OLO with $\R^d$ as the decision set. One of our contributions is an
+\textsc{FTRL}-based scale-free algorithm with a non-trivial regret bound for
+any regularizer and any decision set, bounded or unbounded.
+
+% While the scale-free property has been looked at in the expert setting, in
+% the general OLO setting this issue has been largely ignored.  In particular,
+% the \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014}
+% algorithm, for prediction with expert advice, is specifically designed to be
+% scale-free.  A notable exception in the OLO literature is the discussion of
+% the ``off-by-one'' issue in~\cite{McMahan-2014}, where it is explained that
+% even the popular \textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011} is
+% not completely adaptive; see also our discussion in
+% Section~\ref{section:solo-ftrl}. In particular, existing scale-free
+% algorithms cover only some norms/regularizers and \emph{only} bounded
+% decision sets. The case of \textbf{unbounded decision sets}, practically the
+% most interesting one for machine learning applications, remains completely
+% unsolved.
+
+\subsection{Overview of the Results}
+
+We design and analyze three scale-free algorithms: \textsc{AdaFTRL},
+\textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  \textsc{AdaFTRL} and
+\textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a
+generalization of
 \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
@@
 FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
-regularizers.  The third algorithm, which we call \textsc{Scale-Free Mirror
-Descent}, is a variant of \textsc{Mirror Descent}. It is a generalization of
-\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to arbitrary strongly convex
-regularizers, similar to \textsc{Optimistic MD}~\cite{Rakhlin-Sridharan-2013}.
-The three algorithms are presented and analyzed in
-Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
+regularizers.  \textsc{Scale-Free MD} is based on \textsc{MD}. It is a
+generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
+arbitrary strongly convex regularizers, similar to \textsc{Optimistic
+MD}~\cite{Rakhlin-Sridharan-2013}.  The three algorithms are presented and
+analyzed in Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
 \ref{section:mirror-descent} respectively.
 
-For all three algorithms, we prove that for bounded decision sets the regret
-after $T$ rounds is at most $O \left(\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$
-where the constant hidden in $O(\cdot)$ depends on the decision set and the
-regularizer.  In Section~\ref{section:lower-bound}, we show that the
-$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term is necessary by proving $\Omega
-\left(D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound on the regret of
-any algorithm for OLO for any decision set with diameter $D$ with respect to
-the primal norm $\|\cdot\|$.
+We prove that the regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} after $T$
+rounds is bounded by $O \left(\sqrt{\sup_{v \in K} f(v)
+\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $f$ is a non-negative regularizer that is
+$1$-strongly convex with respect to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its
+dual norm. For \textsc{Scale-Free MD}, we prove $O \left(\sqrt{\sup_{u,v \in K}
+B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $B_f$ is the Bregman
+divergence associated with a $1$-strongly convex regularizer $f$. In
+Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
+\|\ell_t\|_*^2}$ term in the bounds term is necessary by proving
+$\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
+regret of any algorithm for OLO for any decision set with diameter $D$ with
+respect to the primal norm $\|\cdot\|$.
 
-For the \textsc{SOLO FTRL} algorithm, we prove an $O \left(\max_{t=1,2,\dots,T}
-\|\ell_t\|_* \sqrt{T} \right)$ regret bound for any unbounded decision set. The
-constant hidden in $O(\cdot)$ depends not only on the decision set and the
-regularizer, but also on the competitor $u \in K$.~\footnote{An upper bound
-that does not depend on $u$ is impossible. It would contradict the $\Omega
-\left( D \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ lower bound.} This is the
-\textbf{first adaptive algorithm for unbounded decision sets} with a
+For \textsc{SOLO FTRL}, we prove that regret against a competitor $u \in K$ is
+at most $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
+\|\ell_t\|_* \sqrt{T} \right)$.  Same as before, $f$ is a non-negative
+$1$-strongly convex regularizer. The decision can be bounded or unbounded. This
+is the \textbf{first adaptive algorithm for unbounded decision sets} with a
 non-trivial upper bound.
@@
 Our proofs of the regret upper bounds rely on homogeneous inequalities
-(Lemmas~\ref{lemma:recurrence-solution},
-\ref{lemma:sum-of-square-roots-inverses}, \ref{lemma:useful}).  Two of the
-inequalities are new. Previously, many proofs of regret bounds relied on ad-hoc
-non-homogeneous inequalities, or the existing algorithms used quantities (e.g.
-$\sum_{t=1}^T \|\ell_t\|_*^2$) that are not available until all loss vectors
-are processed.
+(Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful},
+\ref{lemma:sum-of-square-roots-inverses}).  Previously, many proofs of regret
+bounds relied on ad-hoc non-homogeneous inequalities, or the existing
+algorithms used quantities (e.g.  $\sum_{t=1}^T \|\ell_t\|_*^2$) that are not
+available until all loss vectors are processed. Two of the inequalities
+(Lemmas~\ref{lemma:recurrence-solution} and \ref{lemma:useful}) are new and
+interesting in their own right.
 
 Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
-negative results for existing popular variants of \textsc{Mirror Descent}. We
-show that variants of \textsc{Mirror Descent} have $\Omega(T)$ regret if the
-Bregman divergence associated with the regularizer is unbounded on the decision
-set.  The setting includes the standard gradient descent with step size
-$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free Mirror Descent} on a
-unbounded decision set. This result suggests that \textsc{Follow The
-Regularized Leader} is superior to \textsc{Mirror Descent}.
+negative results for existing popular variants of \textsc{MD}. We
+show that variants of \textsc{MD} have $\Omega(T)$ regret if the
+Bregman divergence associated with the regularizer is unbounded.
+The setting includes the standard gradient descent with step size
+$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a
+unbounded decision set. This result suggests that \textsc{FTRL}
+is superior to \textsc{MD}.

### 2015-12-11 mirror-descent.tex
Expand introduction

@@
 
-\textsc{Mirror Descent}\footnote{The algorithm was originally proposed in
-\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for convex optimization.
-In the online literature it makes the first appeareance, under a different name, in \cite{Kivinen-Warmuth-1997}.}
-is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
+\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
 quite different in details. The algorithm is stated as

### 2015-12-11 preliminaries.tex
Expand introduction

@@
 Our scale-free online learning algorithms are versions of the \textsc{Follow
-The Regularized Leader}\footnote{The name \textsc{Follow The Regularized
-Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
-It dates back to the potential-based forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was heavily developed by \cite{Shalev-Shwartz-2007}.
-Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for convex optimization under the name \textsc{Dual
-Averaging} and later rediscovered in~\cite{Xiao-2010} for online convex optimization. The version with time-varying regularizers was analyzed in \cite{Orabona-Crammer-Cesa-Bianchi-2014}.
-} (\textsc{FTRL})
-algorithm with \emph{varying regularizers}, presented as
-Algorithm~\ref{algorithm:ftrl-varying-regularizer}. The algorithm is
-paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
-\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
-past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
-following lemma bounds its regret.
+The Regularized Leader} (\textsc{FTRL}) algorithm with \emph{varying
+regularizers}, presented as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.
+The algorithm is paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex
+functions $R_t:K \to \R$ called \emph{regularizers}.  Each regularizer $R_t$
+can depend on the past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an
+arbitrary way.  The following lemma bounds its regret.
 

### 2015-12-11 conclusions.tex
Reword and reorganize some sentences

--- 2015-12-08 conclusions.tex
@@
 of loss vectors, i.e. without any assumption on the norms of the loss vectors
-and the number of rounds. We have designed two scale-free algorithms for online linear
-optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
-\|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly convex
-function with respect to a norm $\|\cdot\|$ defined on the decision set and
-where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
-based on \textsc{Follow The Regularizer Leader}.  We have also designed an
-algorithm with regret $O\left((R(u) + \frac{1}{\lambda}) \sqrt{T}
-\max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any competitor $u
-\in K$, where $R$ is any non-negative $\lambda$-strongly convex function
-defined on the decision set.  The latter result makes sense even when the
-decision set $K$ is unbounded.
+and the number of rounds. We have designed two scale-free algorithms for online
+linear optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v)
+\sum_{t=1}^T \|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly
+convex function with respect to a norm $\|\cdot\|$ defined on the decision set
+and where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
+based on \textsc{Follow The Regularizer Leader}.  We have shown that one of the
+algorithms has regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
+\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T} \right)$ with respect to any
+competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex
+function defined on the decision set.  The latter result makes sense even when
+the decision set $K$ is unbounded.
 
@@
 we gave examples of regularizers for which the associated Bregman divergence is
-unounded and the scale-fre version of \textsc{Mirror Descent} has $\Omega(T)$
+unbounded and the scale-free version of \textsc{Mirror Descent} has $\Omega(T)$
 regret or worse.
 
-We have shown an $\Omega\left( D \sqrt{\sum_{t=1}^T \|\ell\|_*^2}\right)$
+We have shown an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
 lower bound on regret of any algorithm for decision set with diameter $D$.

### 2015-12-11 introduction.tex
Reword and reorganize some sentences

@@
 forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
-developed by \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
+developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
 Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
-Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for
+Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
 convex optimization under the name \textsc{Dual Averaging} and later
@@
 in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
-for convex optimization. In the online literature it makes the first
+for convex optimization. In the online learning literature it makes the first
 appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. Good

### 2015-12-11 mirror-descent.tex
Reword and reorganize some sentences

@@
-\section{Mirror Descent}
+\section{Scale-Free Mirror Descent}
 \label{section:mirror-descent}
@@
 
-\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
-quite different in details. The algorithm is stated as
+\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{Follow The
+Regularized Leader} but quite different in details. The algorithm is stated as
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
@@
 If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the Bregman
-divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
-\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
-R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
-$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
+Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
+u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
+$\grad R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient
+of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
 $R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
+\in K$. The following lemma bounds regret of \textsc{MD}.  Proof
 can be found
@@
 
-Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
-prominent example is \textsc{Projected Gradient Descent} defined by the recurrence
+Several existing algorithms are special cases of \textsc{MD}.
+For example, \textsc{Exponentially-Weighted Average
+Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with a constant learning
+rate $\eta$ for learning with expert advice is an instance of
+\textsc{MD} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
+\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
+\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
+algorithm is \emph{not} scale-free.
+
+The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
+Descent} defined by the recurrence
 $$
@@
 $K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
-an instance of \textsc{Mirror Descent} with regularizer $R_t(w) =
+an instance of \textsc{MD} with regularizer $R_t(w) =
 \frac{1}{2\eta_t}\|w\|_2^2$.  A special case of projected gradient descent with
@@
 Infinitesimal Gradient Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}.
-Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
-Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
-\textsc{Mirror Descent} for any bounded $K$ with regularizers
+Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
+Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
+\textsc{MD} for any bounded $K$ with regularizers
 $$
@@
 $$
-respectively. When $d=1$, both algorithms are equivalent to
+respectively. The diagonal version is simply the per-coordinate construction
+applied to the one-dimensional version. When $d=1$, both algorithms are
+equivalent to
 \begin{equation}
@@
 \eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
-generalization of Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2
-= 1$ for all $i$, \textsc{Mirror Descent} with regularizer
+generalization of Zinkevich's GIGA algorithm. That is, when $\|\ell_i\|_2
+= 1$ for all $i$, \textsc{MD} with regularizer
 \eqref{equation:ada-grad-regularizer} coincides with GIGA.
-
-\textsc{Exponentially-Weighted Average
-Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with constant learning
-rate $\eta$ for the problem of prediction with expert advice is an instance of
-\textsc{Mirror Descent} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
-\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
-\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
-algorithm is \emph{not} scale-free.
 
@@
 \end{equation}
-We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
+We call the resulting algorithm \textsc{Scale-Free MD} with
 regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
 \eqref{equation:scale-free-mirror-descent} does not uniquely define
-the \textsc{Mirror Descent}
-minimizer
+the \textsc{MD} minimizer
 $w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
@@
 
-We can upper bound regret of \textsc{Scale-Free Mirror Descent} (see also \cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).  Proof is
-in~\ref{section:mirror-descent-proofs}.
+The theorem below upper bounds the regret of \textsc{Scale-Free Mirror Descent}
+(see also
+\cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).
+Proof is in~\ref{section:mirror-descent-proofs}.
 

### 2015-12-11 preliminaries.tex
Reword and reorganize some sentences

@@
 $$
-where $\grad f(x)$ is any subgradient of $f$ at point $x$.
+where $\grad f(x)$ is any subgradient of $f$ at the point $x$.
 
@@
 
-\item $f^*$ is finite everywhere and differentiable.
+\item $f^*$ is finite everywhere and differentiable everywhere.
 
@@
 
-Our scale-free online learning algorithms are versions of the \textsc{Follow
-The Regularized Leader} (\textsc{FTRL}) algorithm with \emph{varying
-regularizers}, presented as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.
-The algorithm is paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex
-functions $R_t:K \to \R$ called \emph{regularizers}.  Each regularizer $R_t$
-can depend on the past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an
-arbitrary way.  The following lemma bounds its regret.
+Our scale-free online learning algorithms are instances of \textsc{FTRL} with
+\emph{varying regularizers}, presented as
+Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The algorithm is
+paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
+\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
+past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
+following lemma bounds its regret.
 

### 2015-12-11 lower-bound.tex
Move and fix lower bound section

@@
 
-Our upper bounds are of the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
-\|\ell_t\|_*^2})$ where $f$ is any $1$-strongly convex function with respect to
-$\|\cdot\|$.  The same upper bound is also achieved by \textsc{FTRL} with a
-constant learning rate when the number of rounds $T$ and $\sum_{t=1}^T
-\|\ell_t\|_*^2$ is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The
-lower bound is $\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.
+The upper bounds on regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} are of
+the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2})$ where $f$
+is any $1$-strongly convex function with respect to $\|\cdot\|$.  The same
+upper bound is also achieved by \textsc{FTRL} with a constant learning rate
+when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ is known
+upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
+$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.
 

### 2015-12-11 introduction.tex
minor removals of text

@@
 learning large-scale linear models (e.g. logistic regression) can be reduced to
-OLO with $\R^d$ as the decision set. One of our contributions is an
-\textsc{FTRL}-based scale-free algorithm with a non-trivial regret bound for
-any regularizer and any decision set, bounded or unbounded.
+OLO with $\R^d$ as the decision set.
+% One of our contributions is an
+% \textsc{FTRL}-based scale-free algorithm with a non-trivial regret bound for
+% any regularizer and any decision set, bounded or unbounded.
 
@@
 
-Our proofs of the regret upper bounds rely on homogeneous inequalities
+Our proof tecnique are also new because they rely on new homogeneous inequalities
 (Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful},
-\ref{lemma:sum-of-square-roots-inverses}).  Previously, many proofs of regret
-bounds relied on ad-hoc non-homogeneous inequalities, or the existing
-algorithms used quantities (e.g.  $\sum_{t=1}^T \|\ell_t\|_*^2$) that are not
-available until all loss vectors are processed. Two of the inequalities
-(Lemmas~\ref{lemma:recurrence-solution} and \ref{lemma:useful}) are new and
-interesting in their own right.
+\ref{lemma:sum-of-square-roots-inverses}).
+% Previously, many proofs of regret
+% bounds relied on ad-hoc non-homogeneous inequalities, or the existing
+% algorithms used quantities (e.g.  $\sum_{t=1}^T \|\ell_t\|_*^2$) that are not
+% available until all loss vectors are processed. Two of the inequalities
+% (Lemmas~\ref{lemma:recurrence-solution} and \ref{lemma:useful}) are new and
+% interesting in their own right.
 

### 2015-12-13 ada-ftrl-proofs.tex
Fix typos, grammar, and bibliography.

@@
 where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
-the second inequality from the induction hypothesis and the fact that $f(x) = x
-+ \min\{a_T, a_T^2/x\}$ is an increasing function of $x$.  It remains to prove
-that
+the second inequality from the induction hypothesis and the fact that the
+function $f(x) = x + \min\{a_T, a_T^2/x\}$ is increasing on $[0,\infty)$.  It
+remains to prove that
 $$
@@
 $$
-Dividing through by $a_T$ and making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
+If $a_T = 0$, the inequality clearly holds. Dividing through by $a_T$ and
+making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
 $$

### 2015-12-13 conclusions.tex
Fix typos, grammar, and bibliography.

@@
 weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
-we gave examples of regularizers for which the associated Bregman divergence is
-unbounded and the scale-free version of \textsc{Mirror Descent} has $\Omega(T)$
-regret or worse.
+we gave examples of regularizers for which the scale-free version of
+\textsc{Mirror Descent} has $\Omega(T)$ regret or worse.
 

### 2015-12-13 introduction.tex
Fix typos, grammar, and bibliography.

--- 2015-12-11 introduction.tex
@@
 \cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
-~\cite{Rosenblatt-1958,Freund-Schapire-1999} and
+~\cite{Rosenblatt-1958, Freund-Schapire-1999} and
 regression~\cite{Kivinen-Warmuth-1997},
@@
 problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
-\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch
-and stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
+\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch and
+stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
 Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
@@
 parameter, or the parameter of strong convexity of the regularizer. Examples of
-such algorithms include the \textsc{Hedge} algorithm or online projected
-gradient descent with fixed learning rate.  However, these algorithms have two
-obvious drawbacks.
+such algorithms include the \textsc{Hedge}~\cite{Freund-Schapire-1997} or
+online projected gradient descent with fixed learning rate.  However, these
+algorithms have two obvious drawbacks.
 
@@
 
-% \begin{definition}[Scale-Free Algorithm]
-% \label{definition:scale-free-algorithm}
-% An algorithm for OLO is \emph{scale-free} if for any sequence of loss vectors
-% $\{\ell_t\}_{t=1}^\infty$ and any real number $c > 0$, on the sequences of
-% $\{\ell_t\}_{t=1}^\infty$ and $\{c \cdot \ell_t\}_{t=1}^\infty$, the algorithm
-% chooses the same sequence of points $\{w_t\}_{t=1}^\infty$.
-% \end{definition}
-
 The most important property of scale-free algorithms is that both their loss
@@
 \emph{regularizer}. The function is assumed to be strongly convex with respect
-to a given norm $\|\cdot\|$. Based on different regularizers different
-algorithms with different properties can be instantiated. A summary of
-algorithms for OLO is presented in Table~\ref{table:results}.  All of them,
-except for \textsc{Hedge}, are instances of \textsc{FTRL} or \textsc{MD}.
+to a given norm. Based on different regularizers different algorithms with
+different properties can be instantiated. A summary of algorithms for OLO is
+presented in Table~\ref{table:results}.  All of them, except for
+\textsc{Hedge}, are instances of \textsc{FTRL} or \textsc{MD}.
 
@@
 regret bounds only when the Bregman divergence associated with the regularizer
-is bounded, which happens only when the decision is bounded as well. As we show
-in Section~\ref{subsection:mirror-descent-lower-bound}, this is unavoidable,
-since \textsc{AdaGrad MD} and similar algorithms based on \textsc{MD} can have
-$\Omega(T)$ regret if Bregman divergence is not bounded.
+is bounded, which happens only when the decision set is bounded as well. As we
+show in Section~\ref{subsection:mirror-descent-lower-bound}, this is
+unavoidable, since \textsc{AdaGrad MD} and similar algorithms based on
+\textsc{MD} can have $\Omega(T)$ regret if Bregman divergence is not bounded.
 
@@
 bound was known. Unbounded decision sets are practically important, since
-learning large-scale linear models (e.g. logistic regression) can be reduced to
-OLO with $\R^d$ as the decision set. One of our contributions is an
-\textsc{FTRL}-based scale-free algorithm with a non-trivial regret bound for
-any regularizer and any decision set, bounded or unbounded.
-
-% While the scale-free property has been looked at in the expert setting, in
-% the general OLO setting this issue has been largely ignored.  In particular,
-% the \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014}
-% algorithm, for prediction with expert advice, is specifically designed to be
-% scale-free.  A notable exception in the OLO literature is the discussion of
-% the ``off-by-one'' issue in~\cite{McMahan-2014}, where it is explained that
-% even the popular \textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011} is
-% not completely adaptive; see also our discussion in
-% Section~\ref{section:solo-ftrl}. In particular, existing scale-free
-% algorithms cover only some norms/regularizers and \emph{only} bounded
-% decision sets. The case of \textbf{unbounded decision sets}, practically the
-% most interesting one for machine learning applications, remains completely
-% unsolved.
+learning large-scale linear models (e.g. logistic regression) is done by
+reducing it to OLO with decision set $\R^d$.
 
@@
 rounds is bounded by $O \left(\sqrt{\sup_{v \in K} f(v)
-\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $f$ is a non-negative regularizer that is
-$1$-strongly convex with respect to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its
-dual norm. For \textsc{Scale-Free MD}, we prove $O \left(\sqrt{\sup_{u,v \in K}
-B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $B_f$ is the Bregman
-divergence associated with a $1$-strongly convex regularizer $f$. In
+\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $f$ is a non-negative regularizer
+that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
+$\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
+\left(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} \right)$
+where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
+regularizer $f$.  As we discuss, in Section~\ref{section:mirror-descent}, the
+upper bound for \textsc{Scale-Free MD} is never better than the bound for bound
+for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In
 Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
@@
 \|\ell_t\|_* \sqrt{T} \right)$.  Same as before, $f$ is a non-negative
-$1$-strongly convex regularizer. The decision can be bounded or unbounded. This
-is the \textbf{first adaptive algorithm for unbounded decision sets} with a
-non-trivial upper bound.
+$1$-strongly convex regularizer. This bound is non-trivial for any decision
+set, bounded or unbounded.  The result makes \textsc{SOLO FTRL} the
+\textbf{first adaptive algorithm for unbounded decision sets} with a
+non-trivial regret bound.
 
@@
 
-Our proofs of the regret upper bounds rely on homogeneous inequalities
-(Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful},
-\ref{lemma:sum-of-square-roots-inverses}).  Previously, many proofs of regret
-bounds relied on ad-hoc non-homogeneous inequalities, or the existing
-algorithms used quantities (e.g.  $\sum_{t=1}^T \|\ell_t\|_*^2$) that are not
-available until all loss vectors are processed. Two of the inequalities
-(Lemmas~\ref{lemma:recurrence-solution} and \ref{lemma:useful}) are new and
-interesting in their own right.
+Our proof technique are also new because they rely on new homogeneous
+inequalities (Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful}).
 
 Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
-negative results for existing popular variants of \textsc{MD}. We
-show that variants of \textsc{MD} have $\Omega(T)$ regret if the
-Bregman divergence associated with the regularizer is unbounded.
-The setting includes the standard gradient descent with step size
-$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a
-unbounded decision set. This result suggests that \textsc{FTRL}
-is superior to \textsc{MD}.
+negative results for existing popular variants of \textsc{MD}. We show that
+variants of \textsc{MD} have $\Omega(T)$ regret if the Bregman divergence
+associated with the regularizer is unbounded.  The setting includes the
+standard gradient descent with step size $1/\sqrt{t}$, \textsc{AdaGrad MD} and
+\textsc{Scale-Free MD} on a unbounded decision set. This result suggests that
+\textsc{FTRL} is superior to \textsc{MD}.

### 2015-12-13 limits.tex
Fix typos, grammar, and bibliography.

@@
 $$
-The minimizer with smallest value of $R(w)$ is $(-1,0)$. Hence the lemma implies
-that
+The minimizer with the smallest value of $R(w)$ is $(-1,0)$. Hence the lemma
+implies that
 $$
@@
 
-\begin{proof}
+\begin{proof}[Proof of Lemma~\ref{lemma:prediction-limit-existence}]
 Without loss of generality, we can assume that $R(w)$ is non-negative for any

### 2015-12-13 mirror-descent.tex
minor removals of text
Fix typos, grammar, and bibliography.

@@
 
-\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{Follow The
-Regularized Leader} but quite different in details. The algorithm is stated as
+\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
+but quite different in details. The algorithm is stated as
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
@@
 $R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{MD}.  Proof
-can be found
+\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
 in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
@@
 an instance of \textsc{MD} with regularizer $R_t(w) =
-\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of projected gradient descent with
-step size $\eta_t = \frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized
-Infinitesimal Gradient Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}.
-Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
-Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
-\textsc{MD} for any bounded $K$ with regularizers
+\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
+Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
+Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.  Full-matrix and
+diagonal versions of \textsc{AdaGrad with Composite MD
+Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{MD} for any
+bounded $K$ with regularizers
 $$
@@
 \eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
-generalization of Zinkevich's GIGA algorithm. That is, when $\|\ell_i\|_2
+generalization of GIGA. That is, when $\|\ell_i\|_2
 = 1$ for all $i$, \textsc{MD} with regularizer
@@
 This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
-zero (and in particular for $t=0$). In this case, we define
+zero. In this case, we define
 $w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
 $w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
+Similarly, $w_1 = \argmin_{w \in K} R(w)$.
 
-The theorem below upper bounds the regret of \textsc{Scale-Free Mirror Descent}
-(see also
+The theorem below upper bounds the regret of \textsc{Scale-Free MD} (see also
 \cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).
@@
 \to \R$ is a $\lambda$-strongly convex function with respect to a norm
-$\|\cdot\|$.  \textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
+$\|\cdot\|$.  \textsc{Scale-Free MD} with regularizer $R$ satisfies
 for any $u \in K$,
@@
 that $f:K \to \R$ is a $1$-strongly convex function with respect to a norm
-$\|\cdot\|$.  \textsc{Scale-Free Mirror Descent} with regularizer
+$\|\cdot\|$.  \textsc{Scale-Free MD} with regularizer
 $$
@@
 Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
-Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
+MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
 Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
@@
 Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
-factors, the regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
-better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
-\textsc{SOLO FTRL}.  In fact, it is not hard to come up with examples where
-$R(u)$ is finite whereas $\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention
-two such examples. The first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined
-on the whole space $V$, where for any $u \in V$, $R(u)$ is a finite value but
-$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 =
-+\infty$. The second example is the negative entropy regularizer $R(w) = \ln d
-+ \sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex
-$K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u
-\in K$, $R(u)$ is finite and in fact lies in the interval $[-\ln d, 0]$ but
-$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
-\infty$. We return to these examples in the following subsection.
+factors, the regret bound for \textsc{Scale-Free MD} is inferior to the regret
+bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In fact, it is not hard to
+come up with examples where $R(u)$ is finite whereas $\sup_{v \in K}
+\Breg_R(u,v)$ is infinite. We mention two such examples. The first example is
+$R(w) = \frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u
+\in V$, $R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in
+V} \frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative
+entropy regularizer $R(w) = \ln d + \sum_{j=1}^d w_j \ln w_j$ defined on the
+$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
+\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
+lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v
+\in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples
+in the following subsection.
 
-\subsection{Lower Bounds}
+\subsection{Lower Bounds for Mirror Descent}
 \label{subsection:mirror-descent-lower-bound}
@@
 $\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
-$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free Mirror
-Descent} to have a ``small'' regret. We show necessity of this assumption on
-two counter-examples.  In these counter-examples, we consider strongly convex
+$\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free MD} to
+have a ``small'' regret. We show necessity of this assumption on two
+counter-examples.  In these counter-examples, we consider strongly convex
 regularizers $R$ such that $\Breg_R(u,v)$ is not bounded and we construct
@@
 $\|\ell_1\|_* = \|\ell_2\|_* = \dots = \|\ell_T\|_* = 1$ and \textsc{Scale-Free
-Mirror Descent} has regret $\Omega(T)$ or worse.
+MD} has regret $\Omega(T)$ or worse.
 
@@
 the first vector of the standard unit basis.  On such sequence, Zinkevich's
-\textsc{GIGA} and both versions of \textsc{AdaGrad with Composite Mirror
-Update} have predictions identical to \textsc{Scale-Free Mirror Descent}. Hence
-the lower bound applies to all four algorithms.
+\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
+identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
+algorithms.
 
@@
 $\ell_1, \ell_2, \dots, \ell_T \in V^*$ such that $\|\ell_1\|_2 = \|\ell_2\|_2
-= \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free Mirror Descent} with
-regularizer $R(w) = \frac{1}{2}\|w\|_2^2$, Zinkevich's \textsc{GIGA} algorithm,
-and both versions of \textsc{AdaGrad with Composite Mirror Descent Update}
-satisfy
+= \dots = \|\ell_T\|_2 = 1$ and \textsc{Scale-Free MD} with
+regularizer $R(w) = \frac{1}{2}\|w\|_2^2$, \textsc{GIGA},
+and both versions of \textsc{AdaGrad MD} satisfy
 $$

### 2015-12-13 preliminaries.tex
Fix typos, grammar, and bibliography.

@@
 
-Our scale-free online learning algorithms are instances of \textsc{FTRL} with
+Two of our scale-free algorithms are instances of \textsc{FTRL} with
 \emph{varying regularizers}, presented as
 Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The algorithm is
-paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
-\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
-past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
-following lemma bounds its regret.
+paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of functions $R_t:K \to \R$
+called \emph{regularizers}.  Each regularizer $R_t$ can depend on the past loss
+vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The following
+lemma bounds its regret.
 
@@
 \label{lemma:generic-regret-bound}
-For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
-semi-continuous regularizers, the regret of
-Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
-bounded as
+If the regularizers $R_1, R_2, \dots$ chosen by
+Algorithm~\ref{algorithm:ftrl-varying-regularizer} are strongly convex and lower
+semi-continuous, the algorithm's regret is upper bounded as
 $$

### 2015-12-13 ada-ftrl-proofs.tex
Fix README, capitalize titles in bibfile, etc.

@@
 $$
-If $a_T = 0$, the inequality clearly holds. Dividing through by $a_T$ and
-making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
+If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
+divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
+a_t^2}}{a_T}$, which leads to
 $$

### 2015-12-14 conclusions.tex
Mention Pistol and fixes

@@
 regret and do not need to know or assume \textbf{anything} about the sequence
-of loss vectors, i.e. without any assumption on the norms of the loss vectors
-and the number of rounds. We have designed two scale-free algorithms for online
-linear optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v)
+of loss vectors. In particular, the algorithms do not assume any upper or lower
+bounds on the norms of the loss vectors or the number of rounds.
+
+We have designed two scale-free algorithms based on \textsc{Follow The
+Regularizer Leader}.  Their regret is $O \left(\sqrt{\sup_{v \in K} f(v)
 \sum_{t=1}^T \|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly
 convex function with respect to a norm $\|\cdot\|$ defined on the decision set
-and where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
-based on \textsc{Follow The Regularizer Leader}.  We have shown that one of the
-algorithms has regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
-\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T} \right)$ with respect to any
-competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex
-function defined on the decision set.  The latter result makes sense even when
-the decision set $K$ is unbounded.
+and where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.
 
@@
 lower bound on regret of any algorithm for decision set with diameter $D$.
+
+We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
+regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
+\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
+$f$ is any non-negative $1$-strongly convex function defined on the decision
+set.  The latter result makes sense even when the decision set $K$ is
+unbounded. Notice that with regularizer $f(w) = \frac{1}{2}\|w\|_2^2$ the
+regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
+exist non-scale-free algorithms \cite{ McMahan-Streeter-2012,
+McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
+have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
+dependency.  These algorithms, however, assume an a priori bound on the norm of
+the loss vectors.  This leads to another open problem:
+%
+\begin{quotation}
+\noindent
+\emph{Design a scale-free algorithm that has $O(\|u\|_2 \sqrt{\log \|u\|_2})$
+dependency on the norm $\|u\|_2$ of the competitor.}
+\end{quotation}

### 2015-12-14 introduction.tex
Mention Pistol and fixes

@@
 Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
-forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
+forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
 developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
@@
 bound was known. Unbounded decision sets are practically important, since
-learning large-scale linear models (e.g. logistic regression) is done by
-reducing it to OLO with decision set $\R^d$.
+learning of large-scale linear models (e.g. logistic regression) is done by
+gradient methods that can be reduced to OLO with decision set $\R^d$.
 
@@
 
-Our proof technique are also new because they rely on new homogeneous
-inequalities (Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful}).
+Our proof techniques rely on new homogeneous
+inequalities (Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful})
+which might be of independent interest.
 
 Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
-negative results for existing popular variants of \textsc{MD}. We show that
-variants of \textsc{MD} have $\Omega(T)$ regret if the Bregman divergence
-associated with the regularizer is unbounded.  The setting includes the
-standard gradient descent with step size $1/\sqrt{t}$, \textsc{AdaGrad MD} and
-\textsc{Scale-Free MD} on a unbounded decision set. This result suggests that
-\textsc{FTRL} is superior to \textsc{MD}.
+negative results for existing popular variants of \textsc{MD}. We show two
+examples that demonstrate variants of \textsc{MD} can have $\Omega(T)$ regret
+(on sequences of loss vector of unit norm) if the Bregman divergence associated
+with the regularizer is unbounded.  First example demonstrates
+$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
+$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
+decision set.  Second example demonstrates $\Omega(T)$ regret for
+\textsc{Scale-Free MD} for learning with expert advice with negative entropy
+regularizer.  Both results indicate that \textsc{FTRL} is superior to
+\textsc{MD}.

### 2015-12-15 introduction.tex
Fix grammar

@@
 negative results for existing popular variants of \textsc{MD}. We show two
-examples that demonstrate variants of \textsc{MD} can have $\Omega(T)$ regret
-(on sequences of loss vector of unit norm) if the Bregman divergence associated
-with the regularizer is unbounded.  First example demonstrates
+examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
+regret (on sequences of loss vectors of unit norm) if the Bregman divergence
+associated with the regularizer is unbounded.  First example demonstrates
 $\Omega(T^{3/2})$ regret of the standard gradient descent with step size

### 2015-12-16 ada-ftrl-proofs.tex
Improve wording

@@
 $$
-where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
-the second inequality from the induction hypothesis and the fact that the
+where the second inequality from the induction hypothesis and the fact that the
 function $f(x) = x + \min\{a_T, a_T^2/x\}$ is increasing on $[0,\infty)$.  It

### 2015-12-16 mirror-descent.tex
Mention Pistol and fixes
Improve wording

@@
 Replacing $\frac{1}{2}\|w\|_2^2$ in equation
-\eqref{equation:ada-grad-regularizer} with an arbitrary convex function $R:K
-\to \R$ yields
+\eqref{equation:ada-grad-regularizer} with an arbitrary strongly convex
+function $R:K \to \R$ yields
 \begin{equation}
@@
 The theorem below upper bounds the regret of \textsc{Scale-Free MD} (see also
-\cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).
-Proof is in~\ref{section:mirror-descent-proofs}.
+\cite{Duchi-Hazan-Singer-2011, Duchi-Shalev-Shwartz-Singer-Tewari-2010,
+Rakhlin-Sridharan-2013}).  Proof is in~\ref{section:mirror-descent-proofs}.
 
@@
 
-\subsection{Lower Bounds for Mirror Descent}
+\subsection{Lower Bounds for Scale-Free Mirror Descent}
 \label{subsection:mirror-descent-lower-bound}
@@
 The first counter-example is stated as
-Theorem~\ref{theorem:first-counter-example} below. The decision set is the
-whole space $K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note
-that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
-norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
+Theorem~\ref{theorem:first-counter-example} below; proof is
+in~\ref{section:mirror-descent-proofs}. The decision set is the whole space
+$K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note that $R(w)$ is
+$1$-strongly convex with respect to $\|\cdot\|_2$ and the dual norm of
+$\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
 $\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
@@
 The second counter-example is stated as
-Theorem~\ref{theorem:second-counter-example} below.  The decision set is the
+Theorem~\ref{theorem:second-counter-example} below; proof is
+in~\ref{section:mirror-descent-proofs}.  The decision set is the
 $d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,

### 2015-12-21 abstract.tex
Polished grammar and typography.

@@
 and at the same time do not need to know any upper or lower bounds on the norm
-of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors by scale
-invariance, i.e., our algorithms make exactly the same decisions if the
-sequence of loss vectors is multiplied by any positive constant.  Our
-algorithms work for any decision set, bounded or unbounded.  For unbounded
-decisions sets, these are the first adaptive algorithms for online linear
+of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors
+by scale invariance, i.e., our algorithms make exactly the same decisions if
+the sequence of loss vectors is multiplied by any positive constant.  One of
+our algorithms works for any decision set, bounded or unbounded.  For unbounded
+decisions sets, it is the first adaptive algorithm for online linear
 optimization with a non-vacuous bound.

### 2015-12-21 ada-ftrl.tex
typos and small changes
Polished grammar and typography.

@@
 
-Assume that $K$ is bounded and that $R(w)$ is a strongly convex lower
+Assume that $K$ is bounded and that $R:K \to \R$ is a strongly convex lower
 semi-continuous function bounded from above.  We instantiate
@@
 \label{theorem:ada-ftrl-regret-bound}
-Suppose $K \subseteq V$ is a non-empty bounded closed convex subset. Let $D =
+Suppose $K \subseteq V$ is a non-empty bounded closed convex set. Let $D =
 \sup_{x,y \in K} \|x - y\|$ be its diameter with respect to a norm $\|\cdot\|$.
@@
 semi-continuous function that is $\lambda$-strongly convex with respect to
-$\|\cdot\|$ and is bounded from above.  The regret of \textsc{AdaFTRL} satisfies
+$\|\cdot\|$ and is bounded from above.  The regret of \textsc{AdaFTRL}
+satisfies
 $$
@@
 \label{corollary:ada-ftrl-regret-bound}
-Suppose $K \subseteq V$ is a non-empty bounded closed convex subset. Suppose
-$f:K \to \R$ is a non-negative lower semi-continuous function that is
-$1$-strongly convex with respect to $\|\cdot\|$ and is bounded from above.  The
-regret of \textsc{AdaFTRL} with regularizer
+Suppose $K \subseteq V$ is a non-empty bounded closed convex set. Suppose $f:K
+\to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
+convex with respect to $\|\cdot\|$ and is bounded from above.  The regret of
+\textsc{AdaFTRL} with regularizer
 $$
@@
 
-\begin{lemma}[Recurrence]
+\begin{lemma}[Solution of the recurrence]
 \label{lemma:recurrence-solution}

### 2015-12-21 introduction.tex
Polished grammar and typography.

@@
 algorithm.  That is, the sequence of decisions of the algorithm does not change
-if the sequence of loss vectors is multiplied by a positive constant.
-
-The most important property of scale-free algorithms is that both their loss
-and their regret is scales linearly with the maximum norm of the loss vector
-appearing in the sequence.
+if the sequence of loss vectors is multiplied by a positive constant.  The most
+important property of scale-free algorithms is that both their loss and their
+regret scale linearly with the maximum norm of the loss vector appearing in the
+sequence.
 
@@
 for convex optimization. In the online learning literature it makes the first
-appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. Good
+appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. A good
 description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

### 2015-12-21 limits.tex
Polished grammar and typography.

@@
 \end{align*}
-The limits on the right exist because of compactness of $K$. They are simply
-the minimizers $u= \argmin_{w \in K} - \langle x, w \rangle$ and $v= \argmin_{w
-\in K} - \langle y, w \rangle$ where ties in $\argmin$ are broken according to
-smaller value of $R(w)$.
+The limits on the right exist according to
+Lemma~\ref{lemma:prediction-limit-existence}.  They are simply the minimizers
+$u = \argmin_{w \in K} - \langle x, w \rangle$ and $v= \argmin_{w \in K} -
+\langle y, w \rangle$ where ties in $\argmin$ are broken according to smaller
+value of $R(w)$.
 

### 2015-12-21 mirror-descent.tex
Polished grammar and typography.

@@
 $$
-w_{t+1} = \Pi_K(w_t - \eta_t \ell_t)
+w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right)
 $$
@@
 \label{theorem:regret-scale-free-mirror-descent}
-Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose that $R:K
+Suppose $K \subseteq V$ is a non-empty closed convex set. Suppose that $R:K
 \to \R$ is a $\lambda$-strongly convex function with respect to a norm
@@
 \label{corollary:regret-scale-free-mirror-descent}
-Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
+Suppose $K \subseteq V$ is a non-empty bounded closed convex set.  Suppose
 that $f:K \to \R$ is a $1$-strongly convex function with respect to a norm
@@
 sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
-the first vector of the standard unit basis.  On such sequence, Zinkevich's
-\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
-identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
-algorithms.
+the first vector of the standard orthnormal basis.  On such sequence,
+Zinkevich's \textsc{GIGA} and both versions of \textsc{AdaGrad MD} have
+predictions identical to \textsc{Scale-Free MD}. Hence the lower bound applies
+to all four algorithms.
 

### 2015-12-21 solo-ftrl.tex
Reword and reorganize some sentences
Polished grammar and typography.

@@
 The closest algorithm to a scale-free one in the OLO literature is the
-\textsc{AdaGrad} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
+\textsc{AdaGrad FTRL} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
 regularizer on each coordinate of the form
@@
 the ``doubling'' power. Hence, even in the case one would prove a regret bound,
-%\footnote{For
-%lack of space, we cannot include the regret bound for the doubling trick
-%version. It would be exactly the same as in
-%Theorem~\ref{theorem:regret-solo-ftrl}, following a similar analysis, but with
-%the additional parameter of the doubling power.},
-such strategy would give the
-feeling that \textsc{FTRL} needs to be ``fixed'' in order to obtain a
-scale-free algorithm.
+such strategy would give the feeling that \textsc{FTRL} needs to be ``fixed''
+in order to obtain a scale-free algorithm.
 
@@
 \label{theorem:regret-solo-ftrl}
-Suppose $K \subseteq V$ is a non-empty closed convex subset.  Let $D =
-\sup_{u,v \in K} \|u - v\|$ be its diameter with respect to a norm $\|\cdot\|$.
-Suppose that the regularizer $R:K \to \R$ is a non-negative lower
-semi-continuous function that is $\lambda$-strongly convex with respect to
-$\|\cdot\|$. The regret of \textsc{SOLO FTRL} satisfies
+Suppose $K \subseteq V$ is a non-empty closed convex set.  Let $D = \sup_{u,v
+\in K} \|u - v\|$ be its diameter with respect to a norm $\|\cdot\|$.  Suppose
+that the regularizer $R:K \to \R$ is a non-negative lower semi-continuous
+function that is $\lambda$-strongly convex with respect to $\|\cdot\|$. The
+regret of \textsc{SOLO FTRL} satisfies
 \begin{align*}
@@
 \label{corollary:regret-solo-ftrl-bounded-set}
-Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
-that $f:K \to \R$ is a non-negative lower semi-continuous function that is
+Suppose $K \subseteq V$ is a non-empty bounded closed convex set.  Suppose that
+$f:K \to \R$ is a non-negative lower semi-continuous function that is
 $1$-strongly convex with respect to $\|\cdot\|$. \textsc{SOLO FTRL} with
@@
 The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
-(Lemma~\ref{lemma:useful}).  Related and weaker inequalities were proved by
-~\cite{Auer-Cesa-Bianchi-Gentile-2002} and ~\cite{Jaksch-Ortner-Auer-2010}.
+(Lemma~\ref{lemma:useful}).  Related and weaker inequalities were proved
+in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}.
 The main property of this inequality is that on the right-hand side $C$ does
@@
 \sum_{t=1}^T \min \left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, \ C a_t \right\}
-\le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
+\le 3.5 \, C \max_{t=1,2,\dots,T} a_t \ + \ 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
 $$
@@
 \begin{multline*}
-\min\left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, C a_t \right\}
+\min\left\{ \frac{a_t^2}{\sqrt{\sum_{i=1}^{t-1} a_i^2}}, \ C a_t \right\}
 \le C a_t

### 2015-12-23 conclusions.tex
Improve wording
Improve wording and fix typos

@@
 set.  The latter result makes sense even when the decision set $K$ is
-unbounded. Notice that with regularizer $f(w) = \frac{1}{2}\|w\|_2^2$ the
+unbounded. Notice that with regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
 regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
-exist non-scale-free algorithms \cite{ McMahan-Streeter-2012,
+exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
 McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
 have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
-dependency.  These algorithms, however, assume an a priori bound on the norm of
-the loss vectors.  This leads to another open problem:
+dependency.  These algorithms assume an a priori bound on the norm of the loss
+vectors.  This leads to another open problem:
 %
@@
 \noindent
-\emph{Design a scale-free algorithm that has $O(\|u\|_2 \sqrt{\log \|u\|_2})$
-dependency on the norm $\|u\|_2$ of the competitor.}
+\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
+which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
 \end{quotation}

### 2015-12-23 introduction.tex
Minor fixes
Improve wording and fix typos

@@
 parameter, or the parameter of strong convexity of the regularizer. Examples of
-such algorithms include the \textsc{Hedge}~\cite{Freund-Schapire-1997} or
-online projected gradient descent with fixed learning rate.  However, these
-algorithms have two obvious drawbacks.
+such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
+projected gradient descent with fixed learning rate.  However, these algorithms
+have two obvious drawbacks.
 
 First, they do not come with any regret guarantee for sequences of loss vectors
-with norms exceeding $B$. Second, on sequences where the norm of loss vectors
-is bounded by $b \ll B$, these algorithms fail to have an optimal regret
-guarantee that depends on $b$ rather than on $B$.
+with norms exceeding $B$. Second, on sequences of loss vectors with norms
+bounded by $b \ll B$, these algorithms fail to have an optimal regret guarantee
+that depends on $b$ rather than on $B$.
 
@@
 There is a clear practical need to design algorithms that adapt automatically
-to norms of the loss vectors.  A natural, yet overlooked, design method to
+to the norms of the loss vectors.  A natural, yet overlooked, design method to
 achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
@@
 different properties can be instantiated. A summary of algorithms for OLO is
-presented in Table~\ref{table:results}.  All of them, except for
-\textsc{Hedge}, are instances of \textsc{FTRL} or \textsc{MD}.
+presented in Table~\ref{table:results}.  All of them are instances of
+\textsc{FTRL} or \textsc{MD}.
 
@@
 regret bounds only when the Bregman divergence associated with the regularizer
-is bounded, which happens only when the decision set is bounded as well. As we
-show in Section~\ref{subsection:mirror-descent-lower-bound}, this is
-unavoidable, since \textsc{AdaGrad MD} and similar algorithms based on
-\textsc{MD} can have $\Omega(T)$ regret if Bregman divergence is not bounded.
+is bounded. In particular, since a bound on the Bregman divergence implies that
+decision set is bounded, the regret bounds are vacuous for unbounded sets.  In
+fact, as we show in Section~\ref{subsection:mirror-descent-lower-bound}
+\textsc{AdaGrad MD} and similar algorithms based on \textsc{MD} have
+$\Omega(T)$ regret if the Bregman divergence is not bounded.
 

### 2015-12-23 mirror-descent.tex
Minor fixes
Improve wording and fix typos

@@
 
-Several existing algorithms are special cases of \textsc{MD}.
-For example, \textsc{Exponentially-Weighted Average
-Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with a constant learning
-rate $\eta$ for learning with expert advice is an instance of
-\textsc{MD} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
-\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
-\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
-algorithm is \emph{not} scale-free.
+Several existing algorithms are special cases of \textsc{MD}.  For example,
+\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
+\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
+regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
+is a (constant) learning rate. Note that this algorithm is \emph{not}
+scale-free.
 
@@
 
-The regret bound for \textsc{Scale-Free Mirror Descent} in the
+The regret bound for \textsc{Scale-Free MD} in the
 Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
@@
 
-Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
-factors, the regret bound for \textsc{Scale-Free MD} is inferior to the regret
-bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In fact, it is not hard to
-come up with examples where $R(u)$ is finite whereas $\sup_{v \in K}
-\Breg_R(u,v)$ is infinite. We mention two such examples. The first example is
-$R(w) = \frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u
-\in V$, $R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in
-V} \frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative
-entropy regularizer $R(w) = \ln d + \sum_{j=1}^d w_j \ln w_j$ defined on the
-$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
-\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
-lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v
-\in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples
-in the following subsection.
+Thus, inequality \eqref{equation:regularizer-vs-divergence} implies
+that---ignoring constant factors---the regret bound for \textsc{Scale-Free MD}
+is inferior to the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.
+In fact, it is not hard to come up with examples where $R(u)$ is finite whereas
+$\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention two such examples. The
+first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined on the whole space $V$,
+where for any $u \in V$, $R(u)$ is a finite value but $\sup_{v \in K}
+\Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 = +\infty$. The second
+example is the shifted negative entropy regularizer $R(w) = \ln(d) +
+\sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex $K
+= \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u \in
+K$, $R(u)$ is finite and in fact lies in the interval $[0, \ln d]$ but
+$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
+\infty$. We revisit these examples in the following subsection.
 
@@
 $\Breg_R(u,v)$ is bounded is necessary in order for \textsc{Scale-Free MD} to
-have a ``small'' regret. We show necessity of this assumption on two
+have a sublinear regret. We show necessity of this assumption on two
 counter-examples.  In these counter-examples, we consider strongly convex
@@
 the first vector of the standard orthnormal basis.  On such sequence,
-Zinkevich's \textsc{GIGA} and both versions of \textsc{AdaGrad MD} have
-predictions identical to \textsc{Scale-Free MD}. Hence the lower bound applies
-to all four algorithms.
+\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
+identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
+algorithms.
 
@@
 \ell_T \in V^*$ such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots =
-\|\ell_T\|_\infty = 1$ and \textsc{Scale-Free Mirror Descent} with regularizer
+\|\ell_T\|_\infty = 1$ and \textsc{Scale-Free MD} with regularizer
 $R(w) = \sum_{j=1}^d w_j \ln w_j$ satisfies

### 2015-12-28 introduction.tex
Minor fixes

@@
 is bounded. In particular, since a bound on the Bregman divergence implies that
-decision set is bounded, the regret bounds are vacuous for unbounded sets.  In
-fact, as we show in Section~\ref{subsection:mirror-descent-lower-bound}
-\textsc{AdaGrad MD} and similar algorithms based on \textsc{MD} have
-$\Omega(T)$ regret if the Bregman divergence is not bounded.
+the decision set is bounded, the regret bounds for \textsc{AdaGrad MD} and
+\textsc{Optimistic MD} are vacuous for unbounded sets. In fact, as we show in
+Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
+similar algorithms based on \textsc{MD} have $\Omega(T)$ regret if the Bregman
+divergence is not bounded.
 
@@
 For unbounded decision sets no scale-free algorithm with a non-trivial regret
-bound was known. Unbounded decision sets are practically important, since
-learning of large-scale linear models (e.g. logistic regression) is done by
-gradient methods that can be reduced to OLO with decision set $\R^d$.
+bound was known. Unbounded decision sets are practically important (see e.g.
+\cite{Mcmahan-Holt-Sculley-2013}), since learning of large-scale linear models
+(e.g. logistic regression) is done by gradient methods that can be reduced to
+OLO with decision set $\R^d$.
 
@@
 Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
-\|\ell_t\|_*^2}$ term in the bounds term is necessary by proving
-$\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
-regret of any algorithm for OLO for any decision set with diameter $D$ with
-respect to the primal norm $\|\cdot\|$.
+\|\ell_t\|_*^2}$ term in the bounds is necessary by proving $\frac{D}{\sqrt{8}}
+\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
+for OLO for any decision set with diameter $D$ with respect to the primal norm
+$\|\cdot\|$.
 
@@
 All three algorithms are \textbf{any-time}, i.e., they do not need to know the
-number of rounds in advance and the regret bounds hold for all time steps
+number of rounds, $T$, in advance and the regret bounds hold for all $T$
 simultaneously.

### 2015-12-28 preliminaries.tex
Polished grammar and typography.
Minor fixes

@@
 Sometimes we will assume that $K$ is also bounded. We denote by $D$ its
-diameter with respect to $\|\cdot\|$, i.e. $D = \sup_{u,v \in K} \|u - v\|$. If
-$K$ is unbounded, $D = +\infty$.
+diameter with respect to $\|\cdot\|$, i.e., $D = \sup_{u,v \in K} \|u - v\|$.
+If $K$ is unbounded, $D = +\infty$.
 
@@
 $$
-\text{$\forall w \in K$, $\forall \ell \in V^*$} \qquad
+\forall w \in K, \ \forall \ell \in V^* \qquad \qquad
 f(w) + f^*(\ell) \ge \langle \ell, w \rangle \; .
@@
 \label{proposition:diameter-vs-range}
-Let $K \subseteq V$ be a non-empty bounded closed convex subset.  Let $D =
+Let $K \subseteq V$ be a non-empty bounded closed convex set.  Let $D =
 \sup_{u,v \in K} \|u - v\|$ be its diameter with respect to $\|\cdot\|$.  Let
@@
 
-\item $f^*$ is $\frac{1}{\lambda}$-strongly smooth i.e. for any $x,y \in V^*$, $\Breg_{f^*}(x, y) \le \frac{1}{2\lambda} \|x - y\|_*^2$.
+\item $f^*$ is $\frac{1}{\lambda}$-strongly smooth, i.e., for any $x,y \in
+V^*$, $\Breg_{f^*}(x, y) \le \frac{1}{2\lambda} \|x - y\|_*^2$.
 
-\item $f^*$ has $\frac{1}{\lambda}$-Lipschitz continuous gradients i.e.
+\item $f^*$ has $\frac{1}{\lambda}$-Lipschitz continuous gradients, i.e.,
 for any $x,y \in V^*$,

### 2016-01-04 conclusions.tex
Minor fixes
minor changes to the text

@@
 We have investigated scale-free algorithms for online linear optimization and
-we have shown that scale-free property leads to algorithms which have optimal
+we have shown that the scale-free property leads to algorithms which have optimal
 regret and do not need to know or assume \textbf{anything} about the sequence
@@
 
-We have shown an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
+We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
 lower bound on regret of any algorithm for decision set with diameter $D$.
@@
 set.  The latter result makes sense even when the decision set $K$ is
-unbounded. Notice that with regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
-regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
+unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
+regret depends quadratically on norm of the competitor $\|u\|_2$. There
 exist non-scale-free algorithms \cite{McMahan-Streeter-2012,

### 2016-01-04 mirror-descent.tex
Minor fixes
minor changes to the text

@@
 u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
-$\grad R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient
+$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
 of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
@@
 
-\begin{lemma}[Regret of Mirror Descent]
+\begin{lemma}[Regret of \textsc{MD}]
 \label{lemma:mirror-descent-regret}
@@
 $$
-w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right)
+w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right),
 $$
@@
 \label{equation:regularizer-vs-divergence}
-\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
+\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v),
 \end{equation}

### 2016-01-04 preliminaries.tex
minor changes to the text

@@
 Monotonicity of Fenchel conjugates follows easily from the definition: If
-$f,g:K \to \R$ satisfy $f(w) \le g(w)$ for all $w \in K$ then $f^*(\ell) \ge
+$f,g:K \to \R$ satisfies $f(w) \le g(w)$ for all $w \in K$ then $f^*(\ell) \ge
 g^*(\ell)$ for every $\ell \in V^*$.
@@
 
-\begin{lemma}[Regret of Follow The Regularized Leader]
+\begin{lemma}[Regret of \textsc{FTRL}]
 \label{lemma:generic-regret-bound}

### 2016-01-05 mirror-descent.tex
Very minor fixes

@@
 $$
-w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right),
+w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
 $$
@@
 \label{equation:regularizer-vs-divergence}
-\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v),
+\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v) \; ,
 \end{equation}

### 2016-09-15 revision/ada-ftrl.tex
new directory for revision and first draft

--- 2015-12-21 ada-ftrl.tex
@@
 
-In this section we generalize the \textsc{AdaHedge}
+Here we show that it is possible to derive scale-free algorithms in an alternative way. We generalize the \textsc{AdaHedge}
 algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
@@
 $f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
-optimization is the following corollary.  Its proof can be found
-in~\ref{section:ada-ftrl-proofs}.
+optimization is the following corollary.
 
@@
 \end{corollary}
+%
+\begin{proof}
+Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:ada-ftrl-regret-bound}
+applied to the regularizer $R(w) = \frac{c}{S} f(w)$ and
+Proposition~\ref{proposition:diameter-vs-range} gives
+$$
+\Regret_T \le \sqrt{3}(1 + c) \max\left\{\sqrt{8}, \frac{1}{\sqrt{2c}} \right\} \sqrt{S \sum_{t=1}^T \|\ell_t\|_*^2} \; .
+$$
+It remains to find the minimum of $g(c) = \sqrt{3}(1 + c) \max\{\sqrt{8},
+1/\sqrt{2c}\}$.  The function $g$ is strictly convex on $(0, \infty)$ and has
+minimum at $c=1/16$ and $g(\frac{1}{16}) = \sqrt{3}(1+\frac{1}{16})\sqrt{8} \le
+5.3$.
+\end{proof}
 
@@
 
-\begin{lemma}[Solution of the recurrence]
+\begin{lemma}[Solution of the Recurrence]
 \label{lemma:recurrence-solution}
 Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
-Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of non-negative real
-numbers satisfying
+Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of real numbers
+satisfying
 $$
@@
 %
-Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred
-to~\ref{section:ada-ftrl-proofs}.  Theorem~\ref{theorem:ada-ftrl-regret-bound}
+\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
+We prove the lemma by induction on $T$.  The base case $T=0$ is trivial.
+For $T \ge 1$, let $f(x) = x + \min\{a_T, a_T^2/x\}$. We have
+\begin{align*}
+\Delta_T
+& \le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\} \\
+& = f \left( \Delta_{T-1} \right) \\
+& \le f \left( \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \right) \\
+& = \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}  \; ,
+\end{align*}
+where the second inequality follows from the inductive hypothesis $\Delta_{T-1} \le
+\sqrt{3 \sum_{t=1}^{T-1} a_t^2}$ and that $f(x)$ is non-positive on $[-\infty,
+0)$ and non-negative increasing on $[0,\infty)$. It remains to prove that
+$$
+\sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
+\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
+$$
+If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
+divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
+a_t^2}}{a_T}$, which leads to
+$$
+z\sqrt{3} + \min\left\{1, \frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
+$$
+which can be easily checked by considering separately the cases $z \in
+[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
+\end{proof}
+
+Theorem~\ref{theorem:ada-ftrl-regret-bound}
 follows from

### 2016-09-15 revision/introduction.tex
new directory for revision and first draft

--- 2015-12-28 introduction.tex
@@
 input $B$ is often disguised as the learning rate, the regularization
-parameter, or the parameter of strong convexity of the regularizer. Examples of
-such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
-projected gradient descent with fixed learning rate.  However, these algorithms
-have two obvious drawbacks.
+parameter, or the parameter of strong convexity of the regularizer.
+%Examples of
+%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
+%projected gradient descent with fixed learning rate.
+However, these algorithms have two obvious drawbacks.
 
@@
 \textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
-\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
+\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func.\footnotemark & \textbf{Yes} \\ \hline
 \textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
@@
 \textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
-\textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
 \textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \le C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
 \textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func.
-\newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
-\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
+\newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label2}} \\ \hline
+%\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
+\textsc{Scale-free MD} \textbf{[folklore]} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
 \textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
@@
 
-\footnotetext{\label{footnote-label} These algorithms attempt to produce an
+\addtocounter{footnote}{-2}
+
+\stepcounter{footnote}\footnotetext{\label{footnote-label1} Even if, in principle the FTRL-Proximal algorithm can be used with any proximal regularizer, to the best of our knowledge a general way to construct proximal regularizers is not known.
+The only proximal regularizer we are aware is based on the 2-norm.}
+
+\stepcounter{footnote}\footnotetext{\label{footnote-label2} These algorithms attempt to produce an
 invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a
@@
 
-Majority of existing algorithms for OLO are based on two generic algorithms:
+The majority of the existing algorithms for OLO are based on two generic algorithms:
 \textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
@@
 Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
-convex optimization under the name \textsc{Dual Averaging} and later
-rediscovered in~\cite{Xiao-2010} for online convex optimization. The general
+convex optimization under the name \textsc{Dual Averaging} and rediscovered
+in~\cite{Xiao-2010} for online convex optimization. The general
 version with time-varying regularizers was analyzed in
@@
 in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
-for convex optimization. In the online learning literature it makes the first
-appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. A good
-description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.
+for convex optimization. In the online learning literature it makes his first
+appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.
+%A good description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.
 
-Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called a
+Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called 
 \emph{regularizer}. The function is assumed to be strongly convex with respect
@@
 Scale-free versions of \textsc{MD} include \textsc{AdaGrad
-MD}~\cite{Duchi-Hazan-Singer-2011} and~\textsc{Optimistic
-MD}~\cite{Rakhlin-Sridharan-2013}. However, these algorithms have non-trivial
-regret bounds only when the Bregman divergence associated with the regularizer
+MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
+MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
 is bounded. In particular, since a bound on the Bregman divergence implies that
-the decision set is bounded, the regret bounds for \textsc{AdaGrad MD} and
-\textsc{Optimistic MD} are vacuous for unbounded sets. In fact, as we show in
+the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
+is vacuous for unbounded sets. In fact, as we show in
 Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
@@
 learning with expert advice, where the decision set is bounded.
-Algorithms based on \textsc{FTRL} that are ``almost'' scale-free include
-\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} and
-\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  These algorithms fail to
-be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and also our
+An algorithm based on \textsc{FTRL} that is ``almost'' scale-free is 
+\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to
+be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and the
 discussion in Section~\ref{section:solo-ftrl}.
+Instead, \textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the off-by-one issues, but it requires proximal regularizers, that in general do not have a simple form and even in the simple 2-norm case requires bounded domains.
 
@@
 strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
-``correct'' scale-free version of \textsc{AdaGrad
+``correct'' scale-free version of the diagonal version of \textsc{AdaGrad
 FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
@@
 generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
-arbitrary strongly convex regularizers, similar to \textsc{Optimistic
-MD}~\cite{Rakhlin-Sridharan-2013}.  The three algorithms are presented and
-analyzed in Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
+arbitrary strongly convex regularizers.  The three algorithms are presented in
+Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
 \ref{section:mirror-descent} respectively.
@@
 We prove that the regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} after $T$
-rounds is bounded by $O \left(\sqrt{\sup_{v \in K} f(v)
-\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $f$ is a non-negative regularizer
+rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v)
+\sum_{t=1}^T\|\ell_t\|_*^2} )$ where $f$ is a non-negative regularizer
 that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
 $\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
-\left(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} \right)$
+(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
 where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
@@
 For \textsc{SOLO FTRL}, we prove that regret against a competitor $u \in K$ is
-at most $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
-\|\ell_t\|_* \sqrt{T} \right)$.  Same as before, $f$ is a non-negative
+at most $O (f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
+\|\ell_t\|_* \sqrt{T} )$.  Same as before, $f$ is a non-negative
 $1$-strongly convex regularizer. This bound is non-trivial for any decision
@@
 regret (on sequences of loss vectors of unit norm) if the Bregman divergence
-associated with the regularizer is unbounded.  First example demonstrates
-$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
-$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
-decision set.  Second example demonstrates $\Omega(T)$ regret for
-\textsc{Scale-Free MD} for learning with expert advice with negative entropy
-regularizer.  Both results indicate that \textsc{FTRL} is superior to
-\textsc{MD}.
+associated with the regularizer is unbounded.
+%First example demonstrates
+%$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
+%$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
+%decision set.  Second example demonstrates $\Omega(T)$ regret for
+%\textsc{Scale-Free MD} for learning with expert advice with negative entropy
+%regularizer.
+These results indicate that \textsc{FTRL} is superior to \textsc{MD}.

### 2016-09-15 revision/lower-bound.tex
new directory for revision and first draft

--- 2015-12-11 lower-bound.tex
@@
 We show a lower bound on the worst-case regret of any algorithm for OLO. The
-proof is a standard probabilistic argument, which we present
-in~\ref{section:lower-bound-proof}.
+proof, presented in~\ref{section:lower-bound-proof}, is a standard
+probabilistic argument.
 
@@
 
-The upper bounds on regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} are of
+The upper bound on regret of \textsc{SOLO FTRL} is of
 the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2})$ where $f$

### 2016-09-15 revision/mirror-descent.tex
new directory for revision and first draft

--- 2016-01-05 mirror-descent.tex
@@
 
-\begin{algorithm}[t]
-\caption{\textsc{Mirror Descent with Varying Regularizer}}
-\label{algorithm:mirror-descent-varying-regularizer}
-\begin{algorithmic}[1]
-\REQUIRE Non-empty closed convex set $K \subseteq V$
-\STATE Choose a regularizer $R_0:K \to \R$
-\STATE $w_1 \leftarrow \argmin_{w \in K} R_0(w)$
-\FOR{$t=1,2,3,\dots$}
-\STATE Predict $w_t$
-\STATE Observe $\ell_t \in V^*$
-\STATE Choose a regularizer $R_t:K \to \R$
-\STATE $w_{t+1} \leftarrow \argmin_{w \in K} \left( \langle \ell, w \rangle + \Breg_{R_t}(w, w_t) \right)$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
-but quite different in details. The algorithm is stated as
-Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
-parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
-\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
-loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
-not differentiable\footnote{Note that this can happen even when $R_t$ is a
-restriction of a differentiable function defined on a superset of $K$.  If $K$
-is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
-If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
-Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
-u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
-$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
-of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
-$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
-in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
-completeness we give a proof in~\ref{section:mirror-descent-proofs}.
-
-\begin{lemma}[Regret of \textsc{MD}]
-\label{lemma:mirror-descent-regret}
-Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
-$u \in K$,
-$$
-\Regret_T(u)
-\le
-\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
-$$
-\end{lemma}
-
-Several existing algorithms are special cases of \textsc{MD}.  For example,
-\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
-\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
-regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
-is a (constant) learning rate. Note that this algorithm is \emph{not}
-scale-free.
-
-The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
-Descent} defined by the recurrence
-$$
-w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
-$$
-where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
-$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
-an instance of \textsc{MD} with regularizer $R_t(w) =
-\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
-Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
-Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.  Full-matrix and
-diagonal versions of \textsc{AdaGrad with Composite MD
+Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
 Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{MD} for any
@@
 Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
-$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
-\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
-Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
+$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bound for
+\textsc{SOLO FTRL} in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}
 depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
@@
 MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
-Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
-for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
+Theorem~\ref{theorem:regret-solo-ftrl}
+for \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
 show that
@@
 that---ignoring constant factors---the regret bound for \textsc{Scale-Free MD}
-is inferior to the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.
+is inferior to the regret bound for \textsc{SOLO FTRL}.
 In fact, it is not hard to come up with examples where $R(u)$ is finite whereas

### 2016-09-15 revision/preliminaries.tex
new directory for revision and first draft

--- 2015-12-28 preliminaries.tex
@@
 For completeness, we include it in~\ref{section:definitions-proofs}.
+
+\subsection{Generic Mirror Descent with Varying Regularizer}
+
+\begin{algorithm}[t]
+\caption{\textsc{Mirror Descent with Varying Regularizer}}
+\label{algorithm:mirror-descent-varying-regularizer}
+\begin{algorithmic}[1]
+\REQUIRE Non-empty closed convex set $K \subseteq V$
+\STATE Choose a regularizer $R_0:K \to \R$
+\STATE $w_1 \leftarrow \argmin_{w \in K} R_0(w)$
+\FOR{$t=1,2,3,\dots$}
+\STATE Predict $w_t$
+\STATE Observe $\ell_t \in V^*$
+\STATE Choose a regularizer $R_t:K \to \R$
+\STATE $w_{t+1} \leftarrow \argmin_{w \in K} \left( \langle \ell, w \rangle + \Breg_{R_t}(w, w_t) \right)$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
+but quite different in details. The algorithm is stated as
+Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
+parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
+\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
+loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
+not differentiable\footnote{Note that this can happen even when $R_t$ is a
+restriction of a differentiable function defined on a superset of $K$.  If $K$
+is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
+If $K$ is a subset of an affine subspace of a dimension smaller than the
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
+Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
+u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
+$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
+of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
+$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
+\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
+in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
+completeness we give a proof in~\ref{section:mirror-descent-proofs}.
+
+\begin{lemma}[Regret of \textsc{MD}]
+\label{lemma:mirror-descent-regret}
+Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
+$u \in K$,
+$$
+\Regret_T(u)
+\le
+\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
+$$
+\end{lemma}
+
+Several existing algorithms are special cases of \textsc{MD}.  For example,
+\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
+\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
+regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
+is a (constant) learning rate. Note that this algorithm is \emph{not}
+scale-free.
+
+The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
+Descent} defined by the recurrence
+$$
+w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
+$$
+where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
+$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
+an instance of \textsc{MD} with regularizer $R_t(w) =
+\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
+Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
+Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.

### 2016-09-15 revision/solo-ftrl.tex
new directory for revision and first draft

--- 2015-12-21 solo-ftrl.tex
@@
 
-The closest algorithm to a scale-free one in the OLO literature is the
-\textsc{AdaGrad FTRL} algorithm~\cite{Duchi-Hazan-Singer-2011}.  It uses a
+The closest algorithm to a scale-free \textsc{FTRL} in the existing literature
+is the \textsc{AdaGrad FTRL} algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a
 regularizer on each coordinate of the form
@@
 Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is
-in~\ref{section:solo-ftrl-proofs}.  It is similar to
-Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}. The
-scaling however is different in the two corollaries.  In
-Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in
-K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we
-have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.
+in~\ref{section:solo-ftrl-proofs}.
+%It is similar to Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}.
+The scaling however is different in the two corollaries.
+In Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.
 
@@
 \begin{align*}
-& \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \\
+\Breg_{R_t^*}&(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \\
 & = \Breg_{R_{t+1}^*}(-L_t, -L_{t-1}) + R^*_{t+1}(-L_{t-1}) - R_t^*(-L_{t-1}) \\

### 2016-09-15 revision/abstract.tex
changes for reviewer 1

--- 2015-12-09 abstract.tex
@@
 and at the same time do not need to know any upper or lower bounds on the norm
-of the loss vectors.  We achieve adaptiveness to the norms of the loss vectors by scale
-invariance, i.e., our algorithms make exactly the same decisions if the
-sequence of loss vectors is multiplied by any positive constant.  Our
-algorithms work for any decision set, bounded or unbounded.  For unbounded
-decisions sets, these are the first adaptive algorithms for online linear
-optimization with a non-vacuous bound.
-
-We also study a popular scale-free variant of online mirror descent algorithm,
-and we show that in two natural settings it has linear or worse regret.
+of the loss vectors. We design them as instantiation of the Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms.
+We achieve adaptiveness to the norms of the loss vectors
+by scale invariance, i.e., our algorithms make exactly the same decisions if
+the sequence of loss vectors is multiplied by any positive constant.
+The algorithm based on FTRL works for any decision set, bounded or unbounded.
+For unbounded decisions sets, this is the first adaptive algorithm for online linear
+optimization with a non-vacuous regret bound.
+We also show lower bounds on scale-free algorithms based on MD on unbounded domains.

### 2016-09-15 revision/introduction.tex
changes for reviewer 1

@@
 %\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
-\textsc{Scale-free MD} \textbf{[folklore]} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
+\textsc{Scale-free MD} \textbf{[this paper]} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
 \textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
@@
 achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
-algorithm.  That is, the sequence of decisions of the algorithm does not change
+algorithm.  That is, with the same parameters, the sequence of decisions of the algorithm does not change
 if the sequence of loss vectors is multiplied by a positive constant.  The most
@@
 convex optimization under the name \textsc{Dual Averaging} and rediscovered
-in~\cite{Xiao-2010} for online convex optimization. The general
-version with time-varying regularizers was analyzed in
-\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
+in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers were analyzed
+in~\cite{Duchi-Hazan-Singer-2011} and generalized in \cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
 in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
@@
 
-We design and analyze three scale-free algorithms: \textsc{AdaFTRL},
-\textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  \textsc{AdaFTRL} and
+We design and analyze two scale-free algorithms: \textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  A third one, \textsc{AdaFTRL}, is present in the Appendix~\ref{section:ada-ftrl}. \textsc{AdaFTRL} and
 \textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a

### 2016-09-15 revision/solo-ftrl-proofs.tex
changes for reviewer 1

--- 2015-11-23 solo-ftrl-proofs.tex
@@
-\section{Proofs for SOLO FTRL}
+\section{Proofs for \textsc{SOLO FTRL}}
 \label{section:solo-ftrl-proofs}
@@
 We use the inequality $x/\sqrt{x+y} \le 2(\sqrt{x+y} - \sqrt{y})$ which holds
-for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{s=1}^{t-1} a_s$,
+for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{i=1}^{t-1} a_i$,
 we get that for any $t \ge 1$,
 $$
-\frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{s=1}^t a_s} \ - \ 2 \sqrt{\sum_{s=1}^{t-1} a_s} \; .
+\frac{a_t}{\sqrt{\sum_{i=1}^t a_i}} \le 2 \sqrt{\sum_{i=1}^t a_i} \ - \ 2 \sqrt{\sum_{i=1}^{t-1} a_i} \; .
 $$
@@
 \end{proof}
-
-\begin{proof}[Proof of Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}]
-Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:regret-solo-ftrl} applied
-to the regularizer $R(w) = \frac{c}{\sqrt{S}} f(w)$, together with
-Proposition~\ref{proposition:diameter-vs-range} and a crude bound
-$\max_{t=1,2,\dots,T} \|\ell_t\|_* \le \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$,
-give
-$$
-\Regret_T \le \left(c + \frac{2.75}{c}  + 3.5\sqrt{8} \right) \sqrt{S \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
-$$
-We choose $c$ by minimizing $g(c) = c + \frac{2.75}{c} + 3.5\sqrt{8}$. Clearly,
-$g(c)$ has minimum at $c = \sqrt{2.75}$ and has minimal value $g(\sqrt{2.75}) =
-2\sqrt{2.75} + 3.5\sqrt{8} \le 13.3$.
-\end{proof}

### 2016-09-15 revision/solo-ftrl.tex
changes for reviewer 1

@@
 \label{section:solo-ftrl}
+
+In this section, we introduces our first scale-free algorithm, based on \textsc{FTRL}.
 
@@
 
-When $K$ is bounded, we can choose the optimal multiple of the regularizer.  We
+When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T} \|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the bound.
+On the other hand, when $K$ is bounded, the second term is a constant and we can choose the optimal multiple of the regularizer.  We
 choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
 optimize $\lambda$.  The result of the optimization is
-Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is
-in~\ref{section:solo-ftrl-proofs}.
+Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}.
+%; the proof is in~\ref{section:solo-ftrl-proofs}.
 %It is similar to Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}.
@@
 \end{corollary}
+%
+\begin{proof}
+Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:regret-solo-ftrl} applied
+to the regularizer $R(w) = \frac{c}{\sqrt{S}} f(w)$, together with
+Proposition~\ref{proposition:diameter-vs-range} and a crude bound
+$\max_{t=1,2,\dots,T} \|\ell_t\|_* \le \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$,
+give
+$$
+\Regret_T \le \left(c + \frac{2.75}{c}  + 3.5\sqrt{8} \right) \sqrt{S \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
+$$
+We choose $c$ by minimizing $g(c) = c + \frac{2.75}{c} + 3.5\sqrt{8}$. Clearly,
+$g(c)$ has minimum at $c = \sqrt{2.75}$ and has minimal value $g(\sqrt{2.75}) =
+2\sqrt{2.75} + 3.5\sqrt{8} \le 13.3$.
+\end{proof}
 
@@
 
-\begin{lemma}[Simple Inequality]
+\begin{lemma}[{{\cite[Lemma~3.5]{Auer-Cesa-Bianchi-Gentile-2002}}}]
 \label{lemma:sum-of-square-roots-inverses}
@@
 %
-Lemma~\ref{lemma:sum-of-square-roots-inverses} is from
-\cite[Lemma~3.5]{Auer-Cesa-Bianchi-Gentile-2002}.  We give a proof
-in~\ref{section:solo-ftrl-proofs} for completeness.
+For completeness, a proof of Lemma~\ref{lemma:sum-of-square-roots-inverses} is in
+in~\ref{section:solo-ftrl-proofs}.
 

### 2016-09-15 revision/conclusions.tex
typos of reviewer 1

--- 2016-01-04 conclusions.tex
@@
 McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
-have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
+have only $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
 dependency.  These algorithms assume an a priori bound on the norm of the loss
@@
 \noindent
-\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
+\emph{Is there a scale-free algorithm for the decision set $\R^d$, the regret of
 which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}

### 2016-09-15 revision/introduction.tex
typos of reviewer 1

@@
 Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
-reduced to OLO, e.g. online convex
+reduced to OLO, e.g., online convex
 optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
@@
 regression~\cite{Kivinen-Warmuth-1997},
-~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
+~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed bandits
 problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
@@
 For this reason, most of the existing OLO algorithms receive as an input---or
-explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
+implicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
 input $B$ is often disguised as the learning rate, the regularization
@@
 Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
-similar algorithms based on \textsc{MD} have $\Omega(T)$ regret if the Bregman
+similar algorithms based on \textsc{MD} might incur $\Omega(T)$ regret if the Bregman
 divergence is not bounded.
@@
 For unbounded decision sets no scale-free algorithm with a non-trivial regret
-bound was known. Unbounded decision sets are practically important (see e.g.
+bound was known. Unbounded decision sets are practically important (see, e.g.,
 \cite{Mcmahan-Holt-Sculley-2013}), since learning of large-scale linear models
-(e.g. logistic regression) is done by gradient methods that can be reduced to
+(e.g., logistic regression) is done by gradient methods that can be reduced to
 OLO with decision set $\R^d$.

### 2016-09-15 revision/preliminaries.tex
typos of reviewer 1

@@
 loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
-not differentiable\footnote{Note that this can happen even when $R_t$ is a
+not differentiable,\footnote{Note that this can happen even when $R_t$ is a
 restriction of a differentiable function defined on a superset of $K$.  If $K$
@@
 If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
+dimension of $V$, then $R_t$ fails to be differentiable everywhere.} the
 Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),

### 2016-09-15 revision/solo-ftrl.tex
typos of reviewer 1

@@
 use Algorithm~\ref{algorithm:ftrl-varying-regularizer} with the regularizer
-\begin{equation*}
+\begin{equation}
 \label{equation:solo-ftrl-regularizer}
 R_t(w) = R(w) \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \; ,
-\end{equation*}
+\end{equation}
 where $R:K \to \R$ is any strongly convex function. Through a refined analysis,
@@
 Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
-result is the Theorem~\ref{theorem:regret-solo-ftrl} below, which is proven in
+result is Theorem~\ref{theorem:regret-solo-ftrl} below, which is proven in
 Section~\ref{section:solo-ftrl-regret-bound}.

### 2016-09-15 revision/ada-ftrl.tex
typos of reviewer 2 and his simplified proof

@@
 Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on
-the regret of the algorithm after $T$ rounds, without the need to know before
-hand an upper bound on $\|\ell_t\|_*$.  We prove the theorem in
+the regret of the algorithm after $T$ rounds, without the need to know beforehand an upper bound on $\|\ell_t\|_*$.  We prove the theorem in
 Section~\ref{section:ada-ftrl-regret-bound}.
@@
 \begin{proof}
-Let $R_t(w) = \Delta_{t-1} R(w)$. Since $R$ is non-negative,
+Recall from \eqref{equation:ada-ftrl} that $R_t(w) = \Delta_{t-1} R(w)$. Since $R$ is non-negative,
 $\{R_t\}_{t=1}^\infty$ is non-decreasing.  Hence, $R_t^*(\ell) \ge
@@
 can upper bound $\Breg_{R^*}\left( - \frac{L_t}{\Delta_{t-1}}, -
-\frac{L_{t-1}}{\Delta_{t-1}} \right)$ with two different quantities.  Taking
+\frac{L_{t-1}}{\Delta_{t-1}} \right)$ with two different quantities.  Taking the
 minimum of the two quantities finishes the proof.
@@
 %
-\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
-We prove the lemma by induction on $T$.  The base case $T=0$ is trivial.
-For $T \ge 1$, let $f(x) = x + \min\{a_T, a_T^2/x\}$. We have
-\begin{align*}
-\Delta_T
-& \le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\} \\
-& = f \left( \Delta_{T-1} \right) \\
-& \le f \left( \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \right) \\
-& = \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}  \; ,
-\end{align*}
-where the second inequality follows from the inductive hypothesis $\Delta_{T-1} \le
-\sqrt{3 \sum_{t=1}^{T-1} a_t^2}$ and that $f(x)$ is non-positive on $[-\infty,
-0)$ and non-negative increasing on $[0,\infty)$. It remains to prove that
-$$
-\sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
-\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
-$$
-If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
-divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
-a_t^2}}{a_T}$, which leads to
-$$
-z\sqrt{3} + \min\left\{1, \frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
-$$
-which can be easily checked by considering separately the cases $z \in
-[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
+\begin{proof}
+Observe that
+\[
+\Delta_T^2
+=
+\sum_{t=1}^T (\Delta_t - \Delta_{t-1})(\Delta_t - \Delta_{t-1} + 2\Delta_{t-1}) \; .
+\]
+The left term of the minimum inequality in the definition of $\Delta_t$ gives
+\[
+(\Delta_t - \Delta_{t-1})^2 \le a_t^2,
+\]
+while the right term gives
+\[
+2 (\Delta_t - \Delta_{t-1}) \Delta_{t-1} \le 2 a_t^2 \; .
+\]
+So, we conclude
+\[
+\Delta_T^2
+\le
+3 \sum_{t=1}^T a_t^2\; .
+\]
 \end{proof}

### 2016-09-15 revision/conclusions.tex
typos of reviewer 2 and his simplified proof

@@
 
-Similar, but weaker result holds for a scale-free algorithm based on
-\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
+A similar, but weaker result holds for a scale-free algorithm based on
+\textsc{Mirror Descent}. However, we have also shown this algorithm to be strictly
 weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
@@
 We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
-lower bound on regret of any algorithm for decision set with diameter $D$.
+lower bound on the regret of any algorithm for any decision set with diameter $D$.
 
-We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
+We have shown that the scale-free \textsc{FTRL}-based algorithm \textsc{SOLO FTRL} has
 regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
@@
 unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
-regret depends quadratically on norm of the competitor $\|u\|_2$. There
+regret depends quadratically on the norm of the competitor $\|u\|_2$. There
 exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
 McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
-have only $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
+have only a $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
 dependency.  These algorithms assume an a priori bound on the norm of the loss

### 2016-09-15 revision/introduction.tex
typos of reviewer 2 and his simplified proof

@@
 in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
-for convex optimization. In the online learning literature it makes his first
+for convex optimization. In the online learning literature it makes its first
 appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.
@@
 \textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to
-be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and the
+be scale-free due to ``off-by-one'' issues; see~\cite{McMahan-2014} and the
 discussion in Section~\ref{section:solo-ftrl}.
@@
 
-We prove that the regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} after $T$
+We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
 rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v)
@@
 where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
-regularizer $f$.  As we discuss, in Section~\ref{section:mirror-descent}, the
-upper bound for \textsc{Scale-Free MD} is never better than the bound for bound
-for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In
-Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
-\|\ell_t\|_*^2}$ term in the bounds is necessary by proving $\frac{D}{\sqrt{8}}
+regularizer $f$.  
+%However, as we discuss in Section~\ref{section:mirror-descent}, the
+%upper bound for \textsc{Scale-Free MD} is never better than the bound
+%for \textsc{SOLO FTRL} and \textsc{AdaFTRL}.
+In Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
+\|\ell_t\|_*^2}$ term in the bounds is necessary by proving a $\frac{D}{\sqrt{8}}
 \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
@@
 
-For \textsc{SOLO FTRL}, we prove that regret against a competitor $u \in K$ is
+For \textsc{SOLO FTRL}, we prove that the regret against a competitor $u \in K$ is
 at most $O (f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
-\|\ell_t\|_* \sqrt{T} )$.  Same as before, $f$ is a non-negative
+\|\ell_t\|_* \sqrt{T} )$.  As before, $f$ is a non-negative
 $1$-strongly convex regularizer. This bound is non-trivial for any decision

### 2016-09-15 revision/lower-bound.tex
typos of reviewer 2 and his simplified proof

@@
 The upper bounds on the regret, which we have proved for our algorithms, have
-the same dependency on the norms of loss vectors.  However, a gap remains
+the same dependency on the norms of the loss vectors.  However, a gap remains
 between the lower bound and the upper bounds.
@@
 upper bound is also achieved by \textsc{FTRL} with a constant learning rate
-when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ is known
+when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ are known
 upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is

### 2016-09-15 revision/mirror-descent.tex
typos of reviewer 2 and his simplified proof

@@
 Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
-Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{MD} for any
+Update}~\cite{Duchi-Hazan-Singer-2011} are instances of \textsc{MD} for any
 bounded $K$ with regularizers
@@
 respectively. The diagonal version is simply the per-coordinate construction
-applied to the one-dimensional version. When $d=1$, both algorithms are
+applied to the one-dimensional version, see Section~\ref{sec:per-coordinate}. When $d=1$, both algorithms are
 equivalent to
@@
 \cite{Duchi-Hazan-Singer-2011, Duchi-Shalev-Shwartz-Singer-Tewari-2010,
-Rakhlin-Sridharan-2013}).  Proof is in~\ref{section:mirror-descent-proofs}.
+Rakhlin-Sridharan-2013}).  The proof is in~\ref{section:mirror-descent-proofs}.
 
@@
 The first counter-example is stated as
-Theorem~\ref{theorem:first-counter-example} below; proof is
+Theorem~\ref{theorem:first-counter-example} below; our proof is
 in~\ref{section:mirror-descent-proofs}. The decision set is the whole space
@@
 $\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
-$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
+$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs a
 sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
-the first vector of the standard orthnormal basis.  On such sequence,
+the first vector of the standard orthnormal basis.  On such a sequence,
 \textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions

### 2016-09-15 revision/per-coordinate.tex
typos of reviewer 2 and his simplified proof

--- 2015-12-09 per-coordinate.tex
@@
 \subsection{Per-Coordinate Learning}
+\label{sec:per-coordinate}
 
 An interesting class of algorithms proposed in~\cite{McMahan-Streeter-2010} and
-\cite{Duchi-Hazan-Singer-2011} are based on the so-called per-coordinate
-learning rates.  As shown in \cite{Streeter-McMahan-2010}, any algorithm for OLO can be used with per-coordinate learning rates as well.
+\cite{Duchi-Hazan-Singer-2011} are based on so-called per-coordinate
+learning rates.  As shown in \cite{Streeter-McMahan-2010}, any algorithm for
+OLO can be used with per-coordinate learning rates as well.
 
@@
 If the algorithm for each factor is scale-free, the overall algorithm is
-clearly scale-free as well. Hence, even if not explicitly mentioned in the text, any algorithm we present can be trivially transformed in a per-coordinate version.
-
-%Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
-%each factor $K_j$, we generalize and improve existing regret bounds
-%\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with
-%per-coordinate learning rates.
+clearly scale-free as well. Hence, even if not explicitly mentioned in the
+text, any algorithm we present can be trivially transformed to a per-coordinate
+version.

### 2016-09-15 revision/preliminaries-proofs.tex
typos of reviewer 2 and his simplified proof

--- 2015-12-08 preliminaries-proofs.tex
@@
 $v^*$ is guaranteed to exist by lower semi-continuity of $f$ and compactness of
-$K$.  Optimality condition for $v^*$ and $1$-strong convexity of $f$ imply that
+$K$.  The optimality condition for $v^*$ and $1$-strong convexity of $f$ imply that
 for any $u \in K$,
@@
 $$
-In other words, $\|u - v^*\| \le \sqrt{2S}$. By triangle inequality,
+In other words, $\|u - v^*\| \le \sqrt{2S}$. By the triangle inequality,
 $$
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:generic-regret-bound}]
-By Fenchel-Young inequality,
+By the Fenchel-Young inequality,
 \begin{align*}

### 2016-09-15 revision/preliminaries.tex
typos of reviewer 2 and his simplified proof

@@
 \textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
-but quite different in details. The algorithm is stated as
+but quite different in the details. The algorithm is stated as
 Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
@@
 $\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
-of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
+of $R_t$ at any point $w$. If $R_t$ is a restriction of a differentiable function
 $R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
+\in K$. The following lemma bounds the regret of \textsc{MD}.  A proof can be found
 in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For

### 2016-09-15 revision/solo-ftrl.tex
typos of reviewer 2 and his simplified proof

@@
 where $R:K \to \R$ is any strongly convex function. Through a refined analysis,
-we show that the regularizer suffices to obtain an optimal regret bound for any
-decision set, bounded or unbounded.  We call such variant \textsc{Scale-free
+we show that this regularizer suffices to obtain an optimal regret bound for any
+decision set, bounded or unbounded.  We call this variant \textsc{Scale-free
 Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
@@
 \ell_2, \dots, \ell_{t-1}$ are all zero (and in particular for $t=1$).  In that
-case, we define the $w_t = \argmin_{w \in K} R(w)$ which is consistent with
+case, we define $w_t = \argmin_{w \in K} R(w)$ which is consistent with
 $w_t = \lim_{a \to 0^+}  \argmin_{w \in K} a R(w)$.
@@
 $\|\ell_t\|_* > 0$ for all $t$, since otherwise we can remove all rounds $t$
-where $\ell_t = 0$ without affecting regret and the predictions of the
+where $\ell_t = 0$ without affecting the regret and the predictions of the
 algorithm on the remaining rounds.  By Lemma~\ref{lemma:generic-regret-bound},

### 2016-09-27 revision/abstract.tex
Improvements

@@
-We design algorithms for online linear optimization that have optimal regret
-and at the same time do not need to know any upper or lower bounds on the norm
-of the loss vectors. We design them as instantiation of the Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms.
-We achieve adaptiveness to the norms of the loss vectors
-by scale invariance, i.e., our algorithms make exactly the same decisions if
-the sequence of loss vectors is multiplied by any positive constant.
-The algorithm based on FTRL works for any decision set, bounded or unbounded.
-For unbounded decisions sets, this is the first adaptive algorithm for online linear
-optimization with a non-vacuous regret bound.
-We also show lower bounds on scale-free algorithms based on MD on unbounded domains.
+We design and analyze algorithms for online linear optimization that have
+optimal regret and at the same time do not need to know any upper or lower
+bounds on the norm of the loss vectors. Our algorithms are instances of the
+Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms. We
+achieve adaptiveness to the norms of the loss vectors by scale invariance, i.e.,
+our algorithms make exactly the same decisions if the sequence of loss vectors
+is multiplied by any positive constant.  The algorithm based on FTRL works for
+any decision set, bounded or unbounded.  For unbounded decisions sets, this is
+the first adaptive algorithm for online linear optimization with a non-vacuous
+regret bound. In contrast, we show lower bounds on scale-free algorithms based
+on MD on unbounded domains.

### 2016-09-27 revision/ada-ftrl.tex
Improvements

@@
 
-Here we show that it is possible to derive scale-free algorithms in an alternative way. We generalize the \textsc{AdaHedge}
+In this section, we show that it is possible to derive a scale-free algorithm
+different from \textsc{SOLO FTRL}. We generalize the \textsc{AdaHedge}
 algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
-showing that it retains its scale-free property. The analysis is very general
-and based on general properties of strongly convex functions, rather than
-specific properties of the entropic regularizer like in \textsc{AdaHedge}.
+showing that it retains its scale-free property.  We call the resulting
+algorithm \textsc{AdaFTRL}. The analysis is very general and based on general
+properties of strongly convex functions, rather than specific properties of the
+entropic regularizer as in the original analysis of \textsc{AdaHedge}.
 
@@
 
-Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on
-the regret of the algorithm after $T$ rounds, without the need to know beforehand an upper bound on $\|\ell_t\|_*$.  We prove the theorem in
-Section~\ref{section:ada-ftrl-regret-bound}.
+Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on the
+regret of the algorithm after $T$ rounds, without the need to know beforehand an
+upper bound on $\|\ell_t\|_*$.  We prove the theorem
+in~\ref{section:ada-ftrl-regret-bound}.
 
@@
 Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
-Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of real numbers
-satisfying
+Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of non-negative
+real numbers satisfying
 $$
@@
 \Delta_T^2
-=
-\sum_{t=1}^T (\Delta_t - \Delta_{t-1})(\Delta_t - \Delta_{t-1} + 2\Delta_{t-1}) \; .
+= \sum_{t=1}^T \Delta_t^2 - \Delta_{t-1}^2
+= \sum_{t=1}^T (\Delta_t - \Delta_{t-1})^2 + 2 (\Delta_t - \Delta_{t-1}) \Delta_{t-1} \; .
 \]
+We now bound each term in the sum.
 The left term of the minimum inequality in the definition of $\Delta_t$ gives

### 2016-09-27 revision/introduction.tex
Improvements

@@
 e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
-Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
-online combinatorial optimization~\cite{Kalai-Vempala-2005,
+Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997}
+and online combinatorial optimization~\cite{Kalai-Vempala-2005,
 Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
@@
 parameter, or the parameter of strong convexity of the regularizer.
-%Examples of
-%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
-%projected gradient descent with fixed learning rate.
 However, these algorithms have two obvious drawbacks.
@@
 
-\stepcounter{footnote}\footnotetext{\label{footnote-label1} Even if, in principle the FTRL-Proximal algorithm can be used with any proximal regularizer, to the best of our knowledge a general way to construct proximal regularizers is not known.
-The only proximal regularizer we are aware is based on the 2-norm.}
+\stepcounter{footnote}\footnotetext{\label{footnote-label1} Even if, in
+principle the FTRL-Proximal algorithm can be used with any proximal regularizer,
+to the best of our knowledge a general way to construct proximal regularizers is
+not known. The only proximal regularizer we are aware is based on the 2-norm.}
 
-\stepcounter{footnote}\footnotetext{\label{footnote-label2} These algorithms attempt to produce an
-invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a
-sequence of invariant $w_t$.}
+\stepcounter{footnote}\footnotetext{\label{footnote-label2} These algorithms
+attempt to produce an invariant sequence of predictions $\langle w_t, \ell_t
+\rangle$, rather than a sequence of invariant $w_t$.}
 
-There is a clear practical need to design algorithms that adapt automatically
-to the norms of the loss vectors.  A natural, yet overlooked, design method to
+There is a clear practical need to design algorithms that adapt automatically to
+the norms of the loss vectors.  A natural, yet overlooked, design method to
 achieve this type of adaptivity is by insisting to have a \textbf{scale-free}
-algorithm.  That is, with the same parameters, the sequence of decisions of the algorithm does not change
-if the sequence of loss vectors is multiplied by a positive constant.  The most
-important property of scale-free algorithms is that both their loss and their
-regret scale linearly with the maximum norm of the loss vector appearing in the
-sequence.
+algorithm.  That is, with the same parameters, the sequence of decisions of the
+algorithm does not change if the sequence of loss vectors is multiplied by a
+positive constant.  The most important property of scale-free algorithms is that
+both their loss and their regret scale linearly with the maximum norm of the
+loss vector appearing in the sequence.
 
@@
 
-The majority of the existing algorithms for OLO are based on two generic algorithms:
-\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
-Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
-forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
-developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
-Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
+The majority of the existing algorithms for OLO are based on two generic
+algorithms: \textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and
+\textsc{Mirror Descent} (\textsc{MD}). \textsc{FTRL} dates back to the
+potential-based forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and
+its theory was developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow
+The Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
 Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
 convex optimization under the name \textsc{Dual Averaging} and rediscovered
-in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers were analyzed
-in~\cite{Duchi-Hazan-Singer-2011} and generalized in \cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
-in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
-for convex optimization. In the online learning literature it makes its first
+in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers
+were analyzed in~\cite{Duchi-Hazan-Singer-2011} and generalized in
+\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed in
+\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for
+convex optimization. In the online learning literature it makes its first
 appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.
-%A good description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.
 
@@
 Scale-free versions of \textsc{MD} include \textsc{AdaGrad
-MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
-MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
-is bounded. In particular, since a bound on the Bregman divergence implies that
-the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
-is vacuous for unbounded sets. In fact, as we show in
+MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad MD algorithm has a
+non-trivial regret bounds only when the Bregman divergence associated with the
+regularizer is bounded. In particular, since a bound on the Bregman divergence
+implies that the decision set is bounded, the regret bound for \textsc{AdaGrad
+MD} and is vacuous for unbounded sets. In fact, as we show in
 Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
-similar algorithms based on \textsc{MD} might incur $\Omega(T)$ regret if the Bregman
-divergence is not bounded.
+similar algorithms based on \textsc{MD} incurs $\Omega(T)$ regret, in the worst
+case, if the Bregman divergence is not bounded.
 
@@
 \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
-learning with expert advice, where the decision set is bounded.
-An algorithm based on \textsc{FTRL} that is ``almost'' scale-free is 
-\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to
-be scale-free due to ``off-by-one'' issues; see~\cite{McMahan-2014} and the
-discussion in Section~\ref{section:solo-ftrl}.
-Instead, \textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the off-by-one issues, but it requires proximal regularizers, that in general do not have a simple form and even in the simple 2-norm case requires bounded domains.
+learning with expert advice, where the decision set is bounded. An algorithm
+based on \textsc{FTRL} that is ``almost'' scale-free is \textsc{AdaGrad
+FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to be scale-free due
+to ``off-by-one'' issue; see~\cite{McMahan-2014} and the discussion in
+Section~\ref{section:solo-ftrl}. Instead,
+\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the
+off-by-one issue, but it requires proximal regularizers, that in general do not
+have a simple form and even in the simple 2-norm case requires bounded domains.
 
@@
 
-We design and analyze two scale-free algorithms: \textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  A third one, \textsc{AdaFTRL}, is present in the Appendix~\ref{section:ada-ftrl}. \textsc{AdaFTRL} and
-\textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a
-generalization of
+We design and analyze two scale-free algorithms: \textsc{SOLO FTRL} and
+\textsc{Scale-Free MD}.  A third one, \textsc{AdaFTRL}, is presented in
+the Appendix. \textsc{SOLO FTRL} and \textsc{AdaFTRL} are based on
+\textsc{FTRL}.  \textsc{AdaFTRL} is a generalization of
 \textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
@@
 arbitrary strongly convex regularizers.  The three algorithms are presented in
-Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
-\ref{section:mirror-descent} respectively.
+Sections~\ref{section:solo-ftrl} and \ref{section:mirror-descent}, and
+\ref{section:ada-ftrl}, respectively.
 
 We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
-rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v)
-\sum_{t=1}^T\|\ell_t\|_*^2} )$ where $f$ is a non-negative regularizer
-that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
-$\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
-(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
+rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
+where $f$ is a non-negative regularizer that is $1$-strongly convex with respect
+to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free
+MD}, we prove $O (\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
 where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
-regularizer $f$.  
-%However, as we discuss in Section~\ref{section:mirror-descent}, the
-%upper bound for \textsc{Scale-Free MD} is never better than the bound
-%for \textsc{SOLO FTRL} and \textsc{AdaFTRL}.
-In Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
-\|\ell_t\|_*^2}$ term in the bounds is necessary by proving a $\frac{D}{\sqrt{8}}
-\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
-for OLO for any decision set with diameter $D$ with respect to the primal norm
-$\|\cdot\|$.
+regularizer $f$. In Section~\ref{section:lower-bound}, we show that the
+$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term in the bounds is necessary by proving
+a $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
+regret of any algorithm for OLO for any decision set with diameter $D$ with
+respect to the primal norm $\|\cdot\|$.
 
-For \textsc{SOLO FTRL}, we prove that the regret against a competitor $u \in K$ is
-at most $O (f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
-\|\ell_t\|_* \sqrt{T} )$.  As before, $f$ is a non-negative
-$1$-strongly convex regularizer. This bound is non-trivial for any decision
-set, bounded or unbounded.  The result makes \textsc{SOLO FTRL} the
-\textbf{first adaptive algorithm for unbounded decision sets} with a
-non-trivial regret bound.
+For \textsc{SOLO FTRL}, we prove that the regret against a competitor $u \in K$
+is at most $O (f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
+\|\ell_t\|_* \sqrt{T} )$.  As before, $f$ is a non-negative $1$-strongly convex
+regularizer. This bound is non-trivial for any decision set, bounded or
+unbounded.  The result makes \textsc{SOLO FTRL} the \textbf{first adaptive
+algorithm for unbounded decision sets} with a non-trivial regret bound.
 
@@
 regret (on sequences of loss vectors of unit norm) if the Bregman divergence
-associated with the regularizer is unbounded.
-%First example demonstrates
-%$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
-%$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
-%decision set.  Second example demonstrates $\Omega(T)$ regret for
-%\textsc{Scale-Free MD} for learning with expert advice with negative entropy
-%regularizer.
-These results indicate that \textsc{FTRL} is superior to \textsc{MD}.
+associated with the regularizer is unbounded. These results indicate that
+\textsc{FTRL} is superior to \textsc{MD}.

### 2016-09-27 revision/mirror-descent.tex
Improvements

@@
 
-Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
-Update}~\cite{Duchi-Hazan-Singer-2011} are instances of \textsc{MD} for any
-bounded $K$ with regularizers
-$$
-R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
-\qquad \text{and} \qquad
-R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
-$$
-respectively. The diagonal version is simply the per-coordinate construction
-applied to the one-dimensional version, see Section~\ref{sec:per-coordinate}. When $d=1$, both algorithms are
-equivalent to
-\begin{equation}
-\label{equation:ada-grad-regularizer}
-R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
-\end{equation}
-All three sequences of regularizers give a scale-free algorithm. Regularizer
-\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
-generalization of GIGA. That is, when $\|\ell_i\|_2
-= 1$ for all $i$, \textsc{MD} with regularizer
-\eqref{equation:ada-grad-regularizer} coincides with GIGA.
-
-Replacing $\frac{1}{2}\|w\|_2^2$ in equation
-\eqref{equation:ada-grad-regularizer} with an arbitrary strongly convex
-function $R:K \to \R$ yields
+In this section, we analyze scale-free version of \textsc{Mirror Descent}.
+Our algorithm uses regularizer
 \begin{equation}
@@
 \end{equation}
-We call the resulting algorithm \textsc{Scale-Free MD} with
-regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
+where $R:K \to \R$ an arbitrary strongly convex function.
+We call the resulting algorithm \textsc{Scale-Free MD}.
+Similar to \textsc{SOLO FTRL}, the regularizer
 \eqref{equation:scale-free-mirror-descent} does not uniquely define
@@
 Similarly, $w_1 = \argmin_{w \in K} R(w)$.
+
+Per-coordinate version of \textsc{Scale-Free MD} with regularizer $R(w) =
+\frac{1}{2}\norm{w}_2^2$ is exactly the same as the diagonal version of
+\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011}.
 

### 2016-09-27 revision/preliminaries.tex
Improvements

@@
 $R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds the regret of \textsc{MD}.  A proof can be found
-in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
-completeness we give a proof in~\ref{section:mirror-descent-proofs}.
+\in K$. The following lemma bounds the regret of \textsc{MD}.
 
@@
 
-Several existing algorithms are special cases of \textsc{MD}.  For example,
-\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
-\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
-regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
-is a (constant) learning rate. Note that this algorithm is \emph{not}
-scale-free.
+The proof of the lemma can be found
+in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
+completeness, we give a proof in~\ref{section:mirror-descent-proofs}.
 
-The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
-Descent} defined by the recurrence
-$$
-w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
-$$
-where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
-$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
-an instance of \textsc{MD} with regularizer $R_t(w) =
-\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
-Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
-Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.
+%Several existing algorithms are special cases of \textsc{MD}.  For example,
+%\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
+%\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
+%regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
+%is a (constant) learning rate. Note that this algorithm is \emph{not}
+%scale-free.
+
+%The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
+%Descent} defined by the recurrence
+%$$
+%w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
+%$$
+%where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
+%$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
+%an instance of \textsc{MD} with regularizer $R_t(w) =
+%\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
+%Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
+%Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.

### 2016-09-27 revision/solo-ftrl.tex
Improvements

@@
 
-In this section, we introduces our first scale-free algorithm, based on \textsc{FTRL}.
-
-The closest algorithm to a scale-free \textsc{FTRL} in the existing literature
-is the \textsc{AdaGrad FTRL} algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a
-regularizer on each coordinate of the form
+In this section, we introduce our first scale-free algorithm; it will be based
+on \textsc{FTRL}. The closest algorithm to a scale-free \textsc{FTRL} in the
+existing literature is the \textsc{AdaGrad FTRL}
+algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a regularizer on each
+coordinate of the form
 \begin{equation*}
@@
 knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
-other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
-norm of the gradient vectors. Identical considerations hold for the
+other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
+the norm of the gradient vectors. Identical considerations hold for the
 \textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
@@
 
-When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T} \|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the bound.
-On the other hand, when $K$ is bounded, the second term is a constant and we can choose the optimal multiple of the regularizer.  We
-choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
-optimize $\lambda$.  The result of the optimization is
+When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T}
+\|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the
+bound. On the other hand, when $K$ is bounded, the second term is a constant and
+we can choose the optimal multiple of the regularizer.  We choose $R(w) =
+\lambda f(w)$ where $f$ is a $1$-strongly convex function and optimize
+$\lambda$.  The result of the optimization is
 Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}.
-%; the proof is in~\ref{section:solo-ftrl-proofs}.
-%It is similar to Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}.
-The scaling however is different in the two corollaries.
-In Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.
 
@@
 The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
-(Lemma~\ref{lemma:useful}).  Related and weaker inequalities were proved
-in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}.
-The main property of this inequality is that on the right-hand side $C$ does
-\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.  We will also use the
-well-known technical Lemma~\ref{lemma:sum-of-square-roots-inverses}.
+(Lemma~\ref{lemma:useful}).  Related and weaker inequalities
+(Lemma~\ref{lemma:sum-of-square-roots-inverses}) were proved
+in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}. The
+main property of this inequality is that on the right-hand side $C$ does
+\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.
 
@@
 %
-For completeness, a proof of Lemma~\ref{lemma:sum-of-square-roots-inverses} is in
+For completeness, a proof of Lemma~\ref{lemma:sum-of-square-roots-inverses} is
 in~\ref{section:solo-ftrl-proofs}.

### 2016-09-27 revision/conclusions.tex
Fix conclusions section

@@
 
-We have designed two scale-free algorithms based on \textsc{Follow The
-Regularizer Leader}.  Their regret is $O \left(\sqrt{\sup_{v \in K} f(v)
-\sum_{t=1}^T \|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly
-convex function with respect to a norm $\|\cdot\|$ defined on the decision set
-and where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.
+We have designed a scale-free algorithm based on \textsc{Follow The Regularizer
+Leader}. Its regret is $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
+\min{\sqrt{T}, D} \max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any
+competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex function
+defined on $K$ and $D$ is the diameter of $K$. The result makes sense even when
+the decision set $K$ is unbounded.
 
 A similar, but weaker result holds for a scale-free algorithm based on
-\textsc{Mirror Descent}. However, we have also shown this algorithm to be strictly
-weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
-we gave examples of regularizers for which the scale-free version of
+\textsc{Mirror Descent}. However, we have also shown that this algorithm is
+strictly weaker than algorithms based on \textsc{Follow The Regularizer Leader}.
+Namely, we gave examples of regularizers for which the scale-free version of
 \textsc{Mirror Descent} has $\Omega(T)$ regret or worse.
 
-We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
-lower bound on the regret of any algorithm for any decision set with diameter $D$.
+We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$ lower
+bound on the regret of any algorithm for any decision set with diameter $D$.
 
-We have shown that the scale-free \textsc{FTRL}-based algorithm \textsc{SOLO FTRL} has
-regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
-\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
-$f$ is any non-negative $1$-strongly convex function defined on the decision
-set.  The latter result makes sense even when the decision set $K$ is
-unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
-regret depends quadratically on the norm of the competitor $\|u\|_2$. There
-exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
+Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the regret of
+\textsc{SOLO FTRL} depends quadratically on the norm of the competitor
+$\|u\|_2$. There exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
 McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
@@
 dependency.  These algorithms assume an a priori bound on the norm of the loss
-vectors.  This leads to another open problem:
-%
-\begin{quotation}
-\noindent
-\emph{Is there a scale-free algorithm for the decision set $\R^d$, the regret of
-which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
-\end{quotation}
+vectors. It remains an open problem to design algorithms that have sub-quadratic
+dependency on $\norm{u}_2$ and are adaptive to the scale of the loss vectors.

### 2016-09-27 revision/introduction.tex
Fix conclusions section

@@
 Our proof techniques rely on new homogeneous
-inequalities (Lemmas~\ref{lemma:recurrence-solution}, \ref{lemma:useful})
+inequalities (Lemmas~\ref{lemma:useful}, \ref{lemma:recurrence-solution})
 which might be of independent interest.

### 2016-09-27 revision/mirror-descent.tex
Fix conclusions section

@@
 Per-coordinate version of \textsc{Scale-Free MD} with regularizer $R(w) =
-\frac{1}{2}\norm{w}_2^2$ is exactly the same as the diagonal version of
-\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011}.
+\frac{1}{2}\norm{w}_2^2$ is exactly the same algorithm as the diagonal version
+of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011}.
 
@@
 sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
-the first vector of the standard orthnormal basis.  On such a sequence,
-\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
-identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
+the first vector of the standard orthnormal basis.  On such a sequence, both
+versions of \textsc{AdaGrad MD} as well as  \textsc{Scale-Free MD} are identical
+to gradient descent with step size $1/\sqrt{t}$, i.e., they are identical
+Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (\textsc{GIGA})
+algorithm~\cite{Zinkevich-2003}. Hence the lower bound applies to all these
 algorithms.

### 2016-09-27 revision/introduction.tex
Reword one sentence in the intro

@@
 negative results for existing popular variants of \textsc{MD}. We show two
-examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
-regret (on sequences of loss vectors of unit norm) if the Bregman divergence
-associated with the regularizer is unbounded. These results indicate that
-\textsc{FTRL} is superior to \textsc{MD}.
+examples of decision sets and sequences of loss vectors of unit norm on which
+these variants of \textsc{MD} have $\Omega(T)$ regret.  These results indicate
+that \textsc{FTRL} is superior to \textsc{MD}.

### 2016-09-28 revision/conclusions.tex
Add references to NIPS 2016 papers

@@
 We have designed a scale-free algorithm based on \textsc{Follow The Regularizer
-Leader}. Its regret is $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
-\min{\sqrt{T}, D} \max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any
-competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex function
-defined on $K$ and $D$ is the diameter of $K$. The result makes sense even when
-the decision set $K$ is unbounded.
+Leader}. Its regret with respect to any competitor $u$ is
+$$
+O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}
++ \min\{\sqrt{T}, D\} \max_{t=1,2,\dots,T} \|\ell_t\|_* \right) \; ,
+$$
+where $f$ is any non-negative $1$-strongly convex
+function defined on the decision set and $D$ is the diameter of the decision
+set. The result makes sense even when the decision set is unbounded.
 
@@
 $\|u\|_2$. There exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
-McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
-have only a $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
-dependency.  These algorithms assume an a priori bound on the norm of the loss
-vectors. It remains an open problem to design algorithms that have sub-quadratic
-dependency on $\norm{u}_2$ and are adaptive to the scale of the loss vectors.
+McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014,
+Orabona-Pal-2016} that have only a $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or
+$O(\|u\|_2 \log \|u\|_2)$ dependency.  These algorithms assume an a priori bound
+on the norm of the loss vectors. Recently, an algorithm that adapts to norms of
+loss vectors and has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency was
+proposed~\cite{Cutkosky-Boahen-2016}. However, the trade-off between the
+dependency on $\norm{u}_2$ and the adaptivity to the norms of the loss vectors
+still remains to be explored.

### 2016-09-28 revision/introduction.tex
Add references to NIPS 2016 papers

@@
 Scale-free versions of \textsc{MD} include \textsc{AdaGrad
-MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad MD algorithm has a
-non-trivial regret bounds only when the Bregman divergence associated with the
-regularizer is bounded. In particular, since a bound on the Bregman divergence
-implies that the decision set is bounded, the regret bound for \textsc{AdaGrad
-MD} and is vacuous for unbounded sets. In fact, as we show in
+MD}~\cite{Duchi-Hazan-Singer-2011}. However, the \textsc{AdaGrad MD} algorithm
+has a non-trivial regret bounds only when the Bregman divergence associated with
+the regularizer is bounded. In particular, since a bound on the Bregman
+divergence implies that the decision set is bounded, the regret bound for
+\textsc{AdaGrad MD} is vacuous for unbounded sets. In fact, as we show in
 Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and

### 2016-10-07 revision/introduction.tex
minor fixes

@@
 in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers
-were analyzed in~\cite{Duchi-Hazan-Singer-2011} and generalized in
+were analyzed in~\cite{Duchi-Hazan-Singer-2011} and the analysis tightened in
 \cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed in
@@
 Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called
-\emph{regularizer}. The function is assumed to be strongly convex with respect
-to a given norm. Based on different regularizers different algorithms with
+\emph{regularizer}.
+%The function is assumed to be strongly convex with respect to a given norm.
+Based on different regularizers different algorithms with
 different properties can be instantiated. A summary of algorithms for OLO is
@@
 \textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the
-off-by-one issue, but it requires proximal regularizers, that in general do not
-have a simple form and even in the simple 2-norm case requires bounded domains.
+off-by-one issue, but it requires proximal regularizers. In general, proximal regularizers do not
+have a simple form and even the simple 2-norm case requires bounded domains to achieve non-vacuous regret.
 
@@
 
-We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
+We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} on bounded domains after $T$
 rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
@@
 these variants of \textsc{MD} have $\Omega(T)$ regret.  These results indicate
-that \textsc{FTRL} is superior to \textsc{MD}.
+that \textsc{FTRL} is superior to \textsc{MD} in a worst-case sense.

### 2016-10-07 revision/solo-ftrl.tex
minor fixes

@@
 other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
-the norm of the gradient vectors. Identical considerations hold for the
-\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
+the norm of the gradient vectors. Similar considerations hold for the
+\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: The
 scale-free setting of the learning rate is valid only in the bounded case.
@@
 The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
-(Lemma~\ref{lemma:useful}).  Related and weaker inequalities
-(Lemma~\ref{lemma:sum-of-square-roots-inverses}) were proved
+(Lemma~\ref{lemma:useful}).  Related and weaker inequalities, like
+Lemma~\ref{lemma:sum-of-square-roots-inverses}, were proved
 in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}. The

### 2016-10-07 revision/lower-bound.tex
minor fix

@@
 upper bound is also achieved by \textsc{FTRL} with a constant learning rate
-when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ are known
+when $\sum_{t=1}^T \|\ell_t\|_*^2$ is known
 upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is

### 2016-10-09 revision/response.tex
more fixes and finished the response file

\documentclass{article}

\usepackage{times}
\usepackage{enumerate}

\author{Francesco Orabona
\and
D\'avid P\'al}

\title{Response to the Reviewers}

\begin{document}

\maketitle

\section{Response to Reviewer 1}

We thank the Reviewer to the very detailed reviewed. We have fixed all the typos spotted by the Reviewer. In the following we reply to his concerns.

\begin{itemize}
\item First and foremost, the entire discussion of AdaFTRL---as early as in Section 3---seemed to me like a complete distraction from the main meat of the paper, which is the SOLO-FTRL algorithm and how it is superior over MD.  AdaFTRL is quite unintuitive, its guarantees are inferior to those of SOLO-FTRL, and it is also much harder to implement.

\emph{Following the suggestions of the Reviewer, we have moved the entire description of AdaFTRL to the Appendix. Hence, now the main focus of the paper is only on SOLO FTRL and SCALE-FREE MD.}

\item Furthermore, the Scale Free MD algorithm is not really new (it is simply online MD with adaptive step-size) and steals the focus from SOLO-FTRL.  I believe the authors wished to discuss this algorithm only for appraising the superiority of SOLO-FTRL, but this is not made clear enough in their current presentation, and in fact, only adds to the confusion.

\emph{Given the focus of scale-free algorithms of our submission, we believe it is natural to show meta-algorithms that achieve scale-freeness using the two major design strategies for online learning algorithms. In this sense, it is also important to point out the shortcomings of one against the other. Hence, we believe that the description of SCALE-FREE MD is essential in the paper and we prefer to leave it in the paper. Yet, following the suggestions of the Reviewer, we have heavily reorganized the section of SCALE-FREE MD, moving all the preliminaries in Section 2.}

\item Related results (and Table 1): You argue that Proximal-FTRL is "almost
scale-free" due to the "off-by-one" issue.  However, the proximal version of
FTRL does not suffer from this issue; see, e.g., Theorem 2 in "A Survey of
Algorithms and Analysis for Adaptive Online Learning" by McMahan.  Is the
Proximal-FTRL algorithm actually scale-free?  Also, why do you claim (in Table
1) that Proximal-FTRL is applicable only w.r.t. the 2-norm, while Theorem 2 in
McMahan's survey applies to any strongly convex regularization w.r.t. any norm?
Finally, why is the Optimistic MD algorithm scale-free, and with what
predictions?

\emph{Proximal-FTRL is indeed scale-free, thanks for pointing out the error. However, while it is possible to apply Proximal-FTRL to any strongly convex regularizer, McMahan (2014) does not provide a general way to construct
proximal regularizers. To the best of our knowledge, the only proximal
regularizer we are aware is based on the 2-norm.
References to Optimistic MD have been removed because not precise.
The discussion of related work and the
table, in Section 1, has been changed to reflect all these considerations.}

\item Section 1.2, 1st par, "SOLO FTRL can be viewed as the "correct" scale-free
version of AdaGrad FTRL generalized to arbitrary strongly convex regularizers":
I disagree with this claim; SOLO FTRL is very different from AdaGrad---while
the latter maintains a "matrix learning rate" (diagonal/non-diagonal), SOLO
FTRL maintains a scalar learning rate being adaptive to the scale of the
losses.  Also, how is SOLO FTRL similar to Optimistic MD? Is there anything
"predictive" in the former algorithm?

\emph{We have changed the text in "SOLO FTRL can be viewed as the "correct"
scale-free version of the diagonal version of AdaGrad FTRL generalized to
arbitrary strongly convex regularizers".  Indeed, as explained in Section 2.4,
the diagonal version of AdaGrad is nothing else than running d-copies of a FTRL
algorithm, one for each coordinate. See also M. Streeter, H. B. McMahan, Less
Regret via Online Conditioning, arXiv:1002.4862, 2010.}

\item When discussing related results, I think you should also mention standard
doubling tricks (that are commonly used to make algorithms adaptive) and why
one should prefer your algorithms.

\emph{As far as we know, there is no known doubling trick to have scale-freeness.
The discussion on doubling trick and why our solution has to be preferred is already present in Section 3.}

\item Section 5 is quite messy: part of it belongs to the preliminaries (the
description of MD, its regret bound, the discussion on AdaGrad/GIGA), and the
other part feels like it's there in order to motivate the SOLO-FTRL algorithm
(after the latter was presented in Section 4).  I think the authors have to
rearrange the material and make the purpose of this section more clear.

\emph{Thanks for pointing this out. We have reorganized the section, moving the
  preliminaries to Section 2.}

\end{itemize}

\subsection{Minor comments:}
\begin{itemize}
\item On first reading, the last sentence of the abstract did not make any sense to me. Consider rephrasing/removing it.

\emph{We have rephrased it.}

\item When defining "scale-free" algorithms (bottom of page 2), it is important to remark that the same decisions are produced when the same algorithm *with the same set of parameters* is used on scaled loss sequences. (I assume you do not intend to allow one to magically tune the step-size so as to match the scale of the losses.)

\emph{We added "with the same parameters".}

\item Page 3, 1st par, "the general version with time-varying regularizers was analyzed in [28]": RDA with time-varying regularization was analyzed already in the AdaGrad paper (and even before that).

\emph{Fixed: Orabona et al. (2014) only tightened the analysis in the AdaGrad paper.}

\item Section 1.2, last par, "These results indicate that FTRL is superior to MD": in what sense? Please make a softer statement and give more details.

\emph{Added "in a worst-case sense".}

\item Please mention explicitly why the SOLO FTRL and ScaleFree-MD algorithms are indeed scale free.

\emph{Added "In fact, with this choice of $\delta$ it is easy to see that the prediction $w_t$ in line 4 of Algorithm~\ref{algorithm:ftrl-varying-regularizer} would be independent of the scaling of the $\ell_t$." in Section 3 and "As for \textsc{SOLO FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$ that are scale-free." in Section 4.}

\item Section 4 is missing an exposition and dives right into a technical discussion.

\emph{Section 4 is now completely reorganized, see comments above.}

\item In the short discussion after Theorem 2, please state explicitly what happens when K is unbounded (currently you only discuss the case where K is bounded).  This might be obvious, but it's one of the paper's main results.

\emph{Added "When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T}
\|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the
bound."}

\item I think the proof of Corollary 2 should be included in the body of the paper.

\emph{Done.}

\item Lemma 6: please add the citation ([34, Lemma 3.5]) to the lemma's definition, instead of giving it after the lemma.

\emph{Done.}

\end{itemize}

\section{Response to Reviewer 2}

Thanks for the careful reading, we have fixed all typos. Also, thanks for the more direct proof for the recurrence, we
have included it in the paper.

\end{document}

### 2016-10-09 revision/conclusions.tex
minor fixes
more fixes and finished the response file

@@
 A similar, but weaker result holds for a scale-free algorithm based on
-\textsc{Mirror Descent}. However, we have also shown that this algorithm is
-strictly weaker than algorithms based on \textsc{Follow The Regularizer Leader}.
+\textsc{Mirror Descent}. However, we have also shown this algorithm to be strictly
+weaker than algorithms based on \textsc{Follow The Regularizer Leader}.
 Namely, we gave examples of regularizers for which the scale-free version of
@@
 on the norm of the loss vectors. Recently, an algorithm that adapts to norms of
-loss vectors and has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency was
+loss vectors and has a $O(\|u\|_2 \log \|u\|_2)$ dependency was
 proposed~\cite{Cutkosky-Boahen-2016}. However, the trade-off between the

### 2016-10-09 revision/mirror-descent.tex
minor fixes
more fixes and finished the response file

@@
 In this section, we analyze scale-free version of \textsc{Mirror Descent}.
-Our algorithm uses regularizer
+Our algorithm uses the regularizer
+%
 \begin{equation}
 \label{equation:scale-free-mirror-descent}
-R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
+R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2},
 \end{equation}
+%
 where $R:K \to \R$ an arbitrary strongly convex function.
+As for \textsc{SOLO FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$ that are scale-free.
 We call the resulting algorithm \textsc{Scale-Free MD}.
@@
 The second counter-example is stated as
-Theorem~\ref{theorem:second-counter-example} below; proof is
+Theorem~\ref{theorem:second-counter-example} below; our proof is
 in~\ref{section:mirror-descent-proofs}.  The decision set is the

### 2016-10-09 revision/response.tex
small fixes

@@
 
-We thank the Reviewer to the very detailed reviewed. We have fixed all the typos spotted by the Reviewer. In the following we reply to his concerns.
+We thank the Reviewer for the very detailed reviewed. We have fixed all the typos spotted by the Reviewer. In the following we reply to his concerns.
 
@@
 
-\emph{Following the suggestions of the Reviewer, we have moved the entire description of AdaFTRL to the Appendix. Hence, now the main focus of the paper is only on SOLO FTRL and SCALE-FREE MD.}
+\emph{Following the suggestions of the Reviewer, we have moved the entire description of AdaFTRL to the Appendix.
+Hence, now the main focus of the paper is only on SOLO FTRL and SCALE-FREE MD.
+We still believe it is an important contribution to show that AdaHedge can be generalized to arbitrary regularizers.}
 
@@
 In this sense, it is also important to point out the shortcomings of one against the other.
-Hence, we believe that the description of SCALE-FREE MD is essential in the paper and we prefer to leave it in the paper.
+Hence, we believe that the description of SCALE-FREE MD is essential in the paper and we prefer to keep it in the paper.
 Yet, following the suggestions of the Reviewer, we have heavily reorganized the section of SCALE-FREE MD, moving all the preliminaries in Section 2.}
@@
 proximal regularizers. To the best of our knowledge, the only proximal
-regularizer we are aware is based on the 2-norm.
+regularizer we are aware of is based on the 2-norm.
 References to Optimistic MD have been removed because not precise.
@@
 FTRL maintains a scalar learning rate being adaptive to the scale of the
-losses.  Also, how is SOLO FTRL similar to Optimistic MD? Is there anything
-"predictive" in the former algorithm?
+losses.
 
@@
 \emph{As far as we know, there is no known doubling trick to have scale-freeness.
-The discussion on doubling trick and why our solution has to be preferred is already present in Section 3.}
+The discussion on a possible solution through a doubling trick and why our solution has to be preferred is already present in Section 3.}
 
@@
 
-\emph{Fixed: Orabona et al. (2014) only tightened the analysis in the AdaGrad paper.}
+\emph{Fixed: Orabona et al. (2014) only tightened the analysis of the AdaGrad paper.}
 
@@
 
-\emph{Added "In fact, with this choice of $\delta$ it is easy to see that the prediction $w_t$ in line 4 of Algorithm~\ref{algorithm:ftrl-varying-regularizer} would be independent of the scaling of the $\ell_t$." in Section 3 and "As for \textsc{SOLO FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$ that are scale-free." in Section 4.}
+\emph{Added "In fact, with this choice of $\delta$ it is easy to see that the predictions $w_t$ in line 4 of Algorithm~1 would be independent of the scaling of the $\ell_t$." in Section 3 and "As for \textsc{SOLO FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$ that are scale-free." in Section 4.}
 

### 2016-10-09 revision/mirror-descent.tex
Final fixes

@@
 \label{equation:scale-free-mirror-descent}
-R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2},
+R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; ,
 \end{equation}
 %
-where $R:K \to \R$ an arbitrary strongly convex function.
-As for \textsc{SOLO FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$ that are scale-free.
-We call the resulting algorithm \textsc{Scale-Free MD}.
+where $R:K \to \R$ an arbitrary strongly convex function.  As for \textsc{SOLO
+FTRL}, it is easy to see that such regularizer gives rise to predictions $w_t$
+that are scale-free.  We call the resulting algorithm \textsc{Scale-Free MD}.
 Similar to \textsc{SOLO FTRL}, the regularizer
-\eqref{equation:scale-free-mirror-descent} does not uniquely define
-the \textsc{MD} minimizer
-$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
-w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
-This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
-zero. In this case, we define
-$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
-$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
-Similarly, $w_1 = \argmin_{w \in K} R(w)$.
+\eqref{equation:scale-free-mirror-descent} does not uniquely define the
+\textsc{MD} minimizer $w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w
+\rangle + \Breg_{R_t}(w, w_t) \right)$ when $\sqrt{\sum_{i=1}^t
+\|\ell_i\|_*^2}$ is zero.  This happens when the loss vectors $\ell_1, \ell_2,
+\dots, \ell_t$ are all zero. In this case, we define $w_{t+1} = \argmin_{w \in
+K} R(w)$ which agrees with $w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a
+\Breg_{R}(w, w_t)$.  Similarly, $w_1 = \argmin_{w \in K} R(w)$.
 

### 2016-10-09 revision/solo-ftrl.tex
more fixes and finished the response file
small fixes
Final fixes

@@
 This kind of regularizer would yield a scale-free algorithm \emph{only} for
-$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
-becomes vacuous for such setting in the unbounded case. In fact, it requires
-$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
-knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
-other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
-the norm of the gradient vectors. Similar considerations hold for the
-\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: The
-scale-free setting of the learning rate is valid only in the bounded case.
+$\delta=0$. In fact, with this choice of $\delta$ it is easy to see that the
+predictions $w_t$ in line 4 of
+Algorithm~\ref{algorithm:ftrl-varying-regularizer} would be independent of the
+scaling of the $\ell_t$.  Unfortunately, the regret bound
+in~\cite{Duchi-Hazan-Singer-2011} becomes vacuous for such setting in the
+unbounded case. In fact, it requires $\delta$ to be greater than $\|\ell_t\|_*$
+for all time steps $t$, requiring knowledge of the future (see Theorem~5
+in~\cite{Duchi-Hazan-Singer-2011}). In other words, despite of its name,
+\textsc{AdaGrad FTRL} is not fully adaptive to the norm of the gradient
+vectors. Similar considerations hold for the \textsc{FTRL-Proximal} in
+\cite{McMahan-Streeter-2010,McMahan-2014}: The scale-free setting of the
+learning rate is valid only in the bounded case.
 

### 2016-12-14 first-submision/preliminaries.tex
Move old files into a separate directory

--- 2016-09-15 revision/preliminaries.tex
@@
 For completeness, we include it in~\ref{section:definitions-proofs}.
-
-\subsection{Generic Mirror Descent with Varying Regularizer}
-
-\begin{algorithm}[t]
-\caption{\textsc{Mirror Descent with Varying Regularizer}}
-\label{algorithm:mirror-descent-varying-regularizer}
-\begin{algorithmic}[1]
-\REQUIRE Non-empty closed convex set $K \subseteq V$
-\STATE Choose a regularizer $R_0:K \to \R$
-\STATE $w_1 \leftarrow \argmin_{w \in K} R_0(w)$
-\FOR{$t=1,2,3,\dots$}
-\STATE Predict $w_t$
-\STATE Observe $\ell_t \in V^*$
-\STATE Choose a regularizer $R_t:K \to \R$
-\STATE $w_{t+1} \leftarrow \argmin_{w \in K} \left( \langle \ell, w \rangle + \Breg_{R_t}(w, w_t) \right)$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
-but quite different in details. The algorithm is stated as
-Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
-parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
-\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
-loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
-not differentiable,\footnote{Note that this can happen even when $R_t$ is a
-restriction of a differentiable function defined on a superset of $K$.  If $K$
-is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
-If $K$ is a subset of an affine subspace of a dimension smaller than the
-dimension of $V$, then $R_t$ fails to be differentiable everywhere.} the
-Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
-u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
-$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
-of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
-$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
-\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
-in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
-completeness we give a proof in~\ref{section:mirror-descent-proofs}.
-
-\begin{lemma}[Regret of \textsc{MD}]
-\label{lemma:mirror-descent-regret}
-Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
-$u \in K$,
-$$
-\Regret_T(u)
-\le
-\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
-$$
-\end{lemma}
-
-Several existing algorithms are special cases of \textsc{MD}.  For example,
-\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
-\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
-regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
-is a (constant) learning rate. Note that this algorithm is \emph{not}
-scale-free.
-
-The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
-Descent} defined by the recurrence
-$$
-w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
-$$
-where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
-$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
-an instance of \textsc{MD} with regularizer $R_t(w) =
-\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
-Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
-Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.

### 2016-12-14 first-submission/ada-ftrl-proofs.tex
Rename directory

--- 2015-12-16 ada-ftrl-proofs.tex
@@
 \begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
-We prove the lemma by induction on $T$.  The base case $T=0$ is trivial. For $T
-\ge 1$, we have
+We prove the lemma by induction on $T$.  The base case $T=0$ is trivial.
+For $T \ge 1$, let $f(x) = x + \min\{a_T, a_T^2/x\}$. We have
+\begin{align*}
+\Delta_T
+& \le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\} \\
+& = f \left( \Delta_{T-1} \right) \\
+& \le f \left( \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \right) \\
+& = \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}  \; ,
+\end{align*}
+where the second inequality follows from the inductive hypothesis $\Delta_{T-1} \le
+\sqrt{3 \sum_{t=1}^{T-1} a_t^2}$ and that $f(x)$ is non-positive on $[-\infty,
+0)$ and non-negative increasing on $[0,\infty)$. It remains to prove that
 $$
-\Delta_T
-\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
-\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
-$$
-where the second inequality from the induction hypothesis and the fact that the
-function $f(x) = x + \min\{a_T, a_T^2/x\}$ is increasing on $[0,\infty)$.  It
-remains to prove that
-$$
-\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
+\sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
 \le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
@@
 $$
-z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
+z\sqrt{3} + \min\left\{1, \frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
 $$

### 2016-12-14 first-submission/introduction.tex
Rename directory

--- 2015-12-28 introduction.tex
@@
 \textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  These algorithms fail to
-be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and also our
+be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and the
 discussion in Section~\ref{section:solo-ftrl}.
@@
 arbitrary strongly convex regularizers, similar to \textsc{Optimistic
-MD}~\cite{Rakhlin-Sridharan-2013}.  The three algorithms are presented and
-analyzed in Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
+MD}~\cite{Rakhlin-Sridharan-2013}.  The three algorithms are presented in
+Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
 \ref{section:mirror-descent} respectively.
